### 2019-06-20 latex/paper_main.tex
Add Latex files

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasmathematics} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\graphicspath{{figures/}}

\title{Learning complex models with invertible neural networks: a likelihood-free Bayesian approach}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[1]{Stefan T. Radev}
\author[1]{Ulf K. Mertens} 
\author[1]{Andreas Voss}
\author[2]{Lynton Ardizzone}
\author[2]{Ullrich Köthe}

\affil[1]{Institute of Psychology, Heidelberg University, Hauptstr. 47-51, 69117 Heidelberg, Germany}
\affil[2]{Heidelberg Collaboratory for Image Processing (HCI), Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany}

% Please give the surname of the lead author for the running footer
\leadauthor{Radev}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{This research was supported by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation; grant number GRK 2277 "Statistical Modeling in Psychology")}

\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: stefan.radev\@psychologie.uni-heidelberg.de}

% At least three keywords are required at submission. Please provide two to five keywords, separated by the pipe symbol.
\keywords{Deep learning $|$ Invertible networks $|$ Bayesian inference $|$ Parameter estimation $|$ Stochastic models}

\begin{abstract}
Parametric models of complex processes are ubiquitous throughout the sciences. As the processes under study and the models describing them become increasingly complex, parameter estimation with standard Bayesian and frequentist methods can quickly become intractable. To address this, we propose a novel method for likelihood-free inference based on invertible neural networks. The method is capable of performing fast full Bayesian inference on large amounts of data by training the networks on simulated data and learning to invert the model under study. The method is independent of particular data formats, as it includes a summary network trained to embed the observed data into fixed-size vectors in a data-driven way. This makes the method applicable to various scenarios where standard inference techniques fail. We demonstrate the utility of the method on a toy model with known analytic posterior and on example models from population dynamics, epidemiology, cognitive science and genetics. We argue for a general framework for building reusable parameter estimation machines for potentially any process model from which simulations can be obtained.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about probabilistic phenomena in a probabilistic world. In their most abstract form, mathematical models are specified by a set of latent parameters $\boldsymbol{\theta}$ and a generative model $q$ mapping parameters to manifest quantities $\boldsymbol{x}$:

\begin{align*}
\boldsymbol{x}= q(\boldsymbol{\theta}) \numberthis \label{eqn:1} 
\end{align*}

While $q$ can represent an arbitrarily complex process by an arbitrarily complicated expression, its functional form is usually guided by a well-founded theoretical framework. For instance, it could be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step mechanism dictating the rate of gene expression in certain cells. Thus, it is only through theoretical embedding that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model. Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical}, epidemiology \cite{hethcote2000mathematics}, just to name a few.

Once a mathematical model has been formulated, the next step consists of fitting the model to experimental or observational data and recovering the parameters of interest. However, estimating the parameters of a mathematical model can quickly become one of the most tenacious challenges in applications to real-world problems. It is also one of the most important ones to be tackled, since without reliable parameter estimation methods, it is impossible to test the utility of a model, regardless of its sophistication or theoretical appeal. Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly. However, due to noise, inherent stochasticity or loss of information, the inverse usually does not exist, so researchers need to resort to the sophisticated frameworks of Bayesian or frequentist inference.

Moreover, as mathematical models and processes under description become increasingly complex, parameter estimation and model selection can quickly become intractable with standard Bayesian and frequentist method. Complex models specified by a generative stochastic mechanism do not always provide a closed-form solution for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}. This poses great difficulties for Bayesian and frequentist methods alike, since both depend explicitly on the numerical evaluation of a likelihood function as a proxy for assessing model fit to data. Even if a likelihood function is available, inference could be prohibitively slow for real-world applications. Therefore, the need for powerful and reliable likelihood-free estimation methods arises naturally.

Likelihood-free methods aim at bypassing the intractability problem by resorting to a simulation-based approach to inference and model selection \cite{palestro2018likelihood}. A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate}. ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $\pi(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters. Thus, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.

Performing approximate inference comes at the cost of incurring additional approximation error, which accumulates on top of the irreducible estimation error. Within the context of approximate inference, the most common manifestations of approximation error include: \textit{i)} imprecise form of the posterior; \textit{ii)} imprecise posterior moments; \textit{iii)} under- or overestimation of uncertainty. Different approximation methods usually involve multiple trade-offs between minimizing approximation error and keeping computational time within reasonable bounds \cite{frazier2018asymptotic, palestro2018likelihood}.

Recently, ideas from machine learning and deep learning research have entered the field of likelihood-free inference in an attempt to overcome some of the shortcomings of traditional methods \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}. The most common approach has been to cast the problem of parameter estimation as a supervised learning task. In this setting, a large dataset of the form $\boldsymbol{D} = \{h(\boldsymbol{x}^{(i)}), \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ is created by repeatedly sampling from $\pi(\boldsymbol{\theta})$ and simulating an artificial dataset $\boldsymbol{x}$ by running $q(\boldsymbol{\theta})$ with the sampled parameters. Usually, the dimensionality of the simulated data is reduced by computing summary statistics  with a fixed summary function $h(\boldsymbol{x})$. Then, a supervised learning algorithm $f(h(\boldsymbol{x});\boldsymbol{\phi}) = \widehat{\boldsymbol{\theta}}$ with learnable parameters $\boldsymbol{\phi}$ (e.g., linear regression, random forest, neural network) is trained on the simulated data to later output an estimate of the true data generating parameters. Thus, $f(h(\boldsymbol{x});\boldsymbol{\phi})$ essentially attempts to “learn” the intractable inverse model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$.

Inspired by previous machine learning approaches, the current work proposes a novel and universal likelihood-free method capable of performing full Bayesian inference on any mathematical process model from which simulations can be obtained. It treats parameter inference as a task of inverting a generative model and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}. The method integrates two separate deep neural networks modules (detailed in the \textbf{Methods} section) trained jointly on simulated data: a \textit{summary network} and an \textit{invertible network}.

The \textit{summary network} is responsible for learning the most informative summary statistics directly from data. It should be designed to follow the functional and probabilistic symmetries inherent in the data, e.g. a permutationally invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network for time-series data, or a convolutional network for grid-like data \cite{goodfellow2016deep}. The computation of summary statistics is a crucial aspect in likelihood-free inference. Previous approaches mainly use hand-crafted summary statistics tailored to the specific application. However, in many application, it is not straightforward to settle upon a set of good summary statistics. Thus, the summary network completely eliminates the need to manually specify a fixed number of summary statistics and makes the method independent of the format or the size of the data.

The \textit{invertible network} is responsible for learning the posterior of the model parameters given the observed and summarized data. It is based on the recently developed flow-based architecture \cite{kingma2018glow, grover2018flow, dinh2016density}. Flow-based methods provide exact latent-variable inference and log-likelihood evaluation when operating at optimum. In the \textbf{Methods} section, we show that our method maximizes the posterior over model parameters directly when cast in the context of likelihood-free inference. Furthermore, flow-based methods are capable of approximating very high-dimensional distributions (e.g., images). Once trained with a sufficient amount of simulated data, our invertible network can perform full Bayesian inference on large amounts of real data from a given domain in a single pass.

The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated statistical problems in various scientific domains (see Fig.1). Moreover, the method addresses many of the limitations of previous likelihood-free methods. First, it involves no costly MCMC or rejection sampling, which makes the inference phase lightning fast, as we also make use of GPU-accelerated computation. Second, it involves no fixed summary statistics or kernels, but learns the most informative representation of the data in an end-to-end manner. Third, the method is fully Bayesian, as it directly learns the posterior over model parameters and thus allows for the quantification of uncertainty, which is a crucial requirement in parameter estimation \cite{kendall2017uncertainties, gelman2013bayesian}. Last, the trained networks can be shared and reused by multiple researches within a scientific domain, thus removing the need for wasteful computations and fitting a separate model for each and every dataset. This pooling of computational resources across researches is an important step forward in mathematical modeling, as has been recently argued \cite{mestdagh2018prepaid}.

To illustrate the utility of the new method, we first apply it to a toy Bayesian regression model with known posterior. Then, we present applications to intractable models from cognitive science, population dynamics, epidemiology, and genetics and demonstrate state-of-the art parameter recovery. The outline of the remaining manuscript is as follows: The \textbf{Methods} section introduces the main building blocks of the new method and summarizes the main steps as pseudocode. The \textbf{Results} section presents the various applications of the model to real-world research domains. Finally, the \textbf{Discussion} section lists the advantages of the current method, treats some potential pitfalls and explores future research vistas. Python code and data for all applications are freely available as Jupyter notebooks at \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN} and as a small library based on \textit{TensorFlow} \cite{abadi2016tensorflow} for creating and training custom invertible networks with GPU-support.

\section*{Methods}

\subsection*{Notation}

In the following, we denote observed or simulated univariate datasets from the mathematical model of interest as $\boldsymbol{x} = (x_{1}, x_{2},...,x_{n})$, and multivariate datasets as $\boldsymbol{X} = (\boldsymbol{x}_{1}, \boldsymbol{x}_{1},...,\boldsymbol{x}_{n})$. The parameters of a mathematical model are represented as a vector $\boldsymbol{\theta} = (\theta_{1}, \theta_{2},...,\theta_{d})$, and all trainable parameters of the invertible and summary neural networks as $\boldsymbol{\phi} = (\boldsymbol{\phi}_{inv}, \boldsymbol{\phi}_{sum})$. The number of parameters of a mathematical model will be denoted as $d$, and the number of simulated data points as $n$.

\subsection*{Deep Probabilistic Modeling}

Our method draws on major advances in modern deep probabilistic modeling, also referred to as deep generative modeling \cite{bloem2019probabilistic, kingma2018glow, ardizzone2018analyzing, kingma2014auto}. A hallmark idea in deep probabilistic modeling is to handle intractable target probability distributions by sampling from simpler distributions (e.g., Gaussian or uniform distributions) and transforming these samples via a complex non-linear, learnable transformations. Most popular deep probabilistic models entail two phases. During the \textit{training phase}, a transformation from the simple to the desired target distribution is learned by optimizing a cost function via backpropagation. During the \textit{inference phase}, samples from the target distribution are obtained by sampling from the simple distribution and applying the transformation learned during the training phase. Using this approach, recent applications of deep probabilistic models have achieved unprecedented results on extremely high-dimensional and intractable problems (e.g., complex data distributions such as natural images, music, or text).

In the context of mathematical modeling and Bayesian inference, the target distribution is the posterior distribution of model parameters $p(\boldsymbol{\theta}|\boldsymbol{x})$ capturing our uncertainty about the numerical values of parameters given empirical data. We can leverage the fact that most mathematical models are generative in nature and as such can be used to perform multiple simulations of the process of interest. By specifying a prior distribution over the model parameters $\pi(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters. Thus, at inference time, one can condition the model on observed data $\boldsymbol{x}_{obs}$ and obtain samples $\boldsymbol{\theta}_{1}, \boldsymbol{\theta}_{2},...,\boldsymbol{\theta}_{L}$  approximating the posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$ in the manner described above.

In the current work, we propose to implement and use a conditional invertible neural network (cINN) architecture. Previously, INNs have been successfully employed to model data from astrophysics and medicine\cite{ardizzone2018analyzing}. We adapt the model to suit the task of parameter estimation in the context of mathematical modeling (see Figure 1 for a full graphical illustration of the method) and develop a reusable probabilistic architecture for full Bayesian likelihood-free inference on complex mathematical models.

\subsection*{The Affine Coupling Block}

The basic building block of a cINN is the affine coupling block (ACB) \cite{ardizzone2018analyzing, kingma2018glow, dinh2016density}. Each ACB consists of four separate fully connected neural networks denoted as $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$. An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ is also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ "for free". Denoting the input vector of $f_{\boldsymbol{\phi}_{inv}}$ as $\boldsymbol{u}$ and the output vector as $\boldsymbol{v}$, it follows that $f(\boldsymbol{u}; \boldsymbol{\phi}_{inv}) = \boldsymbol{v}$ and $f^{-1}(\boldsymbol{v}; \boldsymbol{\phi}_{inv}) = \boldsymbol{u}$. Invertibility is achieved by splitting the input vector into two parts $\boldsymbol{u} = (\boldsymbol{u}_{1}, \boldsymbol{u}_{2})$ and performing the following operations on the split input:
\begin{align*} 
\boldsymbol{v}_{1} &= \boldsymbol{u}_{1} \odot exp(s_{1}(\boldsymbol{u}_{2})) + t_{1}(\boldsymbol{u}_{2}) \numberthis \label{eqn:2}  \\ 
\boldsymbol{v}_{2} &= \boldsymbol{u}_{2} \odot exp(s_{1}(\boldsymbol{v}_{1})) + t_{1}(\boldsymbol{v}_{1}) \numberthis \label{eqn:3} 
\end{align*}
The outputs $\boldsymbol{v} = (\boldsymbol{v}_{1}, \boldsymbol{v}_{2})$ are then concatenated again and passed to the next ACB. The inverse operation is given by:
\begin{align*} 
\boldsymbol{u}_{2} &= (\boldsymbol{v}_{2} - t_{2}(\boldsymbol{v}_{1})) \odot exp(-s_{2}(\boldsymbol{v}_{1})) \numberthis \label{eqn:4}  \\ 
\boldsymbol{u}_{1} &= (\boldsymbol{v}_{1} - t_{1}(\boldsymbol{u}_{2})) \odot exp(-s_{1}(\boldsymbol{u}_{2})) \numberthis \label{eqn:5} 
\end{align*}
An additional property of this design, which becomes relevant later for optimization, is that the operations of the ACB have tractable, and cheaply computable Jacobians (strictly upper or lower triangular matrices). Furthermore, the internal networks $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$ can be represented by arbitrarily complex neural networks, which themselves need not be invertible, since they are only ever evaluated in the forward direction during both the forward and the inverse pass through the ACB. To ensure that the model is powerful enough to represent complicated distributions, we chain multiple ACBs, so that the output of each ACB becomes the input of the next. In this way, the whole chain remains invertible from the first input to the last output and can be viewed as a single function parameterized by trainable parameters $\boldsymbol{\phi}_{inv}$.

In our applications, the input to the first ACB is the parameter vector $\boldsymbol{\theta}$, and the output of the final ACB, denoted hitherto as $\boldsymbol{z}$, is encouraged to follow a $d$-dimensional spherical Gaussian via optimization (described in detail later), that is, $p(\boldsymbol{z}) = \mathcal{N}_{d}(\boldsymbol{z}|\boldsymbol{0},\boldsymbol{I})$. Fixed permutation matrices are used before each ACB to ensure that each axis of the latent space encodes information from all components of $\boldsymbol{\theta}$. In order to take into account the observed data $\boldsymbol{x}$, each of the internal networks of each ACB is augmented to take $\boldsymbol{x}$ as an additional input - $s_{1}(\cdot,\boldsymbol{x}), s_{2}(\cdot,\boldsymbol{x}), t_{1}(\cdot,\boldsymbol{x}), t_{2}(\cdot,\boldsymbol{x})$ - so a complete pass through the entire invertible chain can be expressed as:
\begin{align*} 
f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi}_{inv}) = \boldsymbol{z} \numberthis \label{eqn:6}
\end{align*}
together with the inverse operation:
\begin{align*} 
f^{-1}(\boldsymbol{z};\boldsymbol{x},\boldsymbol{\phi}_{inv}) = \boldsymbol{\theta} \numberthis \label{eqn:7}
\end{align*}

This process can be interpreted as follows: the forward pass maps data-generating parameters to $\boldsymbol{z}$-space using conditional information of $\boldsymbol{x}$, while the inverse pass maps data points from $\boldsymbol{z}$-space to the data-generating parameters of interest using the same conditional information provided by the data. In the next section, we describe the optimization procedure used to match the outputs of $f^{-1}(\boldsymbol{z};\boldsymbol{x},\boldsymbol{\phi}_{inv})$ to the posterior $p(\boldsymbol{\theta}|\boldsymbol{x})$.

\subsection*{Summary Network}
Since in practice the conditioning data set $\boldsymbol{x}$ can have variable number of input points (e.g., trial sizes, time points) and exhibit various redundancies, the cINN can profit from some form of dimensionality reduction applied to the data. Ideally, we want to avoid hand-crafted summary statistics, and instead learn the most informative summary statistics directly from data. Therefore, instead of feeding the raw simulated (observed) data to each ACB, we pass the data through an additional summary network to obtain a fixed-sized vector of learned summary statistics $\tilde{\boldsymbol{x}} = h(\boldsymbol{x};\boldsymbol{\phi}_{sum})$ and learn the parameters of the summary network $h$ jointly with those of the cINN chain via backpropagation. Thus, the current method remains completely end-to-end and is capable of generalizing to data sets of variable input size and structure.

\subsection*{Maximum Likelihood Loss}
Broadly speaking, the goal of maximum likelihood estimation is to find a set of parameters which maximize the probability of the data under a parametric model. In our case, we are interested in maximizing the expectation over possible neural network parameters with respect to the parameters of the mathematical model:

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[scale=0.1]{figure1.png}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:frog}
\end{figure}

\subsection*{Digital Figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.

\subsection*{Supporting Information Appendix (SI)}

Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.  PNAS will publish SI uncomposed, as the authors have provided it.  Additional details can be found here: \href{https://www.pnas.org/page/authors/submission}{policy on SI}.  For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.  The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.  Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.

Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

\subsubsection*{SI Datasets}

Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow % Display the acknowledgements section

% Bibliography
\bibliography{references}

\end{document}

### 2019-06-20 viz.py

@@ -181,9 +181,14 @@ def plot_sbc(model, n_samples, X_test, theta_test, param_names, figsize=(15, 5),
 
     # Compute ranks (using broadcasting)    
     ranks = np.sum(theta_samples < theta_test, axis=0)
+
+    # Plot hisograms
     for j in range(len(param_names)):
-        sns.distplot(ranks[:, j], kde=False, ax=axarr[j])
+        sns.distplot(ranks[:, j], kde=False, ax=axarr[j], rug=True)
         axarr[j].set_title(param_names[j])
+        axarr[j].spines['right'].set_visible(False)
+        axarr[j].spines['top'].set_visible(False)
+
     f.tight_layout()
     
     # Show, if specified

### 2019-06-21 latex/paper_main.tex
Add figure

@@
 
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{subfig}
 \graphicspath{{figures/}}
+\DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
 
@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers.}
+\significancestatement{Describing complex stochastic processes with parametric  models lies at the heart of science.
+Simulating models given a set of parameters is relatively easy with the aid of modern computers, but inferring model parameters from observed data can often be a challenging endeavor.
+We combine recent advances in deep learning and Bayesian inference into a powerful method for building reusable parameter estimation networks applicable to various types of models and data encountered in different research fields.}
 
@@
 \dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about probabilistic phenomena in a probabilistic world.
-In their most abstract form, mathematical models are specified by a set of latent parameters $\boldsymbol{\theta}$ and a generative model $q$ mapping parameters to manifest quantities $\boldsymbol{x}$:
-
+In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a generative model $q$ mimicking the process by which manifest quantities $\boldsymbol{x}$ arise from latent parameters:
 \begin{align*}
@@
 \end{align*}
-
 While $q$ can represent an arbitrarily complex process by an arbitrarily complicated expression, its functional form is usually guided by a well-founded theoretical framework.
-For instance, it could be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step mechanism dictating the rate of gene expression in certain cells.
+For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step mechanism dictating the rate of gene expression in certain cells.
 Thus, it is only through theoretical embedding that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
@@
 
-Moreover, as mathematical models and processes under description become increasingly complex, parameter estimation and model selection can quickly become intractable with standard Bayesian and frequentist method.
+However, as mathematical models and processes under description become increasingly complex, parameter estimation and model selection can quickly become intractable with standard Bayesian and frequentist method.
 Complex models specified by a generative stochastic mechanism do not always provide a closed-form solution for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}.
 This poses great difficulties for Bayesian and frequentist methods alike, since both depend explicitly on the numerical evaluation of a likelihood function as a proxy for assessing model fit to data.
-Even if a likelihood function is available, inference could be prohibitively slow for real-world applications.
+Even if a likelihood function is available in closed-form, inference may be prohibitively slow for real-world applications.
+In this case, enforcing simplifying distributional assumptions (i.e., independence or Gaussian assumptions) on the likelihood can increase the speed of inference, but can also lead to model misspecifications and dramatically incorrect estimates.
 Therefore, the need for powerful and reliable likelihood-free estimation methods arises naturally.
 
-Likelihood-free methods aim at bypassing the intractability problem by resorting to a simulation-based approach to inference and model selection \cite{palestro2018likelihood}.
-A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate}.
+Likelihood-free methods aim at bypassing the above problems by resorting to a simulation-based approach to parameter estimation and model selection \cite{palestro2018likelihood, turner2014generalized}.
+A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate, csillery2010approximate}.
 ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $\pi(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters.
@@
 Once trained with a sufficient amount of simulated data, our invertible network can perform full Bayesian inference on large amounts of real data from a given domain in a single pass.
+
+\begin{figure}
+    \centering
+    \subfloat[Training phase]{{\includegraphics[width=.99\linewidth]{training.png}}}
+    \newline
+    \subfloat[Inference phase]{{\includegraphics[width=.99\linewidth]{inference.png}}}
+    \caption{A graphic illustration of the two phases of the method.
+\textbf{(a)} During the training phase, the networks are trained on simulated data from the model; \textbf{(b)} during the inference phase, the posterior is approximated from real data.}
+    \label{fig:first}
+\end{figure}
 
@@
 \begin{align*} 
-\boldsymbol{v}_{1} &= \boldsymbol{u}_{1} \odot exp(s_{1}(\boldsymbol{u}_{2})) + t_{1}(\boldsymbol{u}_{2}) \numberthis \label{eqn:2}  \\ 
-\boldsymbol{v}_{2} &= \boldsymbol{u}_{2} \odot exp(s_{1}(\boldsymbol{v}_{1})) + t_{1}(\boldsymbol{v}_{1}) \numberthis \label{eqn:3} 
+\boldsymbol{v}_{1} &= \boldsymbol{u}_{1} \odot \exp(s_{1}(\boldsymbol{u}_{2})) + t_{1}(\boldsymbol{u}_{2}) \numberthis \label{eqn:2}  \\ 
+\boldsymbol{v}_{2} &= \boldsymbol{u}_{2} \odot \exp(s_{1}(\boldsymbol{v}_{1})) + t_{1}(\boldsymbol{v}_{1}) \numberthis \label{eqn:3} 
 \end{align*}
@@
 \begin{align*} 
-\boldsymbol{u}_{2} &= (\boldsymbol{v}_{2} - t_{2}(\boldsymbol{v}_{1})) \odot exp(-s_{2}(\boldsymbol{v}_{1})) \numberthis \label{eqn:4}  \\ 
-\boldsymbol{u}_{1} &= (\boldsymbol{v}_{1} - t_{1}(\boldsymbol{u}_{2})) \odot exp(-s_{1}(\boldsymbol{u}_{2})) \numberthis \label{eqn:5} 
+\boldsymbol{u}_{2} &= (\boldsymbol{v}_{2} - t_{2}(\boldsymbol{v}_{1})) \odot \exp(-s_{2}(\boldsymbol{v}_{1})) \numberthis \label{eqn:4}  \\ 
+\boldsymbol{u}_{1} &= (\boldsymbol{v}_{1} - t_{1}(\boldsymbol{u}_{2})) \odot \exp(-s_{1}(\boldsymbol{u}_{2})) \numberthis \label{eqn:5} 
 \end{align*}
@@
 
-\subsection*{Maximum Likelihood Loss}
-Broadly speaking, the goal of maximum likelihood estimation is to find a set of parameters which maximize the probability of the data under a parametric model.
+\subsection*{Learning the Posterior}
+The cINN learns to approximate the posterior of model parameters by optimizing  a maximum likelihood (ML) criterion.
+Broadly speaking, the goal of ML estimation is to find a set of parameters which maximize the probability of the data under a parametric model.
 In our case, we are interested in maximizing the expectation over possible neural network parameters with respect to the parameters of the mathematical model:
+\begin{align*} 
+\boldsymbol{\phi}^{*} = \argmax_{\boldsymbol{\phi}}\mathbb{E}_{\boldsymbol{\theta} \sim p(\boldsymbol{\theta}|\boldsymbol{x})}\left[p(\boldsymbol{\phi}|\boldsymbol{\theta},\boldsymbol{x})\right] \numberthis \label{eqn:8}
+\end{align*}
+Applying Bayes' rule to the posterior over all neural network parameters we obtain:
+\begin{align*} 
+p(\boldsymbol{\phi}|\boldsymbol{\theta},\boldsymbol{x}) \propto p(\boldsymbol{\theta}|\boldsymbol{x},\boldsymbol{\phi})p(\boldsymbol{\phi})  \numberthis \label{eqn:9}
+\end{align*}
+Note, that by maximizing eq.\ref{eqn:9} we are maximizing the posterior over model parameters of interest $p(\boldsymbol{\theta}|\boldsymbol{x},\boldsymbol{\phi})$.
+Thus, it remains to find a tractable expression for eq.\ref{eqn:8} to be minimized by backpropagation given a finite number of simulated samples from the model.
+To this end, we recall that we can relate the pdf of $\boldsymbol{\theta}$ to that of $\boldsymbol{z}$ via the change of variable theorem:
+\begin{align*} 
+p(\boldsymbol{\theta}|\boldsymbol{x},\boldsymbol{\phi}) &= p(\boldsymbol{z})\left|\det\left(\frac{\partial \boldsymbol{z}}{\partial \boldsymbol{\theta}}\right)\right| \numberthis \label{eqn:10} \\ 
+&= p\left(f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi})\right)\left|\det\left(\frac{\partial f}{\partial \boldsymbol{\theta}}\right)\right| \numberthis \label{eqn:11}
+\end{align*}
+where $\partial f/\partial \boldsymbol{\theta} = \boldsymbol{J}_{f}$ is the Jacobian of the learned transformation $f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi})$ with respect to the input.
+Both terms in eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the log determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
+\ref{eqn:2} and \ref{eqn:3}.
+We can now formulate the ML loss as the Monte-Carlo approximation of the negative logarithm of eq.\ref{eqn:8} for a batch of size $m$:
+\begin{align*}
+\mathcal{L}(\boldsymbol{\phi}) &= -\frac{1}{m}\sum_{i=1}^{m}\log\left(p(\boldsymbol{\phi}|\boldsymbol{\theta}^{(i)},\boldsymbol{x}^{(i)})\right) \numberthis \label{eqn:12} \\
+&= -\frac{1}{m}\sum_{i=1}^{m}\log\left(p(\boldsymbol{\theta}^{(i)}|\boldsymbol{x}^{(i)},\boldsymbol{\phi})p(\boldsymbol{\phi})\right) \numberthis \label{eqn:13} \\
+&= -\frac{1}{m}\sum_{i=1}^{m}\left(\frac{\norm{f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi})}_{2}^{2}}{2} + \log\left|\det\left(\boldsymbol{J}_{f}^{(i)}\right)\right|\right) + \tau\norm{\boldsymbol{\phi}}_{2}^{2} \numberthis \label{eqn:14}
+\end{align*}
+where we place a Gaussian prior over the neural network parameters with $\tau\equiv 1/\sigma^{2}$, corresponding to a standard $L2$-regularization.
 
-\subsection*{Submitting Manuscripts}
+Minimizing eq.
+\ref{eqn:14} can be interpreted as searching for the optimal neural network parameters $\boldsymbol{\phi}^{*}$ which maximize the probability of model parameters $\boldsymbol{\theta}$ given data $\boldsymbol{x}$.
+This is exactly the probability we are concerned with in Bayesian inference.
+Note that our formulation maximizes the posterior of model parameters directly, in contrast to variational methods which optimize a lower bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
+Once the backpropagation algorithm has settled to a local minimum of the ML loss, one can easily obtain samples from the approximate posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs},\boldsymbol{\phi}=\boldsymbol{\phi}^{*})$, based on an observed dataset $\boldsymbol{x}_{obs}$, by repeatedly sampling $\boldsymbol{z}^{(l)} \sim \mathcal{N}_{d}(\boldsymbol{0},\boldsymbol{I})$, and then passing $\boldsymbol{z}^{(l)}$ in reverse to the cINN in order to compute $\boldsymbol{\theta}^{(l)} = f^{-1}(\boldsymbol{z}^{(l)};\boldsymbol{x}_{obs},\boldsymbol{\phi}^{*})$ for $l=1,...,L$.
+Figure 1b illustrates the inference phase of the method.
+It is worth noting that sampling a large number of parameter values from the approximate posterior takes a negligible amount of time, since it only requires a single pass through the cINN in reverse.
 
-All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
-If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
+\subsection*{Training the Networks}
+We train all cINNs and summary networks described in this paper jointly via backpropagation.
+For all following experiments, we use the Adam optimizer  with a starter learning rate of $10^-3$ and an exponential decay rate of $.95$.
+We set the weight regularization parameter $\tau$ to a value of $10^{-5}$.
+Note, that we did not perform an extensive search for optimal values of the cINN parameter but stick to a cINN with 10 ACBs for each example in this paper.
+Concerning the data generation step, we use two training approaches.
+The first follows the classical approximate Bayesian computation approach to create a large “reference table” or grid of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$.
+The reference table is then used as training data for the neural network and training continues for a pre-specified number of epochs through the entire reference table.
+A separate validation dataset is eventually used to assess the performance of the network.
+This training approach separates the simulation from the training phase but can incur large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
+The second approach follows a different strategy, which is used mainly in the field of active learning.
+Correspondingly, a dataset, or a batch of datasets, is created on the fly and then passed through the neural network.
+This training regime has the advantage that the network never “experiences” the same input data twice.
+Moreover, training can continue as long as the network keeps improving (i.e., the loss keeps decreasing), since overfitting in the classical sense is nearly impossible.
+However, if the simulations are computationally expensive and researchers need to experiment with multiple models, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
+TODO - Describe for which!
 
-\subsection*{Format}
+\subsection*{Putting It All Together}
 
-Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References.
-Other orders and headings are permitted.
+On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $\pi(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\theta}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
+In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the life sciences, since they tend to be unimodal and relatively simple (in contrast to the distribution of natural images or words in a spoken language).
+From these three ingredients, a universal and reusable sampler can be designed for likelihood-free Bayesian estimation of both tractable and intractable mathematical models.
+\textbf{Algorithm} \ref{Algorithm:algo} describes the essential steps of the method using an arbitrary summary network and employing the active learning training regime.
 
-\subsection*{Manuscript Length}
+\begin{algorithm}
+\caption{Bayesian likelihood-free inference with invertible neural networks}\label{Algorithm:algo}
+\begin{algorithmic}[1]
+\State \emph{Training (via active learning):}
+\Repeat
+\State {Sample a batch of $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ from prior $\pi(\boldsymbol{\theta})$}
+\State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = q(\boldsymbol{\theta}^{(i)})$ for $i=1,...,m$}
+\State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$}
+\State {Pass $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ and $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ through cINN $f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi}_{inv})$ to obtain $\{\boldsymbol{z}^{(i)}\}_{i=1}^{m}$}
+\State {Compute ML loss $\mathcal{L}(\boldsymbol{\phi})$ according to eq.\ref{eqn:14}}
+\State {Update neural network parameters $\boldsymbol{\phi}$ via backpropagation}
+\Until {convergence to $\boldsymbol{\phi}^{*}$}
+\State \emph{Inference (given observed or test data $\boldsymbol{x}_{obs}$):}
+\State {Summarize the observed data by computing $\tilde{\boldsymbol{x}}_{obs} = h(\boldsymbol{x}_{obs},\boldsymbol{\phi}_{sum}^{*})$}
+\For{$l = 1,...,L$}
+\State {Sample $\boldsymbol{z}^{(l)} \sim \mathcal{N}_{d}(\boldsymbol{0},\boldsymbol{I})$}
+\State {Compute $\boldsymbol{\theta}^{(l)} = f^{-1}(\boldsymbol{z}^{(l)};\tilde{\boldsymbol{x}}_{obs},\boldsymbol{\phi}_{inv})$}
+\EndFor
+\State {Use $\{\boldsymbol{\theta}^{(l)}\}_{l=1}^{L}$ to approximate the posterior $p(\boldsymbol{\theta}|\boldsymbol{x}_{obs})$}
+\end{algorithmic}
+\end{algorithm}
+
+The backpropagation algorithm works by computing the gradients of the loss function w.r.t.
+the parameters of the neural networks and then adjusting the parameters, so as to drive the loss function to a local minimum.
+We experienced no instability or convergence issues during training with the ML loss.
+Note, that steps $12-15$ of \textbf{Algorithm} \ref{Algorithm:algo} can be executed in parallel with GPU support.
+
+In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of various intractable models from TODO -which?.
+Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.All examples are implemented in Python using the \textit{TensorFlow} library for automatic differentiation with GPU support.
+
+\section*{Results}
+
+\subsection*{Toy Example}
+
+\subsection*{Example 1}
+
+\subsection*{Example 2}
+
+\subsection*{Example 3}
+
+\subsection*{Example 4}
 
@@
 
-\subsection*{Data Archival}
-
-PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.
-
 \subsection*{Language-Editing Services}
@@
 PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[scale=0.1]{figure1.png}
-\caption{Placeholder image of a frog with a long example caption to show justification setting.}
-\label{fig:frog}
-\end{figure}
-
-\subsection*{Digital Figures}
-
-Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size.
-Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-
-Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
-
-\subsection*{Single column equations}
-
-Authors may use 1- or 2-column equations in their article, according to their preference.
-
-To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
 
@@
 
-\subsubsection*{SI Datasets}
+\section*{Discussion}
 
-Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.
+In the current work, we proposed and explored a novel end-to-end likelihood-free method which uses invertible neural networks to perform approximate Bayesian inference on any mathematical process model.
+We demonstrated the utility of the method by applying it to models from various scientific domains exhibiting various data formats and data-generating mechanisms.
+Further, we explored two possible training approaches suitable for different simulation scenarios, namely a reference-table approach and an active learning approach.
+Both training approaches lead to excellent recovery of the true parameters throughout the examples considered in the current work.
+
+Our method combines the universal approximation power of deep learning methods with the important uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
+Besides being capable of performing rapid Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable “parameter estimation machines” for various research domains.
+Moreover, the method is not confined solely to inference on intractable models, but can also prove as a viable alternative in modeling contexts where standard Bayesian or frequentist inference methods are available, but inference is nevertheless prohibitively slow.
+
+Inspired by previous machine learning approaches to likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, our method shares many of the advantages of these methods and further overcomes some important limitations.
+
+First, the introduction of separate summary and inference neural network modules renders the invertible inference module independent of the shape or the size of the observed data.
+This is achieved by learning a fixed-size vector representation of the data through the summary module in an automatic, data-driven manner.
+This is particularly useful in settings where the most informative summary statistics are not known and thus potential information is lost through the choice of suboptimal summary function.
+However, if informative or even \textit{sufficient} statistics are available in a given domain, one might dispose with the summary module altogether and feed the summary statistics directly to the cINN.
+
+Second, we showed that the ML loss optimizes directly the posterior over model parameters of interest, which is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
+Thus, inference is exact when the ML is minimized \cite{kingma2018glow, dinh2016density}.
+This \textit{optimal performance} claim is confirmed by our toy Bayesian regression example, in which we observe negligibly small deviations of the approximate from the true posterior.
+Further, researchers are often interested in some summary of the posterior, for instance, the posterior mean or the posterior variance \cite{radev2019towards, raynal2018abc}.
+We demonstrated that our method exhibits excellent recovery of the posterior means throughout the examples.
+We also showed that the recovery becomes better with increasing number of observed data points, while the variance of the estimates becomes larger.
+This is an important and highly desirable property of any parameter estimation method, as it mirrors the increase oin information following increasing number of data points.
+
+Third, the largest computational cost of our method is paid during training.
+Once trained, the networks can be used and reused to perform inference on large numbers of datasets within seconds and across a given research domain.
+Indeed, there are many instances of research domains where a single model is extensively explored and independently fitted by multiple researches (CITATION DIFFUSION MODELLING, POPULATION DYNAMICS, NEUROSCIENCE) to test scientific theories about the modelled process.
+These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for inference on different datasets.
+In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
+Note, however, that our method does not need to store training data on disk, since the "knowledge" about the relationship between data and parameters is compressed into the networks' weights.
+Moreover, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern graphical processing units (GPUs).
+
+These advantages notwithstanding, some limitations of the method deserve a mention.
+Even though high-level deep learning libraries, such as \textit{TensorFlow} or \textit{Torch}, allow for rapid and relatively straightforward development of various neural network architectures, the implementational burden associated with the current method is still reasonably high.
+In order to ease the understanding and independent application of the method, we provide fully functioning code to reproduce and study all of the examples tackled in this paper (\href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}).
+Moreover, we are currently developing a general user-friendly software, which should abstract away most of the methodological complexities from the user.
+Another potential shortcoming of the method is the seemingly overwhelming number of hyperparameters that might require fine-tuning by the user for optimal performance on a given task.
+However, we observe that many of the default values of hyperparameters are sufficient to achieve good performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training.
+We expect that a single architecture should be able to perform well on almost all models from a given domain (i.e., a single architecture for decision-making models).
+Future research should investigate the question of generality by applying the method to challenging parameter estimation tasks across different research domains.
+
+\section*{Conclusion}
+
+As formal theories in various scientific disciplines (especially in the younger sciences, such as, neuroscience, cognitive science, computational biology, etc.) become increasingly complex, the need for powerful and universally applicable likelihood-free estimation methods becomes increasingly pressing.
+In the present work, we addressed this need by introducing a method potentially applicable to \textit{any} modeling scenario in \textit{any} research domain where simulations from the process model can be obtained.
+We hope that the new method will enable researchers from a variety of fields to accelerate model-based inference and will further prove its utility beyond the examples considered in this paper.
 

### 2019-06-21 viz.py

@@ -53,7 +53,7 @@ def plot_true_est_scatter(model, X_test, theta_test, n_samples,
         
         # Compute R2
         r2 = r2_score(theta_test[:, j], theta_approx_means[:, j])
-        axarr[j].text(0.1, 0.85, '$R^2$={:.3f}'.format(r2),
+        axarr[j].text(0.1, 0.8, '$R^2$={:.3f}'.format(r2),
                      horizontalalignment='left',
                      verticalalignment='center',
                      transform=axarr[j].transAxes, 

### 2019-06-23 latex/paper_main.tex
Compute all metrices

@@
     \caption{A graphic illustration of the two phases of the method.
-\textbf{(a)} During the training phase, the networks are trained on simulated data from the model; \textbf{(b)} during the inference phase, the posterior is approximated from real data.}
+\textbf{(a)} During the training phase, the networks are trained on simulated data from the model; \textbf{(b)} during the inference phase, the true posterior of the model parameters is approximated from real data.}
     \label{fig:first}

### 2019-06-23 inn_utils.py

@@ -1,6 +1,8 @@
 import tensorflow as tf
 import numpy as np
 from losses import maximum_likelihood_loss
+from sklearn.metrics import r2_score
+
 
 def apply_gradients(optimizer, gradients, variables, global_step=None):
     """
@@ -150,3 +152,61 @@ def train_loop_dataset(model, optimizer, dataset, batch_size, p_bar=None, clip_v
         p_bar.update(1)
 
     return losses
+
+
+def compute_metrics(model, param_names, simulate_fun, n_test, 
+                    n_samples_posterior, transform=None, n_min=100, n_max=1000):
+    """
+    Plots a given metric for different numbers of datapoints.
+    ---------
+
+    Arguments:
+    model           : tf.keras.Model -- the invertible chaoin with an optional summary net
+                                        both models are jointly trained
+    param_names     : list of strings -- the names of the parameters
+    simulate_fun    : callable -- the simulate function
+    n_test          : number of test datasets
+    n_samples_posterior : number of samples from the approximate posterior
+    transform       : callable ot None -- a function to transform X and theta, if given
+    n_min           : int -- the minimum number of data points for each dataset
+    n_max           : int -- the maximum number of data points for each dataset
+    ----------
+
+    Returns:
+    ns      : np.array -- the array with time points
+    metrics : dict -- a dictionary with the metrics
+    """
+    
+    #Plot NRMSE for all t
+    ns = np.arange(n_min, n_max+1)
+    metrics = {
+        'nrmse': {k: [] for k in param_names},
+        'r2': {k: [] for k in param_names},
+        'var': {k: [] for k in param_names}
+    }
+    # For each possible number of data points
+    for n_points in ns:
+        # Generate data
+        X_test, theta_test = simulate_fun(n_test, n_trials=n_points)
+        if transform is not None:
+            X_test, theta_test = transform(X_test, theta_test)
+        theta_test = theta_test.numpy()
+
+        # Sample from approx posterior and compute posterior means
+        theta_approx_means = model.sample(X_test, n_samples_posterior, to_numpy=True).mean(axis=0)
+        
+        # --- Plot true vs estimated posterior means on a single row --- #
+        for j, name in enumerate(param_names):
+
+            # Compute NRMSE
+            rmse = np.sqrt(np.mean( (theta_approx_means[:, j] - theta_test[:, j])**2 ))
+            nrmse = rmse / (theta_test[:, j].max() - theta_test[:, j].min())
[…]

### 2019-06-24 latex/paper_main.tex
new latex files

@@
 \dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about probabilistic phenomena in a probabilistic world.
-In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a generative model $q$ mimicking the process by which manifest quantities $\boldsymbol{x}$ arise from latent parameters:
+In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a forward model $q$ mimicking the process by which manifest data $\boldsymbol{x}$ arise from latent parameters:
 \begin{align*}
@@
 While $q$ can represent an arbitrarily complex process by an arbitrarily complicated expression, its functional form is usually guided by a well-founded theoretical framework.
-For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step mechanism dictating the rate of gene expression in certain cells.
+For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm dictating the rate of gene expression in certain cells.
 Thus, it is only through theoretical embedding that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
@@
 Once a mathematical model has been formulated, the next step consists of fitting the model to experimental or observational data and recovering the parameters of interest.
-However, estimating the parameters of a mathematical model can quickly become one of the most tenacious challenges in applications to real-world problems.
+However, estimating the parameters of a mathematical model from data can quickly become one of the most tenacious challenges in applications to real-world problems.
 It is also one of the most important ones to be tackled, since without reliable parameter estimation methods, it is impossible to test the utility of a model, regardless of its sophistication or theoretical appeal.
@@
 A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate, csillery2010approximate}.
-ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $\pi(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters.
+ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters.
 Thus, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
@@
 The most common approach has been to cast the problem of parameter estimation as a supervised learning task.
-In this setting, a large dataset of the form $\boldsymbol{D} = \{h(\boldsymbol{x}^{(i)}), \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ is created by repeatedly sampling from $\pi(\boldsymbol{\theta})$ and simulating an artificial dataset $\boldsymbol{x}$ by running $q(\boldsymbol{\theta})$ with the sampled parameters.
+In this setting, a large dataset of the form $\boldsymbol{D} = \{h(\boldsymbol{x}^{(i)}), \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ is created by repeatedly sampling from $p(\boldsymbol{\theta})$ and simulating an artificial dataset $\boldsymbol{x}$ by running $q(\boldsymbol{\theta})$ with the sampled parameters.
 Usually, the dimensionality of the simulated data is reduced by computing summary statistics  with a fixed summary function $h(\boldsymbol{x})$.
@@
 We can leverage the fact that most mathematical models are generative in nature and as such can be used to perform multiple simulations of the process of interest.
-By specifying a prior distribution over the model parameters $\pi(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters.
+By specifying a prior distribution over the model parameters $p(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters.
 Thus, at inference time, one can condition the model on observed data $\boldsymbol{x}_{obs}$ and obtain samples $\boldsymbol{\theta}_{1}, \boldsymbol{\theta}_{2},...,\boldsymbol{\theta}_{L}$  approximating the posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$ in the manner described above.
@@
 Each ACB consists of four separate fully connected neural networks denoted as $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$.
-An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ is also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ "for free".
+An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ it also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ "for free".
 Denoting the input vector of $f_{\boldsymbol{\phi}_{inv}}$ as $\boldsymbol{u}$ and the output vector as $\boldsymbol{v}$, it follows that $f(\boldsymbol{u}; \boldsymbol{\phi}_{inv}) = \boldsymbol{v}$ and $f^{-1}(\boldsymbol{v}; \boldsymbol{\phi}_{inv}) = \boldsymbol{u}$.
@@
 We set the weight regularization parameter $\tau$ to a value of $10^{-5}$.
-Note, that we did not perform an extensive search for optimal values of the cINN parameter but stick to a cINN with 10 ACBs for each example in this paper.
+We perform $100000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
+Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter but use a cINN with 10 ACBs for each example in this paper (see \textbf{SI} for details of the cINN).
 Concerning the data generation step, we use two training approaches.
@@
 
-On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $\pi(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\theta}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
+On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\theta}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
 In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the life sciences, since they tend to be unimodal and relatively simple (in contrast to the distribution of natural images or words in a spoken language).
@@
 \Repeat
-\State {Sample a batch of $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ from prior $\pi(\boldsymbol{\theta})$}
+\State {Sample a batch of $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ from prior $p(\boldsymbol{\theta})$}
 \State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = q(\boldsymbol{\theta}^{(i)})$ for $i=1,...,m$}
@@
 \EndFor
+\State {\textbf{end for}}
 \State {Use $\{\boldsymbol{\theta}^{(l)}\}_{l=1}^{L}$ to approximate the posterior $p(\boldsymbol{\theta}|\boldsymbol{x}_{obs})$}
@@
 
-\subsection*{Toy Example}
+\subsection*{Toy Example – Bayesian Regression}
 
-\subsection*{Example 1}
+As a proof-of-concept, we demonstrate the utility of our method in recovering the true analytic posteriors of the regression coefficients of a conjugate Bayesian regression model.
+To set the setting, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$ We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, i.e.,: $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
+We place a $d$-dimensional diagonal Gaussian prior on the regression coefficients centered at $0$: $\boldsymbol{\theta} \sim \mathcal{N}_{d}(\boldsymbol{0},b^{-1}\boldsymbol{I})$ where $b$ is the precision of the prior ($b \equiv 1/\sigma_{\boldsymbol{\theta}}^{2}$).
+Thus, the likelihood $p(\boldsymbol{D}|\boldsymbol{\theta})$ admits the following proportionality:
+\begin{align*}
+p(\boldsymbol{D}|\boldsymbol{\theta}) \propto \exp\left(-\frac{a}{2}\left(y-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(y - \boldsymbol{X}\boldsymbol{\theta}\right)\right) \numberthis 
+\end{align*}
+where $\boldsymbol{X}$ denotes the design matrix (with all $\boldsymbol{x}^{(i)}$ stacked row-wise).
+Since the prior of $\boldsymbol{\theta}$ is conjugate to the likelihood (both are Gaussian distributions), the posterior of $\boldsymbol{\theta}$ is also Gaussian, given by:
+\begin{align*}
+p(\boldsymbol{\theta}|\boldsymbol{D}) \propto \exp\left(-\frac{a}{2}\left(y-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(y - \boldsymbol{X}\boldsymbol{\theta}\right)-\frac{b}{2}\boldsymbol{\theta}^{T}\boldsymbol{\theta}\right) \numberthis
+\end{align*}
+Therefore, the posterior has the form $p(\boldsymbol{\theta}|\boldsymbol{D}) = \mathcal{N}_{d}(\boldsymbol{\theta}|\boldsymbol{\mu},\boldsymbol{\Lambda}^{-1})$ where $\boldsymbol{\Lambda}$ denotes the posterior precision matrix (inverse covariance matrix), $\boldsymbol{\mu}$ the posterior mean vector, which are computed as follows:
+\begin{align*}
+\boldsymbol{\Lambda} &= a\boldsymbol{X}^{T}\boldsymbol{X} + b\boldsymbol{I} \numberthis \\
+\boldsymbol{\mu} &= a\boldsymbol{\Lambda}^{-1}\boldsymbol{X}^{T}\boldsymbol{y} \numberthis
+\end{align*} 
+Thus, for known $a$ and $b$, the posterior can be easily computed.
+Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is good for proving the utility of our method.
 
-\subsection*{Example 2}
+For the following application of our method, we set $d=4$, and $a=b=1$.
+The design matrices for each iteration contain a variable number $n$ of \textit{i.i.d.} data points drawn from a unit Gaussian $\mathcal{N}_{4}(\boldsymbol{0},\boldsymbol{I})$.
+The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(50, 500)$ at each training iteration (Lines $2$-$9$ of \textbf{Algorithm} \ref{Algorithm:algo}).
 
-\subsection*{Example 3}
+The results on the toy example are depicted in Figure.XXX.
+The approximate posterior means show negligible deviations from the analytic posterior means as quantified by very small $NRMSE$s (as small as XXX) and very high $R^{2}$ (as high as $1.0$) over all numbers of data points used for training.
+This suggests near-perfect reconstruction of the true posterior means.
 
-\subsection*{Example 4}
+\subsection*{Example 1 - The Ricker Model}
+Discrete population dynamics models describe how the number of individuals in a population changes over discrete units of time \cite{wood2010statistical}.
+In particular, the Ricker model describes the number of individuals $x_{t+1}$ in generation $t+1$ as a function of the number of individuals in the previous generation $t$ by the following non-linear equation:
+\begin{align*}
+N_{t+1} &= rN_{t}\exp(-N_{t} + \epsilon_{t}) \numberthis \\
+x_{t} &\sim Pois(\rho N_{t}) \numberthis
+\end{align*}
+for $t = 1,...,T$ where $N_{t}$ is the expected number of individuals at time $t$, $r$ is the growth rate, $\rho$ is a scaling parameter and $\epsilon_{t} \sim \mathcal{N}(0, \sigma^{2})$ is random Gaussian noise \cite{mestdagh2018prepaid}.
+The Ricker model has no known likelihood function and is thus a suitable candidate for likelihood-free inference.
+The parameter estimation task is thus to recover $\boldsymbol{\theta}=(\rho,r,\sigma)$ from the observed one-dimensional time-series data $\boldsymbol{x}=(x_{1},x_{2},...,x_{T})$ where each $x_{t} \in \mathbb{N}$.
 
-PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
-The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.
-When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.
+During training of the networks, we simulate time-series from the Ricker model with varying lengths.
+The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
 
-\subsection*{Language-Editing Services}
-Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml).
-PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.
+The results on the Ricker model are depicted in Figure XXX... TODO
+
+\subsection*{Example 2 - The Lévy-Flight Model}
+Evidence accumulator models (EAMs) describe (perceptual) decision making by a set of neurocognitively motivated parameters CITE.
+EAMs are most often applied to choice reaction times (RT) data to obtain an estimate of the neurocognitive processes governing observed RT distributions in human and animal participants.
+Most EAM flavors share four underlying assumptions: \textit{i)} information about a stimulus (response option) is accumulated continuously through time; \textit{ii)} stochasticity in the form of noisy accumulation ensures variability; \textit{iii)} empirical response times can be decomposed into a decision time component and a non-decision time component accounting for pre-decisional perceptual (encoding time) and post-decisional motor processes (response execution); and \textit{iv)} a decision is met when the activation of an accumulator exceeds a threshold.
+In its most general formulation, the forward model of EAMs takes the form of a stochastic differential equation given by:
+\begin{align*} 
+dx = vdt + cd\xi  \numberthis \label{eqn:stillno}
+\end{align*}
+where $dx$ denotes a change in activation of an accumulator, $v$ denotes the average speed of information processing (often termed the drift rate), and $d\xi$ represents a stochastic additive component, usually modeled as following a Gaussian distribution centered around $0$: $d\xi \sim \mathcal{N}(0, c^{2})$.
+
+EAMs are particularly amenable for likelihood-free inference, since most members of the family turn out to be intractable.
+This intractability has precluded many interesting applications and empirically driven model refinements.
+Here, we apply our method to estimate the parameters of the recently proposed Lévy-Flight Model (LFM, cite).
+The LFM assumes an \textit{alpha-stable} noise distribution of the evidence accumulation process in order to model "jumps" in the decision process.
+The LFM does not have a known likelihood and a fully Bayesian treatment of the model is still missing from the literature.
+The forward equation of the LFM is given by:
+\begin{align*}
+dx &= vdt + \xi dt^{1/\alpha} \numberthis \label{eqn:stillno2} \\
+\xi &\sim AlphaStable(\alpha,0,1,0) \numberthis \label{eqn:stillno3}
+\end{align*}
+The LFM has three additional parameters: a threshold $a$ determining the amount of evidence needed for the termination of a decision process; a relative starting point, $zr$, determining the amount of starting evidence available to the accumulator before the actual decision alternatives are presented; and an additive non-decision time $t_{0}$.
+
+During training of the networks, we simulate RT data from two experimental conditions with two different drift rates (see \textbf{SI} for details of the simulation).
+The parameter estimation task is thus to recover the parameters $\boldsymbol{\theta} = (v_{0}, v_{1}, a, t_{0}, zr, \alpha)$ from two-dimensional \textit{i.i.d.} RT data $\boldsymbol{X} = (\boldsymbol{x}_{1},\boldsymbol{x}_{2},...,\boldsymbol{x}_{n})$ containing variable number or RT trials.
+The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(100, 1000)$ at each training iteration.
+
+The results on the LFM model are depicted in Figure XXX.
+First, we note that our method achieves state-of-the art parameter recovery, with all $R^{2}$ values above $.97$ and $NRMSE$ below...
+Some of the parameters are almost perfectly reconstructed.
+Describe other stuff...
+
+\subsection*{Example 3 – The Stochastic SIR Model}
+Compartmental models in epidemiology are used to describe the stochastic dynamics of infectious diseases as they spread over a population of individuals \cite{hethcote2000mathematics}.
+The parameters of compartmental models encode important characteristics of diseases, such as the rates of infection or recovery from the disease.
+The stochastic SIR model describes the transition dynamics of N individuals between three discrete states: susceptible (S), infected (I), and recovered (R).
+The transition dynamics are given by the following equations:
+\begin{align*}
+\triangle S &= -\triangle N_{SI} \numberthis \\
+\triangle I &= \triangle N_{SI} - \triangle N_{IR} \numberthis \\
+\triangle R &= \triangle N_{IR} \numberthis \\
+\triangle N_{SI} &\sim Binomial(S, 1 - \exp\left(-\beta\frac{I}{N}\triangle t \right)) \numberthis \\
+\triangle N_{IR} &\sim Binomial(I, 1 - \exp\left(-\gamma\triangle t \right)) \numberthis 
+\end{align*}
+where $S + I + R = N$ give the number of susceptible, infected, and recovered individuals, respectively.
+The parameter $\beta$ controls the transition rate from being susceptible to infected, and $\gamma$ controls the transition rate from being infected to recovered.
+The number of individuals moving from $S$ to $I$, given by $\triangle N_{SI}$, and the number of people moving from $I$ to $R$, given by $\triangle N_{IR}$, over a time interval $\triangle t$ are modeled as binomial random variables.
+The above listed stochastic system has no known analytic solution and thus requires numerical simulation methods for finding optimal parameter values.
+Cast as a parameter estimation task, the challenge is to recover $\boldsymbol{\theta} = \{\beta,\gamma\}$ from three dimensional time-series data $\boldsymbol{X} = (\boldsymbol{x}_{1},\boldsymbol{x}_{2},...,\boldsymbol{x}_{T})$ where each $\boldsymbol{x}_{t} \in \mathbb{N}^{3}$ is a triple containing the number of susceptible ($S$), number of infected ($I$), and recovered ($R$) individuals at time $t$.
+
+During training of the networks, we simulate time-series from the stochastic SIR model with varying lengths.
+The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 200)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+This range is chosen since we observed that the system reaches an equilibrium in almost all parameter settings after $T > 200$.
+
+The results on the SIR model are depicted in Figure XXX.
+
+\subsection*{Example 4 - Single-Cell Sequence RNA}
+dsdsdsds
 

### 2019-06-24 process_models.py

@@ -34,7 +34,7 @@ def simulate_ricker_batch(X, params, n_batch, t_obs):
         X[i, :] = simulate_ricker_single(t_obs, params[i, 0], params[i, 1], params[i, 2])
     
 
-def simulate_ricker(batch_size=64, t_obs_min=100, t_obs_max=500, low_r=1, high_r=90, 
+def simulate_ricker(batch_size=64, t_obs=None, t_obs_min=100, t_obs_max=500, low_r=1, high_r=90, 
                     low_sigma=0.05, high_sigma=0.7, low_phi=0, high_phi=20, to_tensor=True):
     """
     Simulates and returns a batch of 1D timeseries obtained under the Ricker model.
@@ -58,8 +58,9 @@ def simulate_ricker(batch_size=64, t_obs_min=100, t_obs_max=500, low_r=1, high_r
               a batch or time series generated under a batch of Ricker parameters
     """
     
-    # Sample t_obs
-    t_obs = np.random.randint(low=t_obs_min, high=t_obs_max+1)
+    # Sample t_obs, if None given
+    if t_obs is None:
+        t_obs = np.random.randint(low=t_obs_min, high=t_obs_max+1)
 
     # Prepare placeholders
     theta = np.random.uniform(low=[low_r, low_sigma, low_phi], 
@@ -104,7 +105,7 @@ def simulate_sir_single(beta, gamma, t_max=500, N=1000):
     return np.array([S, I, R]).T
 
 @jit
-def simulate_sir(batch_size, low_beta=0.05, high_beta=2., low_gamma=0.01, t_max=500, N=1000, to_tensor=True):
+def simulate_sir(batch_size, low_beta=0.01, high_beta=1., low_gamma=0.01, t_max=500, N=1000, to_tensor=True):
     """
     Simulates and returns a batch of timeseries obtained under the SIR model.
     ----------

### 2019-06-24 simulate_diffusion.py

@@ -48,14 +48,15 @@ def simulate_batch_diffusion_p(x, params):
                                          params[i, 7], params[i, 8], 0.001, 5000)
 
 def simulate_diffusion(batch_size, pbounds, n_trials=None, n_cond=2, 
-                       to_tensor=True, cond_coding=False):
+                       to_tensor=True, cond_coding=False, n_trials_min=100, n_trials_max=1000):
     """Simulates batch_size datasets from the full Ratcliff diffusion model."""
     
     # Get number of parameters
     n_params = len(pbounds)
 
+    # Sample number of trials, if None given
     if n_trials is None:
-        n_trials = np.random.randint(100, 1001)
+        n_trials = np.random.randint(n_trials_min, n_trials_max+1)
 
     # Extract parameter bounds
     lower_bounds = [pbounds['v1'][0], pbounds['v2'][0], pbounds['sv'][0], pbounds['zr'][0],
@@ -71,6 +72,7 @@ def simulate_diffusion(batch_size, pbounds, n_trials=None, n_cond=2,
     X_batch = np.zeros((batch_size, n_trials, n_cond), dtype=np.float32)
     simulate_batch_diffusion_p(X_batch, theta_batch)
 
+    # Return in specified format (condition coding or just stack, tf.Tensor or np.array)
     if cond_coding:
         X_batch = np.stack(([np.c_[X_batch[:, :, 0], X_batch[:, :, 1]], 
                              np.c_[np.zeros((batch_size, n_trials)), np.ones((batch_size, n_trials))]]), axis=-1)

### 2019-06-24 latex/paper_main.tex
new latex files

@@
 Thus, it is only through theoretical embedding that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
-Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical}, epidemiology \cite{hethcote2000mathematics}, just to name a few.
+Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, just to name a few.
 
@@
 In the \textbf{Methods} section, we show that our method maximizes the posterior over model parameters directly when cast in the context of likelihood-free inference.
-Furthermore, flow-based methods are capable of approximating very high-dimensional distributions (e.g., images).
+Furthermore, flow-based methods are capable of approximating very high-dimensional distributions (e.g., the pixels of an image).
 Once trained with a sufficient amount of simulated data, our invertible network can perform full Bayesian inference on large amounts of real data from a given domain in a single pass.
@@
 \subsection*{Example 2 - The Lévy-Flight Model}
-Evidence accumulator models (EAMs) describe (perceptual) decision making by a set of neurocognitively motivated parameters CITE.
+Evidence accumulator models (EAMs) describe (perceptual) decision making by a set of neurocognitively motivated parameters \cite{ratcliff2008diffusion}.
 EAMs are most often applied to choice reaction times (RT) data to obtain an estimate of the neurocognitive processes governing observed RT distributions in human and animal participants.
-Most EAM flavors share four underlying assumptions: \textit{i)} information about a stimulus (response option) is accumulated continuously through time; \textit{ii)} stochasticity in the form of noisy accumulation ensures variability; \textit{iii)} empirical response times can be decomposed into a decision time component and a non-decision time component accounting for pre-decisional perceptual (encoding time) and post-decisional motor processes (response execution); and \textit{iv)} a decision is met when the activation of an accumulator exceeds a threshold.
-In its most general formulation, the forward model of EAMs takes the form of a stochastic differential equation given by:
+Most EAM variants share four underlying assumptions: \textit{i)} information about a stimulus (response option) is accumulated continuously through time; \textit{ii)} stochasticity in the form of noisy accumulation ensures variability; \textit{iii)} empirical response times can be decomposed into a decision time component and a non-decision time component accounting for pre-decisional perceptual (encoding time) and post-decisional motor processes (response execution); and \textit{iv)} a decision is met when the activation of an accumulator exceeds a threshold.
+In its most general formulation, the forward model of EAMs takes the form of a stochastic differential equation given by \cite{usher2001time}:
 \begin{align*} 
@@
 
-EAMs are particularly amenable for likelihood-free inference, since most members of the family turn out to be intractable.
+EAMs are particularly amenable for likelihood-free inference, since most members of the family turn out to be intractable \cite{miletic2017parameter}.
 This intractability has precluded many interesting applications and empirically driven model refinements.
-Here, we apply our method to estimate the parameters of the recently proposed Lévy-Flight Model (LFM, cite).
+Here, we apply our method to estimate the parameters of the recently proposed Lévy-Flight Model (LFM, \cite{voss2019sequential}).
 The LFM assumes an \textit{alpha-stable} noise distribution of the evidence accumulation process in order to model "jumps" in the decision process.
-The LFM does not have a known likelihood and a fully Bayesian treatment of the model is still missing from the literature.
+However, the inclusion of \textit{alpha-stable} noise leads to a model with intractable likelihood; further, to our knowledge, a fully Bayesian treatment of the model is still missing from the literature.
 The forward equation of the LFM is given by:
@@
 \subsection*{Example 3 – The Stochastic SIR Model}
-Compartmental models in epidemiology are used to describe the stochastic dynamics of infectious diseases as they spread over a population of individuals \cite{hethcote2000mathematics}.
+Compartmental models in epidemiology are used to describe the stochastic dynamics of infectious diseases as they spread over a population of individuals \cite{sahneh2017gemfsim,keeling2011modeling, hethcote2000mathematics}.
 The parameters of compartmental models encode important characteristics of diseases, such as the rates of infection or recovery from the disease.
-The stochastic SIR model describes the transition dynamics of N individuals between three discrete states: susceptible (S), infected (I), and recovered (R).
+The stochastic SIR model describes the transition dynamics of $N$ individuals between three discrete states: susceptible ($S$), infected ($I$), and recovered ($R$).
 The transition dynamics are given by the following equations:

### 2019-06-25 latex/paper_main.tex
Prepare seqRNA notebook

@@
 % {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
 
@@
 \usepackage[noend]{algpseudocode}
-\usepackage{subfig}
+%\usepackage{subfig}
+\usepackage{subcaption} 
 \graphicspath{{figures/}}
@@
 It treats parameter inference as a task of inverting a generative model and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
-The method integrates two separate deep neural networks modules (detailed in the \textbf{Methods} section) trained jointly on simulated data: a \textit{summary network} and an \textit{invertible network}.
+The method integrates two separate deep neural networks modules (detailed in the \textbf{Methods} section; see \autoref{fig:Fig.1}) trained jointly on simulated data: a \textit{summary network} and an \textit{invertible network}.
 
@@
 Once trained with a sufficient amount of simulated data, our invertible network can perform full Bayesian inference on large amounts of real data from a given domain in a single pass.
-
 \begin{figure}
-    \centering
-    \subfloat[Training phase]{{\includegraphics[width=.99\linewidth]{training.png}}}
-    \newline
-    \subfloat[Inference phase]{{\includegraphics[width=.99\linewidth]{inference.png}}}
-    \caption{A graphic illustration of the two phases of the method.
-\textbf{(a)} During the training phase, the networks are trained on simulated data from the model; \textbf{(b)} during the inference phase, the true posterior of the model parameters is approximated from real data.}
-    \label{fig:first}
+  \begin{subfigure}[b]{1.0\textwidth}
+    \includegraphics[width=\textwidth]{training.png}
+    \caption{Training phase}
+    \label{fig:Fig.1a}
+  \end{subfigure}
+  %
+  \begin{subfigure}[b]{1.0\textwidth}
+    \includegraphics[width=\textwidth]{inference.png}
+    \caption{Inference phase}
+    \label{fig:Fig.1b}
+  \end{subfigure}
+      \caption{Graphical illustration of the proposed method.
+\textbf{(a)} During the training phase, the networks are trained on simulated data from the model and updated after each batch of simulations; \textbf{(b)} During the inference phase, the true posterior of the model parameters is approximated from real data using the trained networks.
+Thus, knowledge about the relationships between parameters and data (the mathematical model) is compactly encoded within the weights of the two networks.
+} \label{fig:Fig.1}
 \end{figure}
-
-The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated statistical problems in various scientific domains (see Fig.1).
+The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated statistical problems in various scientific domains (\autoref{fig:Fig.1}).
 Moreover, the method addresses many of the limitations of previous likelihood-free methods.
@@
 Most popular deep probabilistic models entail two phases.
-During the \textit{training phase}, a transformation from the simple to the desired target distribution is learned by optimizing a cost function via backpropagation.
-During the \textit{inference phase}, samples from the target distribution are obtained by sampling from the simple distribution and applying the transformation learned during the training phase.
+During the \textit{training phase}, a transformation from the simple to the desired target distribution is learned by optimizing a cost function via backpropagation (see \autoref{fig:Fig.1a}).
+During the \textit{inference phase}, samples from the target distribution are obtained by sampling from the simple distribution and applying the transformation learned during the training phase (see \autoref{fig:Fig.1b}).
 Using this approach, recent applications of deep probabilistic models have achieved unprecedented results on extremely high-dimensional and intractable problems (e.g., complex data distributions such as natural images, music, or text).
@@
 
-The basic building block of a cINN is the affine coupling block (ACB) \cite{ardizzone2018analyzing, kingma2018glow, dinh2016density}.
-Each ACB consists of four separate fully connected neural networks denoted as $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$.
+The basic building block of a cINN is the affine coupling block (ACB, see \autoref{fig:Fig.2a}) \cite{ardizzone2018analyzing, kingma2018glow, dinh2016density}.
+Each cACB consists of four separate fully connected neural networks denoted as $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$.
 An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ it also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ "for free".
@@
 Furthermore, the internal networks $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$ can be represented by arbitrarily complex neural networks, which themselves need not be invertible, since they are only ever evaluated in the forward direction during both the forward and the inverse pass through the ACB.
-To ensure that the model is powerful enough to represent complicated distributions, we chain multiple ACBs, so that the output of each ACB becomes the input of the next.
+To ensure that the model is powerful enough to represent complicated distributions, we chain multiple ACBs, so that the output of each ACB becomes the input of the next (see \autoref{fig:Fig.2b}).
 In this way, the whole chain remains invertible from the first input to the last output and can be viewed as a single function parameterized by trainable parameters $\boldsymbol{\phi}_{inv}$.
@@
 In the next section, we describe the optimization procedure used to match the outputs of $f^{-1}(\boldsymbol{z};\boldsymbol{x},\boldsymbol{\phi}_{inv})$ to the posterior $p(\boldsymbol{\theta}|\boldsymbol{x})$.
+
+% Figures ACB and ACB chain
+\begin{figure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{acb.png}
+    \caption{The cACB}
+    \label{fig:Fig.2a}
+  \end{subfigure}
+  %
+  \begin{subfigure}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{acbchain.png}
+    \caption{A chain of cACBs}
+    \label{fig:Fig.2b}
+  \end{subfigure}
+  \caption{A diagram of the conditional version of the affine coupling block (cACB).
+\textbf{(a)} Each cABC consists of four internal networks performing the invertible operations described in the text; \textbf{(b)} In practice, we chain multiple cACBs to obtain higher representational capacity.} \label{fig:Fig.2}
+\end{figure}
 

### 2019-06-26 latex/paper_main.tex
Add supplementary methods and KL divergence

@@
 \usepackage[noend]{algpseudocode}
-%\usepackage{subfig}
 \usepackage{subcaption} 
+\usepackage{cleveref}
 \graphicspath{{figures/}}
@@
 
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: stefan.radev\@psychologie.uni-heidelberg.de}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: stefan.radev@psychologie.uni-heidelberg.de}
 
@@
   \end{subfigure}
-      \caption{Graphical illustration of the proposed method.
-\textbf{(a)} During the training phase, the networks are trained on simulated data from the model and updated after each batch of simulations; \textbf{(b)} During the inference phase, the true posterior of the model parameters is approximated from real data using the trained networks.
-Thus, knowledge about the relationships between parameters and data (the mathematical model) is compactly encoded within the weights of the two networks.
-} \label{fig:Fig.1}
+      \caption{Graphical illustration of the method.
+\textbf{(a)} During the training phase, the summary and the invertible network are trained on simulated data from the model and updated after each batch of simulations; \textbf{(b)} During the inference phase, the true posterior of the model parameters is approximated from real data using the trained networks.
+Thus, knowledge about the relationship between parameters and data (the mathematical model) is compactly encoded within the weights of the two networks.
+The trained networks can then be shared and used across researchers working on the same model.} \label{fig:Fig.1}
 \end{figure}
@@
   \caption{A diagram of the conditional version of the affine coupling block (cACB).
-\textbf{(a)} Each cABC consists of four internal networks performing the invertible operations described in the text; \textbf{(b)} In practice, we chain multiple cACBs to obtain higher representational capacity.} \label{fig:Fig.2}
+\textbf{(a)} Each cABC consists of four internal networks performing the invertible operations described in the text; \textbf{(b)} In practice, we chain multiple cACBs to obtain higher representational capacity.
+Each cACB layer uses a fixed permutation to ensure that information about each parameter is encoded in each latent dimension of $\boldsymbol{z}$.} \label{fig:Fig.2}
 \end{figure}
@@
 Broadly speaking, the goal of ML estimation is to find a set of parameters which maximize the probability of the data under a parametric model.
-In our case, we are interested in maximizing the expectation over possible neural network parameters with respect to the parameters of the mathematical model:
+In our case, we are interested in maximizing the expectation over all possible neural network parameters with respect to the parameters of the mathematical model:
 \begin{align*} 
@@
 \end{align*}
-Note, that by maximizing eq.\ref{eqn:9} we are maximizing the posterior over model parameters of interest $p(\boldsymbol{\theta}|\boldsymbol{x},\boldsymbol{\phi})$.
-Thus, it remains to find a tractable expression for eq.\ref{eqn:8} to be minimized by backpropagation given a finite number of simulated samples from the model.
+Note, that by maximizing Eq.\ref{eqn:9} we are maximizing the posterior over model parameters of interest $p(\boldsymbol{\theta}|\boldsymbol{x},\boldsymbol{\phi})$.
+Thus, it remains to find a tractable expression for Eq.\ref{eqn:8} to be minimized by backpropagation given a finite number of simulated samples from the model.
 To this end, we recall that we can relate the pdf of $\boldsymbol{\theta}$ to that of $\boldsymbol{z}$ via the change of variable theorem:
@@
 where $\partial f/\partial \boldsymbol{\theta} = \boldsymbol{J}_{f}$ is the Jacobian of the learned transformation $f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi})$ with respect to the input.
-Both terms in eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the log determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
+Both terms in Eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the log determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
 \ref{eqn:2} and \ref{eqn:3}.
-We can now formulate the ML loss as the Monte-Carlo approximation of the negative logarithm of eq.\ref{eqn:8} for a batch of size $m$:
+We can now formulate the ML loss as the Monte-Carlo approximation of the negative logarithm of Eq.\ref{eqn:8} for a batch of size $m$:
 \begin{align*}
@@
 
-Minimizing eq.
+Minimizing Eq.
 \ref{eqn:14} can be interpreted as searching for the optimal neural network parameters $\boldsymbol{\phi}^{*}$ which maximize the probability of model parameters $\boldsymbol{\theta}$ given data $\boldsymbol{x}$.
@@
 From these three ingredients, a universal and reusable sampler can be designed for likelihood-free Bayesian estimation of both tractable and intractable mathematical models.
-\textbf{Algorithm} \ref{Algorithm:algo} describes the essential steps of the method using an arbitrary summary network and employing the active learning training regime.
+\textbf{Algorithm} \ref{alg:1} describes the essential steps of the method using an arbitrary summary network and employing the active learning training regime.
 
 \begin{algorithm}
-\caption{Bayesian likelihood-free inference with invertible neural networks}\label{Algorithm:algo}
+\caption{Bayesian likelihood-free inference with invertible neural networks}\label{alg:1}
 \begin{algorithmic}[1]
@@
 \State {Pass $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ and $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ through cINN $f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi}_{inv})$ to obtain $\{\boldsymbol{z}^{(i)}\}_{i=1}^{m}$}
-\State {Compute ML loss $\mathcal{L}(\boldsymbol{\phi})$ according to eq.\ref{eqn:14}}
+\State {Compute ML loss $\mathcal{L}(\boldsymbol{\phi})$ according to Eq.\ref{eqn:14}}
 \State {Update neural network parameters $\boldsymbol{\phi}$ via backpropagation}
@@
 We experienced no instability or convergence issues during training with the ML loss.
-Note, that steps $12-15$ of \textbf{Algorithm} \ref{Algorithm:algo} can be executed in parallel with GPU support.
+Note, that steps $12-15$ of \textbf{Algorithm} \ref{alg:1} can be executed in parallel with GPU support.
 
-In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of various intractable models from TODO -which?.
+In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of challenging models from population dynamics, cognitive science, epidemiology, and genetics.
 Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.All examples are implemented in Python using the \textit{TensorFlow} library for automatic differentiation with GPU support.
@@
 \section*{Results}
+
+We consider a number of metrics for determining the performance of our method.
+To assess the goodness of parameter recovery, we compute the normalized root mean squared error (NRMSE) and the coefficient of determination ($R^{2}$) metrics of the parameter estimates.
+To assess the recovery of the full posterior, we compute the Kullback-Leibler (KL) divergence \cite{hershey2007approximating} between the true and the approximate distributions for the toy example, and use simulation-based calibration (SBC, \cite{talts2018validating}) for the other examples where the analytic posterior is not available in closed-form.
+Details for computing the $NRMSE$, $R^2$, and SCB can be found in the \textbf{SI}.
 
@@
 \end{align*}
-where $\boldsymbol{X}$ denotes the design matrix (with all $\boldsymbol{x}^{(i)}$ stacked row-wise).
+where $\boldsymbol{X}$ denotes the design matrix containing all $\boldsymbol{x}^{(i)}$ stacked row-wise.
 Since the prior of $\boldsymbol{\theta}$ is conjugate to the likelihood (both are Gaussian distributions), the posterior of $\boldsymbol{\theta}$ is also Gaussian, given by:
@@
 Thus, for known $a$ and $b$, the posterior can be easily computed.
-Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is good for proving the utility of our method.
+Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is good for testing the utility of our method.
 
-For the following application of our method, we set $d=4$, and $a=b=1$.
+For the following application, we set $d=4$, and $a=b=1$.
 The design matrices for each iteration contain a variable number $n$ of \textit{i.i.d.} data points drawn from a unit Gaussian $\mathcal{N}_{4}(\boldsymbol{0},\boldsymbol{I})$.
-The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(50, 500)$ at each training iteration (Lines $2$-$9$ of \textbf{Algorithm} \ref{Algorithm:algo}).
+The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(50, 500)$ at each training iteration (Lines $2$-$9$ of \textbf{Algorithm} \ref{alg:1}).
 
@@
 
-\subsection*{Example 4 - Single-Cell Sequence RNA}
-dsdsdsds
+\subsection*{Example 4 - Single-Cell RNA Sequencing}
+Single-cell RNA sequencing (scRNA-seq) is a method to uncover the dynamics of gene expression within single cells \cite{zappia2017splatter, ozsolak2011rna}.
+Simulation models for scRNA-seq attempt to mimic the process of sequencing real data samples by combining statistical and algorithmic procedures.
+The recently developed \textit{Splat} simulation model \cite{zappia2017splatter} implements a hierarchical model where the mean expression levels for multiple genes are samples from a Gamma distribution, and the number of times a gene has been sequenced in a cell is then sampled from a Poisson distribution.
+The output of the \textit{Splat} simulation is thus a matrix $\boldsymbol{X}$ of \textit{Gene x Cell} counts.
+Figure XXX depicts the \textit{Splat} algorithm for simulating scRNA-seq data.
+
+As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
+We simulate $250000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
+We train the networks for 100 epochs through the entire dataset and evaluate the performance on a separate validation set or matrices.
+
+Figure XXX depicts the results...
 

### 2019-06-26 latex/paper_supplementary.tex
Add supplementary methods and KL divergence

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}
% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.

\graphicspath{{figures/}}
\DeclareMathOperator{\Tr}{Tr}

\title{Learning complex models with invertible neural networks: a likelihood-free Bayesian approach}
\author{Stefan T. Radev, Ulf K. Mertens, Andreas Voss, Lynton Ardizzone, and Ullrich Köthe}
\correspondingauthor{Stefan Radev.\\E-mail: stefan.radev@psychologie.uni-heidelberg.de}

\begin{document}

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section*{Results}

\subsection*{Performance metrics}

In the following, the computation of the performance metrics used throughout the main text is detailed.

\subsubsection*{Normalized Root Mean Squared Error}
The normalized root mean squared error (NRMSE) between a sample of true parameters $\{\theta^{(i)}\}_{i=1}^{n}$ and a sample of estimated parameters $\{\hat{\theta}^{(i)}\}_{i=1}^{n}$ is given by:
\begin{align*}
NRMSE = \sqrt{\sum_{i=1}^{n}\frac{\left(\theta^{(i)}-\hat{\theta}^{(i)}\right)^{2}}{\theta_{max}- \theta_{min}}} \numberthis \label{eqn:1}
\end{align*}
Due to the normalization factor $\theta_{max}-\theta_{min}$, the NRMSE is scale-independent, and thus suitable for comparing the recovery across parameters having different numerical ranges. The NRMSE is zero when the estimates are exactly equal to the true values.
\subsubsection*{Coefficient of Determination }
The coefficient of determination $R^{2}$ gives the proportion of variance in a sample of true parameters $\{\theta^{(i)}\}_{i=1}^{n}$ that is "explained" by a sample of estimated parameters $\{\hat{\theta}^{(i)}\}_{i=1}^{n}$. It is computed as:
\begin{align*}
R^{2} = 1 - \sum_{i=1}^{n}\frac{\left(\theta^{(i)}-\hat{\theta}^{(i)}\right)^{2}}{\left(\theta^{(i)}-\bar{\theta}^{(i)}\right)^{2}} \numberthis \label{eqn:2}
\end{align*}
where $\bar{\theta}$ denotes the mean of the true parameter samples. When $R^{2}$ equals $1$, it means that the estimates are perfect reconstructions of the true parameters.

\subsubsection*{Kullback-Leibler Divergence} The Kullback-Leibler divergence ($D_{KL}$) quantifies the increase in entropy incurred by approximating a target probability distribution $P$ with a distribution $Q$. Its general form for absolutely continuous distributions is given by
\begin{align*}
D_{KL}(P||Q) = \int_{-\infty}^{\infty} p(x)\log\frac{p(x)}{q(x)} dx \numberthis \label{eqn:3}
\end{align*}
where $p$ and $q$ denote the pdfs of $P$ and $Q$. In the case where $P$ and $Q$ are both multivariate Gaussian distributions, the KL divergence can be computed in closed form \cite{hershey2007approximating}:
\begin{align*}
D_{KL}(P||Q) = \frac{1}{2}\left[\log\frac{\det\boldsymbol{\Sigma}_{q}}{\det\boldsymbol{\Sigma}_{p}} + \Tr(\boldsymbol{\Sigma}_{q}^{-1}\boldsymbol{\Sigma}_{p}) - d + (\boldsymbol{\mu}_{p} - \boldsymbol{\mu}_{q})^{T}\boldsymbol{\Sigma}_{q}^{-1}(\boldsymbol{\mu}_{p} - \boldsymbol{\mu}_{q})\right] \numberthis \label{eqn:4}
\end{align*}
where $\boldsymbol{\Sigma}_{p}$ and $\boldsymbol{\Sigma}_{q}$ denote the covariance matrices of $p$ and $q$, $\boldsymbol{\mu}_{p}$ and $\boldsymbol{\mu}_{q}$ the respective mean vectors, and $d$ the number of dimensions of the Gaussian. In the case of diagonal Gaussian distributions, Eq.\ref{eqn:4} reduces to:
\begin{align*}
D_{KL}(P||Q) = \sum_{i=1}^d\left(\log\frac{\sigma_{q,i}}{\sigma_{p,i}} + \frac{\sigma_{p,i}^{2} + (\mu_{q,i} - \mu_{p,i})^{2}}{2\sigma_{q,i}^{2}} - \frac{1}{2} \right) \numberthis \label{eqn:5}
\end{align*}
Even though the KL divergence is not a proper distance metric, as it is not symmetric in its arguments, it can be used to quantify the error of approximation and serve as a metric for comparing different methods.

\subsubsection*{Simulation-Based Calibration}
Simulation-based calibration is a recently proposed method for validating the accuracy of posterior samples generated by a Bayesian sampling method \cite{talts2018validating}. It is based on the so called \textit{self-consistency} of the Bayesian joint distribution. Given a sample from the prior distribution $\tilde{\theta} \sim p(\theta)$ and a sample from the data-generating process $\tilde{x} \sim p(x|\tilde{\theta})$, one can integrate $\tilde{\theta}$ and $\tilde{x}$ out of the Bayesian joint distribution to recover back the prior of $\theta$:
\begin{align*}
p(\theta) &= \int p(\theta,\tilde{\theta},\tilde{x})d\tilde{x}d\tilde{\theta} \numberthis  \label{eqn:6} \\
&= \int p(\theta,\tilde{x}|\tilde{\theta})p(\tilde{\theta})d\tilde{x}d\tilde{\theta} \numberthis \label{eqn:7} \\
&= \int p(\theta|\tilde{x})p(\tilde{x}|\tilde{\theta})p(\tilde{\theta})d\tilde{x}d\tilde{\theta} \numberthis \label{eqn:8}
\end{align*}
If the Bayesian sampling method produces samples from the exact posterior, the equality implied by Eq.\ref{eqn:8} should hold regardless of the particular form of the posterior. Thus, any violation of this equality indicates some error incurred by the sampling method. The authors in \cite{talts2018validating} propose the following algorithm based on rank-statistics for detecting such violations.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{acb.png}
\caption{Second figure}
\end{figure}

%%% Add this line AFTER all your figures and tables
\FloatBarrier

\movie{Type caption for the movie here.}

\movie{Type caption for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}

\movie{A third movie, just for kicks.}

\dataset{dataset_one.txt}{Type or paste caption here.}

\dataset{dataset_two.txt}{Type or paste caption here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

\bibliography{references}

\end{document}

### 2019-06-26 process_models.py

@@ -35,7 +35,7 @@ def simulate_ricker_batch(X, params, n_batch, t_obs):
     
 
 def simulate_ricker(batch_size=64, t_obs=None, t_obs_min=100, t_obs_max=500, low_r=1, high_r=90, 
-                    low_sigma=0.05, high_sigma=0.7, low_phi=0, high_phi=20, to_tensor=True):
+                    low_sigma=0.05, high_sigma=0.7, low_phi=0, high_phi=15, to_tensor=True):
     """
     Simulates and returns a batch of 1D timeseries obtained under the Ricker model.
     ----------

### 2019-06-26 viz.py

@@ -182,7 +182,7 @@ def plot_sbc(model, n_samples, X_test, theta_test, param_names, figsize=(15, 5),
     # Compute ranks (using broadcasting)    
     ranks = np.sum(theta_samples < theta_test, axis=0)
 
-    # Plot hisograms
+    # Plot histograms
     for j in range(len(param_names)):
         sns.distplot(ranks[:, j], kde=False, ax=axarr[j], rug=True)
         axarr[j].set_title(param_names[j])

### 2019-06-28 latex/paper_main.tex
Levy and Ricker ready

@@
 Then, we present applications to intractable models from cognitive science, population dynamics, epidemiology, and genetics and demonstrate state-of-the art parameter recovery.
+Across the examples, we introduce multiple tools to validate the performance of our method.
 The outline of the remaining manuscript is as follows: The \textbf{Methods} section introduces the main building blocks of the new method and summarizes the main steps as pseudocode.
@@
 Finally, the \textbf{Discussion} section lists the advantages of the current method, treats some potential pitfalls and explores future research vistas.
-Python code and data for all applications are freely available as Jupyter notebooks at \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN} and as a small library based on \textit{TensorFlow} \cite{abadi2016tensorflow} for creating and training custom invertible networks with GPU-support.
+Python code and simulation scripts for all current applications are freely available as Jupyter notebooks at \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN} and as a small library based on \textit{TensorFlow} \cite{abadi2016tensorflow} for creating and training custom invertible networks with GPU support, along with some validation tools.
 
@@
 \begin{figure}
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{acb.png}
@@
   %
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{acbchain.png}
@@
 
-On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\theta}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
+On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\phi}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
 In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the life sciences, since they tend to be unimodal and relatively simple (in contrast to the distribution of natural images or words in a spoken language).
@@
 In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of challenging models from population dynamics, cognitive science, epidemiology, and genetics.
-Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.All examples are implemented in Python using the \textit{TensorFlow} library for automatic differentiation with GPU support.
+Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.All examples are implemented in Python using the \textit{TensorFlow} library.
 
@@
 
-The results on the toy example are depicted in Figure.XXX.
-The approximate posterior means show negligible deviations from the analytic posterior means as quantified by very small $NRMSE$s (as small as XXX) and very high $R^{2}$ (as high as $1.0$) over all numbers of data points used for training.
-This suggests near-perfect reconstruction of the true posterior means.
+The results on the toy Bayesian regression are depicted in \autoref{fig:Fig.3}.
+The approximate posterior means show negligible deviations from the analytic posterior means as quantified by very small NRMSEs (as small as $0.002$) and very high $R^{2}$ (as high as $1.0$) over all $n$ sampled during training.
+This suggests near-perfect estimation of the true posterior means.
+Further, the estimates become increasingly more accurate, as the number of data points $n$ increases.
+An inspection of the posterior variances over the different $n$ reveals that the estimated variance follows closely the decrease in analytic variance with increasing $n$.
+However, the analytic variance is slightly underestimated at smaller $n$ and slightly overestimated at larger $n$.
+This pattern is also revealed by the KL divergence plot (see \textbf{SI}).
+This result might be attributable to an underexpressive summary network; another possibility is that the networks need to be trained longer with smaller learning rate decay.
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{Regression_joint.png}
+    \caption{Analytic vs. estimated means joint plot ($n=500$)}
+    \label{fig:Fig.3a}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+	\begin{subfigure}{.95\textwidth}
+ 		\includegraphics[width=\textwidth]{Regression_density.png}
+    	\caption{Full analytic vs. estimated posteriors ($n=500$)}
+    \label{fig:Fig.3b}
+	\end{subfigure}
+	\begin{subfigure}{.95\textwidth}
+ 		\includegraphics[width=\textwidth]{Regression_scatter.png}
+    	\caption{Analytic means recovery ($n=500$)}
+    \label{fig:Fig.3c}
+	\end{subfigure}
+\end{subfigure}
+\begin{subfigure}{.47\textwidth}
+ 		\includegraphics[width=\textwidth]{Regression_metrics.png}
+    	\caption{Performance over all $n$}
+    \label{fig:Fig.3d}
+\end{subfigure}
+\begin{subfigure}{.52\textwidth}
+ 		\includegraphics[width=\textwidth]{Regression_variance.png}
+    	\caption{Analytic vs. estimated variance over all $n$}
+    \label{fig:Fig.3e}
+\end{subfigure}
+\caption[short]{Results on the Bayesian toy regression example.
+\textbf{(a)} Analytic vs.
+estimated posterior means on the validation sample.
+We observe a near-perfect overlap between analytic and estimated posterior means and no spurious covariances; \textbf{(b)} Some example draws from the estimated posteriors.
+Visual inspections reveals a close match between analytic and estimated posteriors; \textbf{(c)} Posterior mean recovery is almost perfect at $n=500$; \textbf{(d)} NRMSE and $R^{2}$ over all $n$.
+Performance improves with increasing number of data points; \textbf{(e)} Analytic vs.
+estimated variance over all $n$.
+The analytic variance is slightly underestimated at lower $n$ and slightly overestimated at higher $n$.} \label{fig:Fig.3}
+\end{figure}
 
@@
 The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+What if the data does not contain information about a given parameter?
+In this case, any good estimation method should detect this, and return the prior of the particular parameter.
+To test this, we add a random uniform variable $u \sim \mathcal{U}(0, 1)$ to the parameter vector $\boldsymbol{\theta}$ and train the model with this additional dummy variable.
+Then, we inspect the posterior of $u$ for uniformity.
 
-The results on the Ricker model are depicted in Figure XXX... TODO
+The results on the Ricker model are depicted in \autoref{fig:Fig.3}.
+As evident from the graphics, the parameter recovery becomes better when more time points with data are available (\autoref{fig:Fig.4c}).
+At $T=500$, the NRMSEs range between $0.015$ and $0.063$, and the $R^{2}$ metrics between $0.997$ and $0.952$, indicating very good recovery of the posterior means (\autoref{fig:Fig.4a}).
+The parameter $\sigma$ seems to be hardest to recover.
+Inspecting the full posteriors, we further see that the posterior distribution of the dummy noise variable $u$ closely resembles the prior, as expected due to the complete lack of mutual information between the data $x$ and $u$ (\autoref{fig:Fig.4b}).
+Finally, the plots of the rank statistic computed for SCB suggest no systematic distortions of the posterior across all parameters (\autoref{fig:Fig.4d}).
+Interpreting deviations from uniformity according to \cite{talts2018validating}, the approximate posteriors of $r$ and $\rho$ slightly underestimate the true posterior means, whereas the approximate posterior of $\sigma$ tends to overestimate the true posterior variance.
+These deviations appear to be due to the fact that recovery worsens at extreme values of the parameters.
+This is unsurprising, as the data generated with these parameters is highly implausible, which in some cases might even render a model unidentifiable.
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Ricker_scatter.png}
+    \caption{Parameter recovery ($T=500$)}
+    \label{fig:Fig.4a}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Ricker_density.png}
+    \caption{Example posterior samples ($T=500$)}
+    \label{fig:Fig.4b}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Ricker_metrics.png}
+    \caption{Performance over all $T$s}
+    \label{fig:Fig.4c}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Ricker_sbc.png}
+    \caption{Simulation-based calibration (SBC)}
+    \label{fig:Fig.4d}
+\end{subfigure}
+\caption[short]{Results on the Ricker model.
+\textbf{(a)} Parameter recovery for the maximum number of generations used during training ($T=500$); \textbf{(b)} Example posteriors for three test datasets.
+We observe that the posterior of the uniform noise variable $u$ is equal to the prior, i.e., the method detects that no information is present in data for this variable; \textbf{(c)} NRMSE and $R^{2}$ performance metrics over all $T$s used in training.
+We observe that the recovery remains good over all $T$s, and becomes progressively better as more data is available; \textbf{(d)} Plots of the rank statistics indicative of the accuracy of the full posterior.
+Accordingly, the approximate posteriors of $r$ and $\rho$ slightly underestimate the true posterior means, whereas the approximate posterior of $\sigma$ tends to overestimate the true posterior variance.} \label{fig:Fig.4}
+\end{figure}
 
@@
 
-The results on the LFM model are depicted in Figure XXX.
-First, we note that our method achieves state-of-the art parameter recovery, with all $R^{2}$ values above $.97$ and $NRMSE$ below...
-Some of the parameters are almost perfectly reconstructed.
-Describe other stuff...
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Levy_scatter.png}
+    \caption{Parameter recovery ($n=1000$)}
+    \label{fig:Fig.5a}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Levy_density.png}
+    \caption{Example posterior samples ($n=1000$)}
+    \label{fig:Fig.5b}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Levy_metrics.png}
+    \caption{Performance over all trial numbers}
+    \label{fig:Fig.5c}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Levy_sbc.png}
+    \caption{Simulation-based calibration (SBC)}
+    \label{fig:Fig.5d}
+\end{subfigure}
+\caption[short]{Results on the LFM model.
+\textbf{(a)} Parameter recovery for the maximum number of trials used during training ($n=1000$); \textbf{(b)} Example posteriors for three test datasets; \textbf{(c)} NRMSE and $R^{2}$ performance metrics over all $n$ trials used during training.
+Again, we observe that the recovery remains good over overall, and becomes progressively better as more data is available; \textbf{(d)} Plots of the rank statistics indicative of the accuracy of the full posterior.
+Accordingly, the approximate posteriors tend to overestimate the true posterior variance.} \label{fig:Fig.5}
+\end{figure}
+
+The results on the LFM model are depicted in \autoref{fig:Fig.4}.
 
@@
 
-\subsection*{Supporting Information Appendix (SI)}
-
-Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.
-PNAS will publish SI uncomposed, as the authors have provided it.
-Additional details can be found here: \href{https://www.pnas.org/page/authors/submission}{policy on SI}.
-For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.
-The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-
-Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-
 \section*{Discussion}
@@
 We demonstrated the utility of the method by applying it to models from various scientific domains exhibiting various data formats and data-generating mechanisms.
-Further, we explored two possible training approaches suitable for different simulation scenarios, namely a reference-table approach and an active learning approach.
+Further, we explored two possible training approaches suitable for different simulation scenarios, namely an active learning approach, and a reference-table approach.
 Both training approaches lead to excellent recovery of the true parameters throughout the examples considered in the current work.

### 2019-06-28 latex/paper_supplementary.tex
Levy and Ricker ready

@@
 
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \graphicspath{{figures/}}
@@
 Thus, any violation of this equality indicates some error incurred by the sampling method.
-The authors in \cite{talts2018validating} propose the following algorithm based on rank-statistics for detecting such violations.
+The authors of \cite{talts2018validating} propose the following algorithm relying on rank-statistics for visually detecting such violations:
+
+\begin{algorithm}
+\caption{Simulation-based calibration}\label{alg:1}
+\begin{algorithmic}[1]
+\State {Lol}
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection*{Model details}
+
+\subsubsection*{Simulation of the Ricker model}
+We place the following uniform priors over the Ricker model parameters:
+\begin{align*}
+\rho &\sim \mathcal{U}(0,15) \numberthis \\
+r &\sim \mathcal{U}(1,90)  \numberthis \\
+\sigma &\sim \mathcal{U}(0.05,0.7)  \numberthis 
+\end{align*}
+These ranges are very broad, as datasets generated by extreme parameter values appear implausible in real-world scenarios.
+Nevertheless, we stick to broad priors for training, even though parameter recovery might degrade at the extremes.
+Figure XXX depicts some Ricker datasets generated by parameters drawn from the specified prior.
+
+\subsubsection*{Simulation of the Lévy-Flight Model}
+The following uniform priors over the LFM parameters are used for simulation:
+\begin{align*}
+v_{0} &\sim \mathcal{U}(0, 6) \numberthis \\
+v_{1} &\sim \mathcal{U}(-6, 0) \numberthis \\
+zr &\sim \mathcal{U}(0.3, 0.7) \numberthis \\
+a &\sim \mathcal{U}(0.6, 3) \numberthis \\
+t_{0} &\sim \mathcal{U}(0.3, 1) \numberthis \\
+\alpha &\sim \mathcal{U}(1, 2) \numberthis 
+\end{align*}
+These priors are broad enough to cover the range of realistic reaction times observed in choice RT experiments.
+
+\subsubsection*{Simulation of the stochastic SIR model}
+We place the following uniform priors over the two rate parameters of the stochastic SIR model:
+\begin{align*}
+\beta &\sim \mathcal{U}(0.01, 1) \numberthis \\
+\gamma &\sim \mathcal{U}(0.01,\beta)  \numberthis 
+\end{align*}
+Figure XXX depicts some SIR datasets generated by parameters drawn from the specified prior.
+
 \begin{figure}

### 2019-06-28 inn_utils.py

@@ -154,8 +154,8 @@ def train_loop_dataset(model, optimizer, dataset, batch_size, p_bar=None, clip_v
     return losses
 
 
-def compute_metrics(model, param_names, simulate_fun, n_test, 
-                    n_samples_posterior, transform=None, n_min=100, n_max=1000):
+def compute_metrics(model, param_names, simulate_fun, n_test, n_samples_posterior, 
+                    p_bar=None, transform=None, n_min=100, n_max=1000):
     """
     Plots a given metric for different numbers of datapoints.
     ---------
@@ -187,13 +187,15 @@ def compute_metrics(model, param_names, simulate_fun, n_test,
     # For each possible number of data points
     for n_points in ns:
         # Generate data
-        X_test, theta_test = simulate_fun(n_test, n_trials=n_points)
+        X_test, theta_test = simulate_fun(n_test, n_points=n_points)
         if transform is not None:
             X_test, theta_test = transform(X_test, theta_test)
         theta_test = theta_test.numpy()
 
         # Sample from approx posterior and compute posterior means
-        theta_approx_means = model.sample(X_test, n_samples_posterior, to_numpy=True).mean(axis=0)
+        theta_approx = model.sample(X_test, n_samples_posterior, to_numpy=True)
+        theta_approx_means = np.mean(theta_approx, axis=0)
+        theta_approx_vars = np.var(theta_approx, axis=0, ddof=1)
         
         # --- Plot true vs estimated posterior means on a single row --- #
         for j, name in enumerate(param_names):
@@ -204,9 +206,12 @@ def compute_metrics(model, param_names, simulate_fun, n_test,
             # Compute R2
             r2 = r2_score(theta_test[:, j], theta_approx_means[:, j])
             # Compute posterior variance
-            var = np.var(theta_approx_means[:, j], ddof=1)
+            var = np.mean(theta_approx_vars[:, j])
             # Add to dict
             metrics['nrmse'][name].append(nrmse)
             metrics['r2'][name].append(r2)
             metrics['var'][name].append(var)
+            
+        if p_bar is not None:
+            p_bar.update(1)
     return ns, metrics

### 2019-06-28 process_models.py

@@ -34,7 +34,7 @@ def simulate_ricker_batch(X, params, n_batch, t_obs):
         X[i, :] = simulate_ricker_single(t_obs, params[i, 0], params[i, 1], params[i, 2])
     
 
-def simulate_ricker(batch_size=64, t_obs=None, t_obs_min=100, t_obs_max=500, low_r=1, high_r=90, 
+def simulate_ricker(batch_size=64, n_points=None, t_obs_min=100, t_obs_max=500, low_r=1, high_r=90, 
                     low_sigma=0.05, high_sigma=0.7, low_phi=0, high_phi=15, to_tensor=True):
     """
     Simulates and returns a batch of 1D timeseries obtained under the Ricker model.
@@ -42,7 +42,7 @@ def simulate_ricker(batch_size=64, t_obs=None, t_obs_min=100, t_obs_max=500, low
 
     Arguments:
     batch_size : int -- number of Ricker processes to simulate
-    t_obs      : int -- length of the observed time-series
+    n_points   : int -- length of the observed time-series
     low_r      : float -- lower bound for the uniform prior on r
     high_r     : float -- upper bound for the uniform prior on r
     low_sigma  : float -- lower bound for the uniform prior on sigma
@@ -59,16 +59,16 @@ def simulate_ricker(batch_size=64, t_obs=None, t_obs_min=100, t_obs_max=500, low
     """
     
     # Sample t_obs, if None given
-    if t_obs is None:
-        t_obs = np.random.randint(low=t_obs_min, high=t_obs_max+1)
+    if n_points is None:
+        n_points = np.random.randint(low=t_obs_min, high=t_obs_max+1)
 
     # Prepare placeholders
     theta = np.random.uniform(low=[low_r, low_sigma, low_phi], 
                           high=[high_r, high_sigma, high_phi], size=(batch_size, 3))
-    X = np.zeros((batch_size, t_obs))
+    X = np.zeros((batch_size, n_points))
 
     # Simulate a batch from the Ricker model
-    simulate_ricker_batch(X, theta, batch_size, t_obs)
+    simulate_ricker_batch(X, theta, batch_size, n_points)
     if to_tensor:
         return tf.convert_to_tensor(X[:, :, np.newaxis], dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)
     return X[:, :, np.newaxis]

### 2019-06-28 simulate_diffusion.py

@@ -47,7 +47,7 @@ def simulate_batch_diffusion_p(x, params):
                                          params[i, 4], params[i, 5], params[i, 6], 
                                          params[i, 7], params[i, 8], 0.001, 5000)
 
-def simulate_diffusion(batch_size, pbounds, n_trials=None, n_cond=2, 
+def simulate_diffusion(batch_size, pbounds, n_points=None, n_cond=2, 
                        to_tensor=True, cond_coding=False, n_trials_min=100, n_trials_max=1000):
     """Simulates batch_size datasets from the full Ratcliff diffusion model."""
     
@@ -55,8 +55,8 @@ def simulate_diffusion(batch_size, pbounds, n_trials=None, n_cond=2,
     n_params = len(pbounds)
 
     # Sample number of trials, if None given
-    if n_trials is None:
-        n_trials = np.random.randint(n_trials_min, n_trials_max+1)
+    if n_points is None:
+        n_points = np.random.randint(n_trials_min, n_trials_max+1)
 
     # Extract parameter bounds
     lower_bounds = [pbounds['v1'][0], pbounds['v2'][0], pbounds['sv'][0], pbounds['zr'][0],
@@ -69,29 +69,13 @@ def simulate_diffusion(batch_size, pbounds, n_trials=None, n_cond=2,
 
     # Draw from priors
     theta_batch  = np.random.uniform(low=lower_bounds, high=upper_bounds,  size=(batch_size, n_params)).astype(np.float32)
-    X_batch = np.zeros((batch_size, n_trials, n_cond), dtype=np.float32)
+    X_batch = np.zeros((batch_size, n_points, n_cond), dtype=np.float32)
     simulate_batch_diffusion_p(X_batch, theta_batch)
 
     # Return in specified format (condition coding or just stack, tf.Tensor or np.array)
     if cond_coding:
         X_batch = np.stack(([np.c_[X_batch[:, :, 0], X_batch[:, :, 1]], 
-                             np.c_[np.zeros((batch_size, n_trials)), np.ones((batch_size, n_trials))]]), axis=-1)
+                             np.c_[np.zeros((batch_size, n_points)), np.ones((batch_size, n_points))]]), axis=-1)
     if to_tensor:
[…]

### 2019-06-28 viz.py

@@ -1,15 +1,18 @@
 import matplotlib.pyplot as plt
+from matplotlib.ticker import FormatStrFormatter
 import seaborn as sns
 import numpy as np
 from sklearn.metrics import r2_score
 
-# Plot settings
-plt.rcParams['font.size'] = 14
 
 
-def plot_true_est_scatter(model, X_test, theta_test, n_samples, 
-                          param_names, figsize=(20, 4), show=True, filename=None):
+def plot_true_est_scatter(model, X_test, theta_test, n_samples, param_names, 
+                          figsize=(20, 4), show=True, filename=None, font_size=12):
     """Plots a scatter plot with abline of the estimated posterior means vs true values."""
+
+
+    # Plot settings
+    plt.rcParams['font.size'] = font_size
 
     # Convert true parameters to numpy
     theta_test = theta_test.numpy()
@@ -100,11 +103,98 @@ def plot_losses(losses, figsize=(15, 5), show=True):
         plt.show()
 
 
+def plot_metrics(metrics, ns, param_names, figsize=(12, 4), show=True, 
+                 xlabel=r'$n$', filename=None, font_size=12):
+    """
+    Plots the nrmse and r2 for all parameters
+    """
+
+    # Plot settings
+    plt.rcParams['font.size'] = font_size
+    
+    # Initialize figure
+    f, axarr = plt.subplots(1, 2, figsize=figsize)
+
+    for i, metric in enumerate(['nrmse', 'r2']):
+        for p in param_names:
+            sns.lineplot(ns, metrics[metric][p], label=p, markers=True, dashes=False, ax=axarr[i])
+            
+        if metric == 'nrmse':
+            axarr[i].set_ylabel('NRMSE')
+        elif metric == 'r2':
+            axarr[i].set_ylabel(r'$R^{2}$')
+        axarr[i].set_xlabel(xlabel)
+        
+        axarr[i].spines['right'].set_visible(False)
+        axarr[i].spines['top'].set_visible(False)
+        axarr[i].legend(fontsize=12)
+    
+    f.tight_layout()
+        
+    if show:
+        plt.show()
+    
+    if filename is not None:
+        f.savefig("figures/{}_metrics.png".format(filename), dpi=600, bbox_inches='tight')
+
+
+def plot_variance(variances, ns, param_names, figsize=(12, 4), show=True, 
+                  xlabel=r'$n$', filename=None, tight=True, std=False, font_size=12):
+    """
[…]

### 2019-06-28 latex/paper_main.tex
Fix SIR

@@
 For the following application, we set $d=4$, and $a=b=1$.
-The design matrices for each iteration contain a variable number $n$ of \textit{i.i.d.} data points drawn from a unit Gaussian $\mathcal{N}_{4}(\boldsymbol{0},\boldsymbol{I})$.
+The design matrices for each iteration contain a variable number $n$ of \textit{i.i.d.} data points drawn from a unit Gaussian $\boldsymbol{x}^{(i)} \sim \mathcal{N}_{4}(\boldsymbol{0},\boldsymbol{I})$ for $i=1,...,n$.
 The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(50, 500)$ at each training iteration (Lines $2$-$9$ of \textbf{Algorithm} \ref{alg:1}).
@@
 \begin{align*}
-N_{t+1} &= rN_{t}\exp(-N_{t} + \epsilon_{t}) \numberthis \\
-x_{t} &\sim Pois(\rho N_{t}) \numberthis
+x_{t} &\sim Pois(\rho N_{t}) \numberthis \\
+N_{t+1} &= rN_{t}e^{-N_{t} + \epsilon_{t}} \numberthis 
 \end{align*}
@@
 The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+
 What if the data does not contain information about a given parameter?
@@
 
-The results on the LFM model are depicted in \autoref{fig:Fig.4}.
+The results on the LFM model are depicted in \autoref{fig:Fig.4}. XXX
 
@@
 Another potential shortcoming of the method is the seemingly overwhelming number of hyperparameters that might require fine-tuning by the user for optimal performance on a given task.
-However, we observe that many of the default values of hyperparameters are sufficient to achieve good performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training.
+However, we observe that many of the default hyperparameter values are sufficient to achieve excellent performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training, even if the model to be learned is relatively simple.
 We expect that a single architecture should be able to perform well on almost all models from a given domain (i.e., a single architecture for decision-making models).

### 2019-06-28 latex/paper_supplementary.tex
Fix SIR

@@
 
+\usepackage{bbm}
 \usepackage{algorithm}
@@
 Thus, any violation of this equality indicates some error incurred by the sampling method.
-The authors of \cite{talts2018validating} propose the following algorithm relying on rank-statistics for visually detecting such violations:
+The authors of \cite{talts2018validating} propose \textbf{Algorithm} \ref{alg:1} for visually detecting such violations:
 
 \begin{algorithm}
-\caption{Simulation-based calibration}\label{alg:1}
+\caption{Simulation-based calibration (SBC) for a single parameter $\theta$ }\label{alg:1}
 \begin{algorithmic}[1]
-\State {Lol}
-
+\For{$i = 1,...,n$}
+\State {Sample $\tilde{\theta}^{(i)} \sim p(\theta)$}
+\State {Simulate a dataset $\boldsymbol{x}^{(i)} = q(\tilde{\theta}^{(i)})$}
+\State {Draw posterior samples $\{\theta^{(l)}\}_{l=1}^{L} \sim p(\theta|\boldsymbol{x}^{(i)})$} 
+\State {Compute rank statistic $r^{(i)} = \sum_{l=1}^{L}\mathbbm{1}_{[\theta^{(l)}<\tilde{\theta}^{(i)}]}$}
+\State {Store $r^{(i)}$}
+\EndFor
+\State {\textbf{end for}}
+\State {Analyze the histogram for uniformity}
 \end{algorithmic}
 \end{algorithm}
+
+\textbf{Algorithm} \ref{alg:1} is justified, since Eq.\ref{eqn:8} implies that the rank statistic defined in line $5$ should be uniformly distributed.
+Hence, any deviations from uniformity indicate some error in the approximate posterior.
 

### 2019-06-28 process_models.py

@@ -105,7 +105,8 @@ def simulate_sir_single(beta, gamma, t_max=500, N=1000):
     return np.array([S, I, R]).T
 
 @jit
-def simulate_sir(batch_size, low_beta=0.01, high_beta=1., low_gamma=0.01, t_max=500, N=1000, to_tensor=True):
+def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gamma=0., 
+                 t_min=200, t_max=500, N=1000, to_tensor=True):
     """
     Simulates and returns a batch of timeseries obtained under the SIR model.
     ----------
@@ -125,16 +126,20 @@ def simulate_sir(batch_size, low_beta=0.01, high_beta=1., low_gamma=0.01, t_max=
               (tf.Tensor of shape (batch_size, t_obs, 1), tf.Tensor of shape (batch_size, 3)) --
               a batch or time series generated under a batch of Ricker parameters
     """
+
+    # Select T
+    if n_points is None:
+        n_points = np.random.randint(low=t_min, high=t_max+1)
     
     # Prepare X and theta
-    X = np.zeros((batch_size, t_max, 3))
+    X = np.zeros((batch_size, n_points, 3))
     beta_samples = np.random.uniform(low=low_beta, high=high_beta, size=batch_size)
     gamma_samples = np.random.uniform(low=low_gamma, high=beta_samples)
     theta = np.c_[beta_samples, gamma_samples]
     
     # Run the SIR simulator for # batch_size
     for j in prange(batch_size):
-        X[j] = simulate_sir_single(theta[j, 0], theta[j, 1], t_max=t_max, N=N)
+        X[j] = simulate_sir_single(theta[j, 0], theta[j, 1], t_max=n_points, N=N)
     
     if to_tensor:
         return tf.convert_to_tensor(X, dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)

### 2019-06-28 viz.py

@@ -236,7 +236,7 @@ def plot_true_est_posterior(model, n_samples, param_names, n_test=None, data_gen
                 axarr[i, j].set_title(param_names[j])       
             
             if i == 0 and j == 0:
-                f.legend(loc='lower center', bbox_to_anchor=(0.5, -0.02), shadow=True, ncol=3, fontsize=10, borderaxespad=1)
+                f.legend(loc='lower center', bbox_to_anchor=(0.5, -0.03), shadow=True, ncol=3, fontsize=10, borderaxespad=1)
                 #axarr[i, j].legend(fontsize=10)
                 
     if tight:

### 2019-06-29 latex/paper_main.tex
Add conv net to SIR

@@
 
-Minimizing Eq.
-\ref{eqn:14} can be interpreted as searching for the optimal neural network parameters $\boldsymbol{\phi}^{*}$ which maximize the probability of model parameters $\boldsymbol{\theta}$ given data $\boldsymbol{x}$.
+Minimizing Eq.\ref{eqn:14} can be interpreted as searching for the optimal neural network parameters $\boldsymbol{\phi}^{*}$ which maximize the probability of model parameters $\boldsymbol{\theta}$ given data $\boldsymbol{x}$.
 This is exactly the probability we are concerned with in Bayesian inference.
@@
 We set the weight regularization parameter $\tau$ to a value of $10^{-5}$.
-We perform $100000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
-Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter but use a cINN with 10 ACBs for each example in this paper (see \textbf{SI} for details of the cINN).
-Concerning the data generation step, we use two training approaches.
-The first follows the classical approximate Bayesian computation approach to create a large “reference table” or grid of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$.
+We perform $50000$ to $100000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
+Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINN).
+All networks were trained on a single-GPU machine equipped with NVIDIA GTX1060 graphics card.
+Regarding the data generation step, we use two training approaches.
+
+The first approach follows the classical approximate Bayesian computation approach to create a large “reference table” or grid of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$.
 The reference table is then used as training data for the neural network and training continues for a pre-specified number of epochs through the entire reference table.
@@
 This training approach separates the simulation from the training phase but can incur large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
+
 The second approach follows a different strategy, which is used mainly in the field of active learning.
@@
 Moreover, training can continue as long as the network keeps improving (i.e., the loss keeps decreasing), since overfitting in the classical sense is nearly impossible.
-However, if the simulations are computationally expensive and researchers need to experiment with multiple models, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
-TODO - Describe for which!
+However, if the simulations are computationally expensive and researchers need to experiment with different networks or training hyperparameters, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
 
@@
 As a proof-of-concept, we demonstrate the utility of our method in recovering the true analytic posteriors of the regression coefficients of a conjugate Bayesian regression model.
-To set the setting, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$ We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, i.e.,: $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
+To set the setting, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$.
+We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, i.e., $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
 We place a $d$-dimensional diagonal Gaussian prior on the regression coefficients centered at $0$: $\boldsymbol{\theta} \sim \mathcal{N}_{d}(\boldsymbol{0},b^{-1}\boldsymbol{I})$ where $b$ is the precision of the prior ($b \equiv 1/\sigma_{\boldsymbol{\theta}}^{2}$).
@@
 The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(100, 1000)$ at each training iteration.
-
 \begin{figure}[H]
@@
 
-The results on the LFM model are depicted in \autoref{fig:Fig.4}. XXX
+The results on the LFM model are depicted in \autoref{fig:Fig.4}.
+To our knowledge, this is the first Bayesian treatment of the LFM, as its intractability makes traditional methods prohibitively slow.
+We observe state-of-the art recovery of all LFM parameters with NRMSEs ranging between $0.008$ and $0.048$ and $R^{(2)}$ between $0.972$ and $0.999$.
+Further, estimation remains very good across all trial sizes, with goodness increasing as more trials become available.
+The parameter $\alpha$ appears to be most challenging to estimate, requiring more data for good estimation, whereas the non-decision time parameter $t_{0}$ is almost perfectly reconstructed for all trial sizes.
+Last, inspecting the SCB plots, we notice that the histograms tend to be slightly peaked, indicating a slight overestimation of the posterior variance \cite{talts2018validating}.
 
@@
 Once trained, the networks can be used and reused to perform inference on large numbers of datasets within seconds and across a given research domain.
-Indeed, there are many instances of research domains where a single model is extensively explored and independently fitted by multiple researches (CITATION DIFFUSION MODELLING, POPULATION DYNAMICS, NEUROSCIENCE) to test scientific theories about the modelled process.
-These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for inference on different datasets.
+Indeed, there are many instances of research domains where a single model is extensively explored and independently fitted by multiple researches to test scientific hypotheses \cite{voss2019sequential, zappia2017splatter, ratcliff2008diffusion, de2002fitting}.
+These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for rapid inference on different datasets.
 In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
 Note, however, that our method does not need to store training data on disk, since the "knowledge" about the relationship between data and parameters is compressed into the networks' weights.
-Moreover, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern graphical processing units (GPUs).
+Moreover, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern GPUs.
 
@@
 In order to ease the understanding and independent application of the method, we provide fully functioning code to reproduce and study all of the examples tackled in this paper (\href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}).
+In addition, we also provide implementation of all tools for performance validation used throughout the paper.
 Moreover, we are currently developing a general user-friendly software, which should abstract away most of the methodological complexities from the user.

### 2019-06-29 latex/paper_supplementary.tex
Add conv net to SIR

@@
 \section*{Results}
+Code for all examples used in this paper is available at \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
@@
 \State {\textbf{end for}}
-\State {Analyze the histogram for uniformity}
+\State {Create a histogram of $\{r^{(i)}\}_{i=1}^{n}$ and analyze it for uniformity}
 \end{algorithmic}
@@
 \end{align*}
-Figure XXX depicts some SIR datasets generated by parameters drawn from the specified prior.
+We decide to Figure XXX depicts some SIR datasets generated by parameters drawn from the specified priors.
 

### 2019-06-29 process_models.py

@@ -106,7 +106,7 @@ def simulate_sir_single(beta, gamma, t_max=500, N=1000):
 
 @jit
 def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gamma=0., 
-                 t_min=200, t_max=500, N=1000, to_tensor=True):
+                 t_min=200, t_max=500, N=1000, normalize=True, to_tensor=True):
     """
     Simulates and returns a batch of timeseries obtained under the SIR model.
     ----------
@@ -140,7 +140,9 @@ def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gam
     # Run the SIR simulator for # batch_size
     for j in prange(batch_size):
         X[j] = simulate_sir_single(theta[j, 0], theta[j, 1], t_max=n_points, N=N)
-    
+
+    if normalize:
+        X /= N
     if to_tensor:
         return tf.convert_to_tensor(X, dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)
     return X, theta

### 2019-06-29 latex/paper_main.tex
Sir ready

@@
 \begin{figure}
-  \begin{subfigure}[b]{1.0\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{training.png}
@@
   \end{subfigure}
-  %
-  \begin{subfigure}[b]{1.0\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \includegraphics[width=\textwidth]{inference.png}
@@
 for $t = 1,...,T$ where $N_{t}$ is the expected number of individuals at time $t$, $r$ is the growth rate, $\rho$ is a scaling parameter and $\epsilon_{t} \sim \mathcal{N}(0, \sigma^{2})$ is random Gaussian noise \cite{mestdagh2018prepaid}.
-The Ricker model has no known likelihood function and is thus a suitable candidate for likelihood-free inference.
+The likelihood function for the Ricker model is not available in closed-form, and the model is known to exhibit chaotic behavior.
+Thus, it is a suitable candidate for likelihood-free inference.
 The parameter estimation task is thus to recover $\boldsymbol{\theta}=(\rho,r,\sigma)$ from the observed one-dimensional time-series data $\boldsymbol{x}=(x_{1},x_{2},...,x_{T})$ where each $x_{t} \in \mathbb{N}$.
@@
 \textbf{(a)} Parameter recovery for the maximum number of trials used during training ($n=1000$); \textbf{(b)} Example posteriors for three test datasets; \textbf{(c)} NRMSE and $R^{2}$ performance metrics over all $n$ trials used during training.
-Again, we observe that the recovery remains good over overall, and becomes progressively better as more data is available; \textbf{(d)} Plots of the rank statistics indicative of the accuracy of the full posterior.
+Again, we observe that the recovery remains overall very good, and becomes progressively better as more data is available; \textbf{(d)} Plots of the rank statistics indicative of the accuracy of the full posterior.
 Accordingly, the approximate posteriors tend to overestimate the true posterior variance.} \label{fig:Fig.5}
@@
 To our knowledge, this is the first Bayesian treatment of the LFM, as its intractability makes traditional methods prohibitively slow.
-We observe state-of-the art recovery of all LFM parameters with NRMSEs ranging between $0.008$ and $0.048$ and $R^{(2)}$ between $0.972$ and $0.999$.
+We observe state-of-the art recovery of all LFM parameters with NRMSEs ranging between $0.008$ and $0.048$ and $R^{2}$ between $0.972$ and $0.999$.
 Further, estimation remains very good across all trial sizes, with goodness increasing as more trials become available.
@@
 where $S + I + R = N$ give the number of susceptible, infected, and recovered individuals, respectively.
-The parameter $\beta$ controls the transition rate from being susceptible to infected, and $\gamma$ controls the transition rate from being infected to recovered.
+The parameter $\beta$ controls the transition rate from being susceptible to infected, and $\gamma$ controls the transition rate from being infected to recovered (see \autoref{fig:Fig.6}).
 The number of individuals moving from $S$ to $I$, given by $\triangle N_{SI}$, and the number of people moving from $I$ to $R$, given by $\triangle N_{IR}$, over a time interval $\triangle t$ are modeled as binomial random variables.
@@
 During training of the networks, we simulate time-series from the stochastic SIR model with varying lengths.
-The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 200)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
-This range is chosen since we observed that the system reaches an equilibrium in almost all parameter settings after $T > 200$.
+The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(200, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+Usually, at lower $T$s, the system has not converged to an equilibrium (i.e., not all individuals have moved from being $I$ to $R$).
+Thus, it is especially interesting to see if our method can recover the parameters, even if the process dynamics are still unfolding over time.
 
-The results on the SIR model are depicted in Figure XXX.
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{SIR_plot.png}
+    \caption{Example SIR timeseries with $\beta=0.5$ and $\gamma=0.2$}
+    \label{fig:Fig.6a}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{SIR_states.png}
+    \caption{State-transition diagram of the SIR model}
+    \label{fig:Fig.6b}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{SIR_scatter.png}
+    \caption{Parameter recovery ($T=500$)}
+    \label{fig:Fig.6c}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \begin{subfigure}{.49\textwidth}
+    	\includegraphics[width=\textwidth]{SIR_metrics.png}
+    	\caption{Performance over all $T$s}
+    	\label{fig:Fig.6d}
+	\end{subfigure}
+	\begin{subfigure}{.49\textwidth}
+    	\includegraphics[width=\textwidth]{SIR_variance.png}
+    	\caption{Posterior variance over all $Ts$}
+    	\label{fig:Fig.6e}
+	\end{subfigure}
+	\begin{subfigure}{.49\textwidth}
+    	\includegraphics[width=\textwidth]{SIR_density.png}
+    	\caption{Example posterior samples ($T=500$}
+    	\label{fig:Fig.6f}
+	\end{subfigure}
+	\begin{subfigure}{.49\textwidth}
+    	\includegraphics[width=\textwidth]{SIR_sbc.png}
+    	\caption{Simulation-based calibration (SBC)}
+    	\label{fig:Fig.6f}
+	\end{subfigure}
+\end{subfigure}
+\caption[short]{Results on the stochastic SIR model.
+\textbf{(a)} Example SIR timeseries generated with $\beta=0.5$ and $\gamma=0.2$; \textbf{(b)} State transition diagram of the SIR model; \textbf{(c)} Parameter recovery depicting also NRMSE and $R^{2}$ metrics; \textbf{(d)} NRMSE and $R^{2}$ performance over all $T$s seen by the networks during training.
+We observe similar patterns as in the previous examples; \textbf{(e)} Posterior variances of the SIR parameters over all $T$s; \textbf{(e)} Example full posteriors of the parameters recovered from three test datasets; \textbf{(f)} Plots of the rank statistic, according to which the posterior of $\beta$ slightly overestimates the true mean.} \label{fig:Fig.6}
+\end{figure}
+
+The results on the SIR model are depicted in \autoref{fig:Fig.6}.
+Similar to the previous examples, we observe very good recovery of the true parameters, with NRMSE at $T=500$ around $0.04$, and $R^{2}$s around $0.98$.
+Further, we observe decent performance even at smaller $T$s, with performance increasing as $T$ increases.
+Conversely, the estimated posterior variance drops as $T$ increases, capturing the fact that more information becomes available to the networks.
+Finally, the SCB plots indicate a slight underestimation of the true posterior means by the estimated posterior means.
+This trend is visible when $\beta$ and $\gamma$ both assume very similar values, and the generated timeseries are far from equilibrium at the final time point $T$.
+Thus, such datasets contain very little information about the parameters given the considered time range of $T_{max}=500$.
 
@@
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $250000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
+We simulate $500000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
 We train the networks for 100 epochs through the entire dataset and evaluate the performance on a separate validation set or matrices.

### 2019-06-29 process_models.py

@@ -1,5 +1,7 @@
 from numba import jit, prange
 import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
 import scipy
 import tensorflow as tf
 
@@ -106,6 +108,7 @@ def simulate_sir_single(beta, gamma, t_max=500, N=1000):
 
 @jit
 def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gamma=0., 
+
                  t_min=200, t_max=500, N=1000, normalize=True, to_tensor=True):
     """
     Simulates and returns a batch of timeseries obtained under the SIR model.
@@ -146,3 +149,28 @@ def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gam
     if to_tensor:
         return tf.convert_to_tensor(X, dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)
     return X, theta
+
+def plot_sir(beta, gamma, n_points=500, figsize=(8, 4), N=1000, filename=None):
+    """
+    Simulates a single SIR process.
+    """
+
+    # Generate SIR dataframes
+    X = simulate_sir_single(beta, gamma, t_max=n_points, N=N)
+    t = np.arange(1, n_points+1)
+
+    # Prepare figure
+    f, ax = plt.subplots(1, 1, figsize=figsize)
+    sns.lineplot(t, X[:, 0], ax=ax, label='Susceptible')
+    sns.lineplot(t, X[:, 1], ax=ax, label='Infected')
+    sns.lineplot(t, X[:, 2], ax=ax, label='Recovered')
+
+    ax.spines['right'].set_visible(False)
+    ax.spines['top'].set_visible(False)
+    ax.set_xlabel(r'Number of time points ($T$)', fontsize=12)
+    ax.set_ylabel('Number of individuals', fontsize=12)
+    ax.legend(fontsize=10)
+
+    # Save if specified
+    if filename is not None:
+        f.savefig("figures/{}_plot.png".format(filename), dpi=600)

### 2019-06-29 viz.py

@@ -227,7 +227,7 @@ def plot_true_est_posterior(model, n_samples, param_names, n_test=None, data_gen
             axarr[i, j].axvline(theta_test[i, j], color='#e55e5e', label='True')
             axarr[i, j].spines['right'].set_visible(False)
             axarr[i, j].spines['top'].set_visible(False)
-            axarr[i, j].xaxis.set_major_formatter(FormatStrFormatter('%.1f'))
+            axarr[i, j].xaxis.set_major_formatter(FormatStrFormatter('%.2f'))
             axarr[i, j].get_yaxis().set_ticks([])
             
             

### 2019-07-01 latex/paper_main.tex
Add seqRNA files

@@
 
-\title{Learning complex models with invertible neural networks: a likelihood-free Bayesian approach}
+\title{Learning complex stochastic models with invertible neural networks: a likelihood-free Bayesian approach}
 
@@
     	\caption{Simulation-based calibration (SBC)}
-    	\label{fig:Fig.6f}
+    	\label{fig:Fig.6g}
 	\end{subfigure}
@@
 \textbf{(a)} Example SIR timeseries generated with $\beta=0.5$ and $\gamma=0.2$; \textbf{(b)} State transition diagram of the SIR model; \textbf{(c)} Parameter recovery depicting also NRMSE and $R^{2}$ metrics; \textbf{(d)} NRMSE and $R^{2}$ performance over all $T$s seen by the networks during training.
-We observe similar patterns as in the previous examples; \textbf{(e)} Posterior variances of the SIR parameters over all $T$s; \textbf{(e)} Example full posteriors of the parameters recovered from three test datasets; \textbf{(f)} Plots of the rank statistic, according to which the posterior of $\beta$ slightly overestimates the true mean.} \label{fig:Fig.6}
+We observe similar patterns as in the previous examples; \textbf{(e)} Posterior variances of the SIR parameters over all $T$s; \textbf{(f)} Example full posteriors of the parameters recovered from three test datasets; \textbf{(g)} Plots of the rank statistic, according to which the posterior of $\beta$ slightly overestimates the true mean.} \label{fig:Fig.6}
 \end{figure}
@@
 The output of the \textit{Splat} simulation is thus a matrix $\boldsymbol{X}$ of \textit{Gene x Cell} counts.
-Figure XXX depicts the \textit{Splat} algorithm for simulating scRNA-seq data.
-
+\autoref{fig:Fig.7a} depicts the \textit{Splat} algorithm for simulating scRNA-seq data \cite{zappia2017splatter}.
+\begin{table} 
+\centering
+\caption{Splat target parameters estimated by our method}
+\begin{tabular}{lll}
+Parameter & Name & Description\\
+\midrule
+$\alpha$ & Mean shape & Shape of the mean gene expression gamma distribution \\
+$\beta$ & Mean rate & Rate of the mean gene expression gamma distribution \\
+$\mu^{L}$ & Library size location & Location of the library size log-normal distribution \\
+$\sigma^{L}$ & Library size scale & Scale of the library size log-normal distribution \\
+$\pi^{O}$ & Outlier probability & Probability that a gene is an expression outlier \\
+$\mu^{O}$ & Outlier location & Location of the expression outlier factor log-normal distribution \\
+$\sigma^{O}$ & Outlier scale & Scale of the expression outlier factor log-normal distribution \\
+$\phi$ & Common BCV & Common BCV dispersion across all genes \\
+\bottomrule
+\end{tabular}
+\label{table:Table 1}
+\end{table}
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $500000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
-We train the networks for 100 epochs through the entire dataset and evaluate the performance on a separate validation set or matrices.
-
-Figure XXX depicts the results...
+We simulate $500000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $50\%$ of the counts are $0$).
+The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
+The parameter estimation task is thus to recover $\boldsymbol{\theta} = (\alpha,\beta,\mu^{L},\sigma^{L},\pi^{O},\mu^{O},\sigma^{O},\phi)$ from an observed count matrix $\boldsymbol{X}$.
+We train the networks for 50 epochs through the entire dataset and evaluate the performance on a separate validation set of $500$ count matrices.
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Splatter_plot.png}
+    \caption{The \textit{Splat} simulation diagram. Adapted after \cite{zappia2017splatter}.}
+    \label{fig:Fig.7a}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Splatter_scatter.png}
+    \caption{Parameter recovery }
+    \label{fig:Fig.7b}
+\end{subfigure}
+\caption[short]{Results on the \textit{Splat} scRNA simulation model. \textbf{(a)} The \textit{Splat} simulation diagram; \textbf{(b)} Parameter recovery.} \label{fig:Fig.7}
+\end{figure}
+The results on the scRNA example are depicted in \autoref{fig:Fig.7}.
+We observe excellent parameter recovery on the validation set.
+\textbf{TODO: Train the full method.
+These results are preliminary}.
 

### 2019-07-04 latex/paper_main.tex
Paper update

@@
 
-\dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about probabilistic phenomena in a probabilistic world.
+\dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about deterministic or probabilistic phenomena.
 In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a forward model $q$ mimicking the process by which manifest data $\boldsymbol{x}$ arise from latent parameters:
@@
 \end{align*}
-While $q$ can represent an arbitrarily complex process by an arbitrarily complicated expression, its functional form is usually guided by a well-founded theoretical framework.
+The functional form of $q$, which can represent an arbitrarily complex process by an arbitrarily complicated expression, is usually guided by a well-founded theoretical framework.
 For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm dictating the rate of gene expression in certain cells.
-Thus, it is only through theoretical embedding that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
-Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, just to name a few.
+Thus, it is only within the context of a theory that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
+Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical, geritz2004mechanistic}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, just to name a few.
 
-Once a mathematical model has been formulated, the next step consists of fitting the model to experimental or observational data and recovering the parameters of interest.
+Once a mathematical model has been formulated, the next step consists in fitting the model to experimental or observational data and recovering the parameters of interest.
 However, estimating the parameters of a mathematical model from data can quickly become one of the most tenacious challenges in applications to real-world problems.
-It is also one of the most important ones to be tackled, since without reliable parameter estimation methods, it is impossible to test the utility of a model, regardless of its sophistication or theoretical appeal.
+It is also one of the most important ones to be tackled, since it is impossible to test the utility of a model without reliable parameter estimation methods, regardless of the model's sophistication or theoretical appeal.
 Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
@@
 Complex models specified by a generative stochastic mechanism do not always provide a closed-form solution for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}.
-This poses great difficulties for Bayesian and frequentist methods alike, since both depend explicitly on the numerical evaluation of a likelihood function as a proxy for assessing model fit to data.
+This poses great difficulties for Bayesian and frequentist methods alike, since both depend on the numerical evaluation of a likelihood function as a proxy for assessing model fit to data.
 Even if a likelihood function is available in closed-form, inference may be prohibitively slow for real-world applications.
-In this case, enforcing simplifying distributional assumptions (i.e., independence or Gaussian assumptions) on the likelihood can increase the speed of inference, but can also lead to model misspecifications and dramatically incorrect estimates.
-Therefore, the need for powerful and reliable likelihood-free estimation methods arises naturally.
+In this case, enforcing simplifying distributional assumptions (i.e., independence or Gaussian assumptions) can increase computational speed, but can also lead to model misspecifications and dramatically incorrect estimates.
+Therefore, the need for powerful and reliable likelihood-free estimation methods becomes obvious.
 
-Likelihood-free methods aim at bypassing the above problems by resorting to a simulation-based approach to parameter estimation and model selection \cite{palestro2018likelihood, turner2014generalized}.
+Likelihood-free methods bypass the above problems by resorting to a simulation-based approach to parameter estimation and model selection \cite{palestro2018likelihood, turner2014generalized}.
 A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate, csillery2010approximate}.
 ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters.
-Thus, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
+Correspondingly, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
 
@@
 
-Recently, ideas from machine learning and deep learning research have entered the field of likelihood-free inference in an attempt to overcome some of the shortcomings of traditional methods \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
+In an attempt to balance the shortcomings of traditional methods, ideas from machine learning and deep learning research have recently entered the field of likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
 The most common approach has been to cast the problem of parameter estimation as a supervised learning task.
@@
 Usually, the dimensionality of the simulated data is reduced by computing summary statistics  with a fixed summary function $h(\boldsymbol{x})$.
-Then, a supervised learning algorithm $f(h(\boldsymbol{x});\boldsymbol{\phi}) = \widehat{\boldsymbol{\theta}}$ with learnable parameters $\boldsymbol{\phi}$ (e.g., linear regression, random forest, neural network) is trained on the simulated data to later output an estimate of the true data generating parameters.
+Then, a supervised learning algorithm $f(h(\boldsymbol{x});\boldsymbol{\phi}) = \widehat{\boldsymbol{\theta}}$ with learnable parameters $\boldsymbol{\phi}$ (e.g., linear regression, random forest, neural network) is trained on the simulated data to output an estimate of the true data generating parameters.
 Thus, $f(h(\boldsymbol{x});\boldsymbol{\phi})$ essentially attempts to “learn” the intractable inverse model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$.
 
-Inspired by previous machine learning approaches, the current work proposes a novel and universal likelihood-free method capable of performing full Bayesian inference on any mathematical process model from which simulations can be obtained.
+Inspired by previous machine learning approaches \cite{hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, the current work proposes a novel and universal likelihood-free method capable of performing full Bayesian inference on any mathematical process model from which simulations can be obtained.
 It treats parameter inference as a task of inverting a generative model and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
@@
 However, in many application, it is not straightforward to settle upon a set of good summary statistics.
-Thus, the summary network completely eliminates the need to manually specify a fixed number of summary statistics and makes the method independent of the format or the size of the data.
+The summary network completely eliminates the need to manually decide on and compute a fixed number of summary statistics and, thus, makes the method independent of the format or the size of the data.
 
@@
 \end{figure}
+
 The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated statistical problems in various scientific domains (\autoref{fig:Fig.1}).
@@
 Last, the trained networks can be shared and reused by multiple researches within a scientific domain, thus removing the need for wasteful computations and fitting a separate model for each and every dataset.
-This pooling of computational resources across researches is an important step forward in mathematical modeling, as has been recently argued \cite{mestdagh2018prepaid}.
+This pooling of computational resources across researches is an important step forward in mathematical modeling \cite{mestdagh2018prepaid}.
 
@@
 Across the examples, we introduce multiple tools to validate the performance of our method.
-The outline of the remaining manuscript is as follows: The \textbf{Methods} section introduces the main building blocks of the new method and summarizes the main steps as pseudocode.
+The outline of the remaining manuscript is as follows: The \textbf{Methods} section introduces the main building blocks of the new method and summarizes the main steps in pseudocode.
 The \textbf{Results} section presents the various applications of the model to real-world research domains.

### 2019-07-08 latex/paper_main.tex
Modify text

@@
 \begin{abstract}
-Parametric models of complex processes are ubiquitous throughout the sciences.
+Mathematical models of complex processes are ubiquitous throughout the sciences.
 As the processes under study and the models describing them become increasingly complex, parameter estimation with standard Bayesian and frequentist methods can quickly become intractable.
 To address this, we propose a novel method for likelihood-free inference based on invertible neural networks.
-The method is capable of performing fast full Bayesian inference on large amounts of data by training the networks on simulated data and learning to invert the model under study.
-The method is independent of particular data formats, as it includes a summary network trained to embed the observed data into fixed-size vectors in a data-driven way.
+The method is capable of performing fast and full Bayesian inference on large datasets by training the networks on simulated data and learning a probabilistic mapping between parameters and data.
+The method is independent of any particular data representation, as it incorporates a summary network trained to embed the observed data into fixed-size vectors in a data-driven way.
 This makes the method applicable to various scenarios where standard inference techniques fail.
@@
 
-Once a mathematical model has been formulated, the next step consists in fitting the model to experimental or observational data and recovering the parameters of interest.
+Once a mathematical model has been formulated, the next step consists in fitting the model to experimental or observational data to recover the parameters of interest.
 However, estimating the parameters of a mathematical model from data can quickly become one of the most tenacious challenges in applications to real-world problems.
-It is also one of the most important ones to be tackled, since it is impossible to test the utility of a model without reliable parameter estimation methods, regardless of the model's sophistication or theoretical appeal.
+Parameter estimation is also one of the most important ones to be tackled, since, otherwise, it is impossible to test the utility of a model, regardless of the model's sophistication or theoretical appeal.
 Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
@@
 
-However, as mathematical models and processes under description become increasingly complex, parameter estimation and model selection can quickly become intractable with standard Bayesian and frequentist method.
+However, as mathematical models and processes under description become increasingly complex, parameter estimation can quickly become prohibitively slow or even intractable with standard Bayesian and frequentist method.
 Complex models specified by a generative stochastic mechanism do not always provide a closed-form solution for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}.
@@
 
-In an attempt to balance the shortcomings of traditional methods, ideas from machine learning and deep learning research have recently entered the field of likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
+In an attempt to balance the shortcomings of traditional methods, ideas from machine learning and deep learning research have recently entered the field of likelihood-free inference \cite{mertens2019deep, radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
 The most common approach has been to cast the problem of parameter estimation as a supervised learning task.
@@
 Furthermore, flow-based methods are capable of approximating very high-dimensional distributions (e.g., the pixels of an image).
-Once trained with a sufficient amount of simulated data, our invertible network can perform full Bayesian inference on large amounts of real data from a given domain in a single pass.
+Once trained with a sufficient amount of simulated data, the invertible network can perform rapid Bayesian inference on large datasets from a given research domain.
 \begin{figure}
@@
 
-The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated statistical problems in various scientific domains (\autoref{fig:Fig.1}).
+The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated mathematical models in various scientific domains (\autoref{fig:Fig.1}).
 Moreover, the method addresses many of the limitations of previous likelihood-free methods.
-First, it involves no costly MCMC or rejection sampling, which makes the inference phase lightning fast, as we also make use of GPU-accelerated computation.
-Second, it involves no fixed summary statistics or kernels, but learns the most informative representation of the data in an end-to-end manner.
+First, it involves no costly MCMC or rejection sampling, which makes the inference phase lightning fast.
+Second, it involves no fixed summary statistics, but instead learns the most informative representation of the data in an end-to-end manner.
 Third, the method is fully Bayesian, as it directly learns the posterior over model parameters and thus allows for the quantification of uncertainty, which is a crucial requirement in parameter estimation \cite{kendall2017uncertainties, gelman2013bayesian}.
 Last, the trained networks can be shared and reused by multiple researches within a scientific domain, thus removing the need for wasteful computations and fitting a separate model for each and every dataset.
-This pooling of computational resources across researches is an important step forward in mathematical modeling \cite{mestdagh2018prepaid}.
+This pooling of computational resources across researchers is an important step forward in mathematical modeling \cite{mestdagh2018prepaid}.
 
@@
 The parameters of a mathematical model are represented as a vector $\boldsymbol{\theta} = (\theta_{1}, \theta_{2},...,\theta_{d})$, and all trainable parameters of the invertible and summary neural networks as $\boldsymbol{\phi} = (\boldsymbol{\phi}_{inv}, \boldsymbol{\phi}_{sum})$.
-The number of parameters of a mathematical model will be denoted as $d$, and the number of simulated data points as $n$.
+The number of parameters of a mathematical model will be denoted as $d$, and the number of data points in $\boldsymbol{x}$ or $\boldsymbol{X}$ as $n$.
 
@@
 Our method draws on major advances in modern deep probabilistic modeling, also referred to as deep generative modeling \cite{bloem2019probabilistic, kingma2018glow, ardizzone2018analyzing, kingma2014auto}.
-A hallmark idea in deep probabilistic modeling is to handle intractable target probability distributions by sampling from simpler distributions (e.g., Gaussian or uniform distributions) and transforming these samples via a complex non-linear, learnable transformations.
+A hallmark idea in deep probabilistic modeling is to handle intractable target probability distributions by sampling from simpler distributions (e.g., Gaussian or uniform distributions) and transforming these samples via complex non-linear, learnable transformations.
 Most popular deep probabilistic models entail two phases.
@@
 Fixed permutation matrices are used before each ACB to ensure that each axis of the latent space encodes information from all components of $\boldsymbol{\theta}$.
-In order to take into account the observed data $\boldsymbol{x}$, each of the internal networks of each ACB is augmented to take $\boldsymbol{x}$ as an additional input - $s_{1}(\cdot,\boldsymbol{x}), s_{2}(\cdot,\boldsymbol{x}), t_{1}(\cdot,\boldsymbol{x}), t_{2}(\cdot,\boldsymbol{x})$ - so a complete pass through the entire invertible chain can be expressed as:
+In order to take into account the observed or simulated data $\boldsymbol{x}$, each of the internal networks of each ACB is augmented to take $\boldsymbol{x}$ as an additional input - $s_{1}(\cdot,\boldsymbol{x}), s_{2}(\cdot,\boldsymbol{x}), t_{1}(\cdot,\boldsymbol{x}), t_{2}(\cdot,\boldsymbol{x})$ - so a complete pass through the entire invertible chain can be expressed as:
 \begin{align*} 
@@
 where $\partial f/\partial \boldsymbol{\theta} = \boldsymbol{J}_{f}$ is the Jacobian of the learned transformation $f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi})$ with respect to the input.
-Both terms in Eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the log determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
+Both terms in Eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
 \ref{eqn:2} and \ref{eqn:3}.
@@
 We train all cINNs and summary networks described in this paper jointly via backpropagation.
-For all following experiments, we use the Adam optimizer  with a starter learning rate of $10^-3$ and an exponential decay rate of $.95$.
+For all following experiments, we use the Adam optimizer  with a starter learning rate of $10^{-3}$ and an exponential decay rate of $.95$.
 We set the weight regularization parameter $\tau$ to a value of $10^{-5}$.
-We perform $50000$ to $100000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
+We perform $50\ 000$ to $100\ 000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
 Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINN).
-All networks were trained on a single-GPU machine equipped with NVIDIA GTX1060 graphics card.
+All networks were trained on a single-GPU machine equipped with NVIDIA\textsuperscript{\textregistered} GTX1060 graphics card.
 Regarding the data generation step, we use two training approaches.
 
-The first approach follows the classical approximate Bayesian computation approach to create a large “reference table” or grid of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$.
+The first approach follows the classical approximate Bayesian computation approach to create a large \textit{reference table} or grid of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$.
 The reference table is then used as training data for the neural network and training continues for a pre-specified number of epochs through the entire reference table.
 A separate validation dataset is eventually used to assess the performance of the network.
-This training approach separates the simulation from the training phase but can incur large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
+Training on large pre-simulated datasets separates the simulation from the training phase but can incur large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
 
@@
 Correspondingly, a dataset, or a batch of datasets, is created on the fly and then passed through the neural network.
-This training regime has the advantage that the network never “experiences” the same input data twice.
+This training regime has the advantage that the network never \textit{experiences} the same input data twice.
 Moreover, training can continue as long as the network keeps improving (i.e., the loss keeps decreasing), since overfitting in the classical sense is nearly impossible.
@@
 \State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = q(\boldsymbol{\theta}^{(i)})$ for $i=1,...,m$}
-\State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$}
+\State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ for $i=1,...,m$}
 \State {Pass $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ and $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ through cINN $f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi}_{inv})$ to obtain $\{\boldsymbol{z}^{(i)}\}_{i=1}^{m}$}
@@
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $500000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $50\%$ of the counts are $0$).
+We simulate $500\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $50\%$ of the counts are $0$).
 The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
@@
 In the current work, we proposed and explored a novel end-to-end likelihood-free method which uses invertible neural networks to perform approximate Bayesian inference on any mathematical process model.
-We demonstrated the utility of the method by applying it to models from various scientific domains exhibiting various data formats and data-generating mechanisms.
+We demonstrated the utility of the method by applying it to models from various scientific domains dealing with several different data formats and data-generating mechanisms.
 Further, we explored two possible training approaches suitable for different simulation scenarios, namely an active learning approach, and a reference-table approach.
@@
 Our method combines the universal approximation power of deep learning methods with the important uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
-Besides being capable of performing rapid Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable “parameter estimation machines” for various research domains.
+Besides being capable of performing rapid Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable \textit{parameter estimation machines} for various research domains.
 Moreover, the method is not confined solely to inference on intractable models, but can also prove as a viable alternative in modeling contexts where standard Bayesian or frequentist inference methods are available, but inference is nevertheless prohibitively slow.
 
-Inspired by previous machine learning approaches to likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, our method shares many of the advantages of these methods and further overcomes some important limitations.
+Inspired by previous machine learning approaches to likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, our method shares many of the advantages of these methods and further overcomes some of their important limitations.
 
@@
 
-Second, we showed that the ML loss optimizes directly the posterior over model parameters of interest, which is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
+Second, we showed that the ML loss directly maximizes the posterior over model parameters of interest, which is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
 Thus, inference is exact when the ML is minimized \cite{kingma2018glow, dinh2016density}.
@@
 We also showed that the recovery becomes better with increasing number of observed data points, while the variance of the estimates becomes larger.
-This is an important and highly desirable property of any parameter estimation method, as it mirrors the increase oin information following increasing number of data points.
+This is an important and highly desirable property of any parameter estimation method, as it mirrors the increase in information following increasing number of data points.
 
@@
 In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
-Note, however, that our method does not need to store training data on disk, since the "knowledge" about the relationship between data and parameters is compressed into the networks' weights.
+Note, however, that our method does not need to store training data on disk, since the \textit{knowledge} about the relationship between data and parameters is compressed into the networks' weights.
+This not only makes the global sharing of pre-trained parameter estimation networks across researchers extremely easy, but also makes their local storage very efficient.
 Moreover, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern GPUs.
 
-These advantages notwithstanding, some limitations of the method deserve a mention.
+These advantages notwithstanding, some limitations of the method deserve mention.
 Even though high-level deep learning libraries, such as \textit{TensorFlow} or \textit{Torch}, allow for rapid and relatively straightforward development of various neural network architectures, the implementational burden associated with the current method is still reasonably high.
@@
 Another potential shortcoming of the method is the seemingly overwhelming number of hyperparameters that might require fine-tuning by the user for optimal performance on a given task.
-However, we observe that many of the default hyperparameter values are sufficient to achieve excellent performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training, even if the model to be learned is relatively simple.
+However, we observe that many of the default hyperparameter values are sufficient to achieve excellent performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training, even if the mathematical model to be learned is relatively simple.
 We expect that a single architecture should be able to perform well on almost all models from a given domain (i.e., a single architecture for decision-making models).
@@
 As formal theories in various scientific disciplines (especially in the younger sciences, such as, neuroscience, cognitive science, computational biology, etc.) become increasingly complex, the need for powerful and universally applicable likelihood-free estimation methods becomes increasingly pressing.
-In the present work, we addressed this need by introducing a method potentially applicable to \textit{any} modeling scenario in \textit{any} research domain where simulations from the process model can be obtained.
+In the present work, we addressed this need by introducing a method potentially applicable to \textit{any} modeling scenario in \textit{any} research domain where simulations from the mechanistic model can be obtained.
 We hope that the new method will enable researchers from a variety of fields to accelerate model-based inference and will further prove its utility beyond the examples considered in this paper.

### 2019-07-10 latex/paper_main.tex
Add paper after new comments

@@
 As the processes under study and the models describing them become increasingly complex, parameter estimation with standard Bayesian and frequentist methods can quickly become intractable.
-To address this, we propose a novel method for likelihood-free inference based on invertible neural networks.
-The method is capable of performing fast and full Bayesian inference on large datasets by training the networks on simulated data and learning a probabilistic mapping between parameters and data.
+To address this issue, we propose a novel method for likelihood-free inference based on invertible neural networks.
+The method is capable of performing full Bayesian inference on large datasets by training the networks on simulated data to learn a probabilistic mapping between parameters and data.
 The method is independent of any particular data representation, as it incorporates a summary network trained to embed the observed data into fixed-size vectors in a data-driven way.
@@
 We demonstrate the utility of the method on a toy model with known analytic posterior and on example models from population dynamics, epidemiology, cognitive science and genetics.
-We argue for a general framework for building reusable parameter estimation machines for potentially any process model from which simulations can be obtained.
+We argue that our method provides a general framework for building reusable parameter estimation machines for any process model from which data can be simulated.
 \end{abstract}
@@
 The functional form of $q$, which can represent an arbitrarily complex process by an arbitrarily complicated expression, is usually guided by a well-founded theoretical framework.
-For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm dictating the rate of gene expression in certain cells.
+For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm controlling the rate of gene expression in certain cells.
 Thus, it is only within the context of a theory that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
-Examples of mathematical models can be found in various scientific domains, e.g., genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical, geritz2004mechanistic}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, just to name a few.
+Examples of mathematical models can be found in various scientific domains, for instance, genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical, geritz2004mechanistic}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, to name just a few.
 
-Once a mathematical model has been formulated, the next step consists in fitting the model to experimental or observational data to recover the parameters of interest.
-However, estimating the parameters of a mathematical model from data can quickly become one of the most tenacious challenges in applications to real-world problems.
-Parameter estimation is also one of the most important ones to be tackled, since, otherwise, it is impossible to test the utility of a model, regardless of the model's sophistication or theoretical appeal.
+Once a theoretical model has been translated to an exact mathematical formulation, the model can be fitted to observed data to recover the parameter values describing the data-generating process.
+However, estimating the parameters of a mathematical model from data can quickly become the most tenacious challenge a researcher faces.
+At the same time, parameter estimation is the most important, since, otherwise, it is impossible to test the utility of a model, regardless of the model's sophistication or theoretical appeal.
 Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
@@
 The computation of summary statistics is a crucial aspect in likelihood-free inference.
-Previous approaches mainly use hand-crafted summary statistics tailored to the specific application.
-However, in many application, it is not straightforward to settle upon a set of good summary statistics.
-The summary network completely eliminates the need to manually decide on and compute a fixed number of summary statistics and, thus, makes the method independent of the format or the size of the data.
+Previous approaches typically used hand-crafted summary statistics tailored to the specific application \cite{raynal2018abc, jiang2017learning, sunnaaker2013approximate}.
+However, in many applications, finding an optimal set of summary statistics is far from obvious.
+The summary network eliminates the need to manually decide on and compute a fixed set of summary statistics and, thus, makes the method independent of the format or the size of the data.
 
@@
 In the \textbf{Methods} section, we show that our method maximizes the posterior over model parameters directly when cast in the context of likelihood-free inference.
-Furthermore, flow-based methods are capable of approximating very high-dimensional distributions (e.g., the pixels of an image).
+Furthermore, flow-based methods are capable of approximating high-dimensional distributions (e.g., the pixels of an image).
 Once trained with a sufficient amount of simulated data, the invertible network can perform rapid Bayesian inference on large datasets from a given research domain.
@@
 Moreover, the method addresses many of the limitations of previous likelihood-free methods.
-First, it involves no costly MCMC or rejection sampling, which makes the inference phase lightning fast.
+First, it involves no costly MCMC or rejection sampling, which makes inference lightning fast, once the networks have been trained.
 Second, it involves no fixed summary statistics, but instead learns the most informative representation of the data in an end-to-end manner.
@@
 The parameters of a mathematical model are represented as a vector $\boldsymbol{\theta} = (\theta_{1}, \theta_{2},...,\theta_{d})$, and all trainable parameters of the invertible and summary neural networks as $\boldsymbol{\phi} = (\boldsymbol{\phi}_{inv}, \boldsymbol{\phi}_{sum})$.
-The number of parameters of a mathematical model will be denoted as $d$, and the number of data points in $\boldsymbol{x}$ or $\boldsymbol{X}$ as $n$.
+The number of parameters of a mathematical model will be denoted as $d$, and the number of observations in $\boldsymbol{x}$ or $\boldsymbol{X}$ as $n$.
 
@@
 Our method draws on major advances in modern deep probabilistic modeling, also referred to as deep generative modeling \cite{bloem2019probabilistic, kingma2018glow, ardizzone2018analyzing, kingma2014auto}.
-A hallmark idea in deep probabilistic modeling is to handle intractable target probability distributions by sampling from simpler distributions (e.g., Gaussian or uniform distributions) and transforming these samples via complex non-linear, learnable transformations.
+A hallmark idea in deep probabilistic modeling is to handle intractable target probability distributions by sampling from simpler distributions (e.g., Gaussian or uniform distributions) and transforming these samples via learned complex non-linear transformations.
 Most popular deep probabilistic models entail two phases.
@@
 During the \textit{inference phase}, samples from the target distribution are obtained by sampling from the simple distribution and applying the transformation learned during the training phase (see \autoref{fig:Fig.1b}).
-Using this approach, recent applications of deep probabilistic models have achieved unprecedented results on extremely high-dimensional and intractable problems (e.g., complex data distributions such as natural images, music, or text).
+Using this approach, recent applications of deep probabilistic models have achieved unprecedented results on extremely high-dimensional and intractable problems (e.g., complex data distributions such as images, music, or text)  \cite{bloem2019probabilistic, kingma2018glow, grover2018flow}.
 
-In the context of mathematical modeling and Bayesian inference, the target distribution is the posterior distribution of model parameters $p(\boldsymbol{\theta}|\boldsymbol{x})$ capturing our uncertainty about the numerical values of parameters given empirical data.
+In the context of mathematical modeling and Bayesian inference, the target distribution is the posterior distribution of model parameters $p(\boldsymbol{\theta}|\boldsymbol{x})$ capturing the uncertainty about the numerical values of parameters given empirical data.
 We can leverage the fact that most mathematical models are generative in nature and as such can be used to perform multiple simulations of the process of interest.
 By specifying a prior distribution over the model parameters $p(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters.
-Thus, at inference time, one can condition the model on observed data $\boldsymbol{x}_{obs}$ and obtain samples $\boldsymbol{\theta}_{1}, \boldsymbol{\theta}_{2},...,\boldsymbol{\theta}_{L}$  approximating the posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$ in the manner described above.
+Thus, during the inference phase, the model generates samples $\boldsymbol{\theta}_{1}, \boldsymbol{\theta}_{2},...,\boldsymbol{\theta}_{L}$ from the observed data $\boldsymbol{x}_{obs}$, approximating the target posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$.
 
@@
 Ideally, we want to avoid hand-crafted summary statistics, and instead learn the most informative summary statistics directly from data.
-Therefore, instead of feeding the raw simulated (observed) data to each ACB, we pass the data through an additional summary network to obtain a fixed-sized vector of learned summary statistics $\tilde{\boldsymbol{x}} = h(\boldsymbol{x};\boldsymbol{\phi}_{sum})$ and learn the parameters of the summary network $h$ jointly with those of the cINN chain via backpropagation.
+Therefore, instead of feeding the raw simulated or observed data to each ACB, we pass the data through an additional summary network to obtain a fixed-sized vector of learned summary statistics $\tilde{\boldsymbol{x}} = h(\boldsymbol{x};\boldsymbol{\phi}_{sum})$ and learn the parameters of the summary network $h$ jointly with those of the cINN chain via backpropagation.
 Thus, the current method remains completely end-to-end and is capable of generalizing to data sets of variable input size and structure.
@@
 We set the weight regularization parameter $\tau$ to a value of $10^{-5}$.
-We perform $50\ 000$ to $100\ 000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
+With these settings, we perform $50\ 000$ to $100\ 000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
 Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINN).
-All networks were trained on a single-GPU machine equipped with NVIDIA\textsuperscript{\textregistered} GTX1060 graphics card.
+All networks were implemented in Python using the \textit{TensorFlow} library \cite{abadi2016tensorflow} and trained on a single-GPU machine equipped with NVIDIA\textsuperscript{\textregistered} GTX1060 graphics card.
 Regarding the data generation step, we use two training approaches.
@@
 A separate validation dataset is eventually used to assess the performance of the network.
-Training on large pre-simulated datasets separates the simulation from the training phase but can incur large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
+Training on large pre-simulated datasets separates the simulation from the training phase but can cause large memory overhead, as the reference table must be stored on disk and then loaded in chunks or in its entirety into memory.
 
-The second approach follows a different strategy, which is used mainly in the field of active learning.
+The second approach follows a different strategy, which resembles ideas from \textit{active learning} \cite{mnih2015human}.
 Correspondingly, a dataset, or a batch of datasets, is created on the fly and then passed through the neural network.
@@
 Moreover, training can continue as long as the network keeps improving (i.e., the loss keeps decreasing), since overfitting in the classical sense is nearly impossible.
-However, if the simulations are computationally expensive and researchers need to experiment with different networks or training hyperparameters, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
+However, if simulations are computationally expensive and researchers need to experiment with different networks or training hyperparameters, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
 
@@
 On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\phi}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
-In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the life sciences, since they tend to be unimodal and relatively simple (in contrast to the distribution of natural images or words in a spoken language).
+In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the cognitive or life sciences, since they tend to be unimodal and relatively simple (in contrast to the distributions required to represent images or words in a spoken language).
 From these three ingredients, a universal and reusable sampler can be designed for likelihood-free Bayesian estimation of both tractable and intractable mathematical models.
@@
 \State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = q(\boldsymbol{\theta}^{(i)})$ for $i=1,...,m$}
-\State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ for $i=1,...,m$}
+\State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$}
 \State {Pass $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ and $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$ through cINN $f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi}_{inv})$ to obtain $\{\boldsymbol{z}^{(i)}\}_{i=1}^{m}$}
@@
 
-The backpropagation algorithm works by computing the gradients of the loss function w.r.t.
-the parameters of the neural networks and then adjusting the parameters, so as to drive the loss function to a local minimum.
+The backpropagation algorithm works by computing the gradients of the loss function with respect to the parameters of the neural networks and then adjusting the parameters, so as to drive the loss function to a local minimum.
 We experienced no instability or convergence issues during training with the ML loss.
@@
 In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of challenging models from population dynamics, cognitive science, epidemiology, and genetics.
-Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.All examples are implemented in Python using the \textit{TensorFlow} library.
+Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
@@
 
-We consider a number of metrics for determining the performance of our method.
-To assess the goodness of parameter recovery, we compute the normalized root mean squared error (NRMSE) and the coefficient of determination ($R^{2}$) metrics of the parameter estimates.
+To assess the performance of our method on the following application examples, we consider a number of different metrics.
+To assess the precision of point estimates (posterior means), we compute the normalized root mean squared error (NRMSE) and the coefficient of determination ($R^{2}$) between estimated and true parameter values.
 To assess the recovery of the full posterior, we compute the Kullback-Leibler (KL) divergence \cite{hershey2007approximating} between the true and the approximate distributions for the toy example, and use simulation-based calibration (SBC, \cite{talts2018validating}) for the other examples where the analytic posterior is not available in closed-form.
@@
 As a proof-of-concept, we demonstrate the utility of our method in recovering the true analytic posteriors of the regression coefficients of a conjugate Bayesian regression model.
-To set the setting, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$.
-We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, i.e., $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
+To set the stage, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$.
+We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, that is, $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
 We place a $d$-dimensional diagonal Gaussian prior on the regression coefficients centered at $0$: $\boldsymbol{\theta} \sim \mathcal{N}_{d}(\boldsymbol{0},b^{-1}\boldsymbol{I})$ where $b$ is the precision of the prior ($b \equiv 1/\sigma_{\boldsymbol{\theta}}^{2}$).
@@
 \caption[short]{Results on the Bayesian toy regression example.
-\textbf{(a)} Analytic vs.
-estimated posterior means on the validation sample.
+\textbf{(a)} Pair-plots of the analytic vs.
+estimated posterior means on the whole validation sample.
 We observe a near-perfect overlap between analytic and estimated posterior means and no spurious covariances; \textbf{(b)} Some example draws from the estimated posteriors.
@@
 
-What if the data does not contain information about a given parameter?
+What if the data does not contain information about a particular parameter?
 In this case, any good estimation method should detect this, and return the prior of the particular parameter.
-To test this, we add a random uniform variable $u \sim \mathcal{U}(0, 1)$ to the parameter vector $\boldsymbol{\theta}$ and train the model with this additional dummy variable.
-Then, we inspect the posterior of $u$ for uniformity.
+To test this, we append a random uniform variable $u \sim \mathcal{U}(0, 1)$ to the parameter vector $\boldsymbol{\theta}$ and train the model with this additional dummy parameter.
+We expect that the networks learn to ignore this dummy parameter, that is, the estimated posterior of $u$ is (approximately) equal to the uniform prior.
 
-The results on the Ricker model are depicted in \autoref{fig:Fig.3}.
-As evident from the graphics, the parameter recovery becomes better when more time points with data are available (\autoref{fig:Fig.4c}).
+The results on the Ricker model are depicted in \autoref{fig:Fig.4}.
+As evident from the graphs, parameter recovery becomes better when more time points with data are available (\autoref{fig:Fig.4c}).
 At $T=500$, the NRMSEs range between $0.015$ and $0.063$, and the $R^{2}$ metrics between $0.997$ and $0.952$, indicating very good recovery of the posterior means (\autoref{fig:Fig.4a}).
-The parameter $\sigma$ seems to be hardest to recover.
+The parameter $\sigma$ seems to be the hardest to recover.
 Inspecting the full posteriors, we further see that the posterior distribution of the dummy noise variable $u$ closely resembles the prior, as expected due to the complete lack of mutual information between the data $x$ and $u$ (\autoref{fig:Fig.4b}).
@@
 \subsection*{Example 2 - The Lévy-Flight Model}
-Evidence accumulator models (EAMs) describe (perceptual) decision making by a set of neurocognitively motivated parameters \cite{ratcliff2008diffusion}.
-EAMs are most often applied to choice reaction times (RT) data to obtain an estimate of the neurocognitive processes governing observed RT distributions in human and animal participants.
-Most EAM variants share four underlying assumptions: \textit{i)} information about a stimulus (response option) is accumulated continuously through time; \textit{ii)} stochasticity in the form of noisy accumulation ensures variability; \textit{iii)} empirical response times can be decomposed into a decision time component and a non-decision time component accounting for pre-decisional perceptual (encoding time) and post-decisional motor processes (response execution); and \textit{iv)} a decision is met when the activation of an accumulator exceeds a threshold.
+Evidence accumulator models (EAMs) describe human decision making by a set of neurocognitively motivated parameters \cite{ratcliff2008diffusion}.
+EAMs are most often applied to choice reaction times (RT) data to obtain an estimate of the neurocognitive processes governing observed RT distributions in human (or animal) participants.
+Most EAM variants share four underlying assumptions: \textit{i)} information about a stimulus (response option) is accumulated continuously through time; \textit{ii)} stochasticity in the form of noisy accumulation ensures variability; \textit{iii)} empirical response times can be decomposed into a decision time component and a non-decision time component accounting for pre-decisional perceptual (encoding time) and post-decisional motor processes (response execution); and \textit{iv)} a decision is met when the activation of an accumulator reaches a threshold.
 In its most general formulation, the forward model of EAMs takes the form of a stochastic differential equation given by \cite{usher2001time}:
@@
 \end{align*}
-where $dx$ denotes a change in activation of an accumulator, $v$ denotes the average speed of information processing (often termed the drift rate), and $d\xi$ represents a stochastic additive component, usually modeled as following a Gaussian distribution centered around $0$: $d\xi \sim \mathcal{N}(0, c^{2})$.
+where $dx$ denotes a change in activation of an accumulator, $v$ denotes the average speed of information accumulation (often termed the drift rate), and $d\xi$ represents a stochastic additive component, usually modeled as following a Gaussian distribution centered around $0$: $d\xi \sim \mathcal{N}(0, c^{2})$.
 
-EAMs are particularly amenable for likelihood-free inference, since most members of the family turn out to be intractable \cite{miletic2017parameter}.
+EAMs are particularly amenable for likelihood-free inference, since the likelihood of some members of this model family turn out to be intractable \cite{miletic2017parameter}.
 This intractability has precluded many interesting applications and empirically driven model refinements.
@@
 The LFM assumes an \textit{alpha-stable} noise distribution of the evidence accumulation process in order to model "jumps" in the decision process.
-However, the inclusion of \textit{alpha-stable} noise leads to a model with intractable likelihood; further, to our knowledge, a fully Bayesian treatment of the model is still missing from the literature.
+However, the inclusion of \textit{alpha-stable} noise leads to a model with intractable likelihood; further, to our knowledge, a fully Bayesian treatment of the model is still missing in the literature.
 The forward equation of the LFM is given by:
@@
 \end{align*}
-The LFM has three additional parameters: a threshold $a$ determining the amount of evidence needed for the termination of a decision process; a relative starting point, $zr$, determining the amount of starting evidence available to the accumulator before the actual decision alternatives are presented; and an additive non-decision time $t_{0}$.
+The LFM has three additional parameters: the threshold $a$ determining the amount of evidence needed for the termination of a decision process; a relative starting point, $zr$, determining the amount of starting evidence available to the accumulator before the actual decision alternatives are presented; and an additive non-decision time $t_{0}$.
 
-During training of the networks, we simulate RT data from two experimental conditions with two different drift rates (see \textbf{SI} for details of the simulation).
+During training of the networks, we simulate response times data from two experimental conditions with two different drift rates (see \textbf{SI} for details of the simulation), since this such a design is often encountered in psychological research.
 The parameter estimation task is thus to recover the parameters $\boldsymbol{\theta} = (v_{0}, v_{1}, a, t_{0}, zr, \alpha)$ from two-dimensional \textit{i.i.d.} RT data $\boldsymbol{X} = (\boldsymbol{x}_{1},\boldsymbol{x}_{2},...,\boldsymbol{x}_{n})$ containing variable number or RT trials.
@@
 
-The results on the LFM model are depicted in \autoref{fig:Fig.4}.
+The results on the LFM model are depicted in \autoref{fig:Fig.5}.
 To our knowledge, this is the first Bayesian treatment of the LFM, as its intractability makes traditional methods prohibitively slow.
-We observe state-of-the art recovery of all LFM parameters with NRMSEs ranging between $0.008$ and $0.048$ and $R^{2}$ between $0.972$ and $0.999$.
+We observe excellent recovery of all LFM parameters with NRMSEs ranging between $0.008$ and $0.048$ and $R^{2}$ between $0.972$ and $0.999$.
 Further, estimation remains very good across all trial sizes, with goodness increasing as more trials become available.
@@
 The number of individuals moving from $S$ to $I$, given by $\triangle N_{SI}$, and the number of people moving from $I$ to $R$, given by $\triangle N_{IR}$, over a time interval $\triangle t$ are modeled as binomial random variables.
-The above listed stochastic system has no known analytic solution and thus requires numerical simulation methods for finding optimal parameter values.
+The above listed stochastic system has no known analytic solution and thus requires numerical simulation methods for recovering parameter values from data.
 Cast as a parameter estimation task, the challenge is to recover $\boldsymbol{\theta} = \{\beta,\gamma\}$ from three dimensional time-series data $\boldsymbol{X} = (\boldsymbol{x}_{1},\boldsymbol{x}_{2},...,\boldsymbol{x}_{T})$ where each $\boldsymbol{x}_{t} \in \mathbb{N}^{3}$ is a triple containing the number of susceptible ($S$), number of infected ($I$), and recovered ($R$) individuals at time $t$.
@@
 The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(200, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
-Usually, at lower $T$s, the system has not converged to an equilibrium (i.e., not all individuals have moved from being $I$ to $R$).
+Usually, at lower $T$s, the system has not converged to an equilibrium (i.e., not all individuals have transitioned from being $I$ to $R$).
 Thus, it is especially interesting to see if our method can recover the parameters, even if the process dynamics are still unfolding over time.
@@
 Thus, inference is exact when the ML is minimized \cite{kingma2018glow, dinh2016density}.
-This \textit{optimal performance} claim is confirmed by our toy Bayesian regression example, in which we observe negligibly small deviations of the approximate from the true posterior.
+This \textit{optimal performance} claim is confirmed by our toy Bayesian regression example, in which we observe negligibly small deviations from the form of the true posterior.
 Further, researchers are often interested in some summary of the posterior, for instance, the posterior mean or the posterior variance \cite{radev2019towards, raynal2018abc}.
 We demonstrated that our method exhibits excellent recovery of the posterior means throughout the examples.
-We also showed that the recovery becomes better with increasing number of observed data points, while the variance of the estimates becomes larger.
-This is an important and highly desirable property of any parameter estimation method, as it mirrors the increase in information following increasing number of data points.
+We also showed that the recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters becomes smaller.
+These are important and highly desirable properties of any parameter estimation method, as it mirrors the increase in information following increasing number of data points.
 
@@
 In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
-Note, however, that our method does not need to store training data on disk, since the \textit{knowledge} about the relationship between data and parameters is compressed into the networks' weights.
+Note, however, that our method does not need to store training data, since the \textit{knowledge} about the relationship between data and parameters is compressed into the networks' weights.
 This not only makes the global sharing of pre-trained parameter estimation networks across researchers extremely easy, but also makes their local storage very efficient.

### 2019-07-10 process_models.py
Measure speed

@@ -150,6 +150,7 @@ def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gam
         return tf.convert_to_tensor(X, dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)
     return X, theta
 
+
 def plot_sir(beta, gamma, n_points=500, figsize=(8, 4), N=1000, filename=None):
     """
     Simulates a single SIR process.
@@ -170,6 +171,7 @@ def plot_sir(beta, gamma, n_points=500, figsize=(8, 4), N=1000, filename=None):
     ax.set_xlabel(r'Number of time points ($T$)', fontsize=12)
     ax.set_ylabel('Number of individuals', fontsize=12)
     ax.legend(fontsize=10)
+    f.tight_layout()
 
     # Save if specified
     if filename is not None:

### 2019-07-11 latex/paper_main.tex
Measure speed
New edit

@@
 
-\dropcap{M}athematical models are formal descriptions of scientific theories allowing a clear and unambiguous way to formulate and test scientific hypotheses about deterministic or probabilistic phenomena.
+\dropcap{M}athematical models are formal descriptions of scientific theories.
 In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a forward model $q$ mimicking the process by which manifest data $\boldsymbol{x}$ arise from latent parameters:
@@
 
-Once a theoretical model has been translated to an exact mathematical formulation, the model can be fitted to observed data to recover the parameter values describing the data-generating process.
-However, estimating the parameters of a mathematical model from data can quickly become the most tenacious challenge a researcher faces.
-At the same time, parameter estimation is the most important, since, otherwise, it is impossible to test the utility of a model, regardless of the model's sophistication or theoretical appeal.
+An important goal of mathematical modeling is recovering the parameters of the model from observed data.
 Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
-However, due to noise, inherent stochasticity or loss of information, the inverse usually does not exist, so researchers need to resort to the sophisticated frameworks of Bayesian or frequentist inference.
+Unfortunately, this inverse rarely exists, and estimating parameters remains computationally inconvenient in a wide range of settings.
 
-However, as mathematical models and processes under description become increasingly complex, parameter estimation can quickly become prohibitively slow or even intractable with standard Bayesian and frequentist method.
-Complex models specified by a generative stochastic mechanism do not always provide a closed-form solution for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}.
+Parameter estimation becomes notably difficult when the forward model in Eq.\ref{eqn:1} does not provide a closed-form for the \textit{likelihood function} \cite{palestro2018likelihood, csillery2010approximate, toni2009simulation}.
 This poses great difficulties for Bayesian and frequentist methods alike, since both depend on the numerical evaluation of a likelihood function as a proxy for assessing model fit to data.
@@
 In this case, enforcing simplifying distributional assumptions (i.e., independence or Gaussian assumptions) can increase computational speed, but can also lead to model misspecifications and dramatically incorrect estimates.
-Therefore, the need for powerful and reliable likelihood-free estimation methods becomes obvious.
+Therefore, there is a need for accurate and computationally convenient likelihood-free estimation methods.
 
@@
 A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate, csillery2010approximate}.
-ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the generative model $q(\boldsymbol{\theta})$ using the sampled parameters.
+ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the forward model $q(\boldsymbol{\theta})$ using the sampled parameters.
 Correspondingly, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
@@
 
-In an attempt to balance the shortcomings of traditional methods, ideas from machine learning and deep learning research have recently entered the field of likelihood-free inference \cite{mertens2019deep, radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
+To address the shortcomings of traditional methods, ideas from machine learning and deep learning research have recently entered the field of likelihood-free inference \cite{mertens2019deep, radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}.
 The most common approach has been to cast the problem of parameter estimation as a supervised learning task.
@@
 Inspired by previous machine learning approaches \cite{hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, the current work proposes a novel and universal likelihood-free method capable of performing full Bayesian inference on any mathematical process model from which simulations can be obtained.
-It treats parameter inference as a task of inverting a generative model and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
+It treats parameter inference as a task of inverting the forward model in Eq.\ref{eqn:1} and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
 The method integrates two separate deep neural networks modules (detailed in the \textbf{Methods} section; see \autoref{fig:Fig.1}) trained jointly on simulated data: a \textit{summary network} and an \textit{invertible network}.
@@
 It should be designed to follow the functional and probabilistic symmetries inherent in the data, e.g.
-a permutationally invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network for time-series data, or a convolutional network for grid-like data \cite{goodfellow2016deep}.
+a permutationally invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network for time-series data \cite{gers1999learning}, or a convolutional network for grid-like data \cite{goodfellow2016deep}.
 The computation of summary statistics is a crucial aspect in likelihood-free inference.
@@
 
-The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of inverting complicated mathematical models in various scientific domains (\autoref{fig:Fig.1}).
+The joint training of a summary network and an invertible network results in a powerful and universal parameter estimation machine capable of estimating complex mathematical models in various scientific domains (\autoref{fig:Fig.1}).
 Moreover, the method addresses many of the limitations of previous likelihood-free methods.
@@
 
-In the context of mathematical modeling and Bayesian inference, the target distribution is the posterior distribution of model parameters $p(\boldsymbol{\theta}|\boldsymbol{x})$ capturing the uncertainty about the numerical values of parameters given empirical data.
+In the context of mathematical modeling and Bayesian inference, the target distribution is the posterior distribution of model parameters $p(\boldsymbol{\theta}|\boldsymbol{x})$ capturing the uncertainty about the numerical values of parameters.
 We can leverage the fact that most mathematical models are generative in nature and as such can be used to perform multiple simulations of the process of interest.
 By specifying a prior distribution over the model parameters $p(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters.
-Thus, during the inference phase, the model generates samples $\boldsymbol{\theta}_{1}, \boldsymbol{\theta}_{2},...,\boldsymbol{\theta}_{L}$ from the observed data $\boldsymbol{x}_{obs}$, approximating the target posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$.
+Thus, during the inference phase, the model generates samples $\boldsymbol{\theta}^{(1)}, \boldsymbol{\theta}^{(2)},...,\boldsymbol{\theta}^{(L)}$ from the observed data $\boldsymbol{x}_{obs}$, approximating the target posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$.
 
@@
 To assess the recovery of the full posterior, we compute the Kullback-Leibler (KL) divergence \cite{hershey2007approximating} between the true and the approximate distributions for the toy example, and use simulation-based calibration (SBC, \cite{talts2018validating}) for the other examples where the analytic posterior is not available in closed-form.
-Details for computing the $NRMSE$, $R^2$, and SCB can be found in the \textbf{SI}.
+Details for computing the NRMSE, $R^2$, and SCB can be found in the \textbf{SI}.
 
@@
 To set the stage, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$.
-We model each $y^{(i)}$ as being conditionally Gaussian given $x^{(i)}$, that is, $y^{(i)} \sim \mathcal{N}(\boldsymbol{\theta}^{T}\boldsymbol{x}^{(i)},a^{-1})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
+We stack all $\boldsymbol{x}^{(i)}$ row-wise in a design matrix $\boldsymbol{X}$ and model the vector of outcomes $\boldsymbol{y}$ as being conditionally Gaussian given $\boldsymbol{X}$, that is, $\boldsymbol{y} \sim \mathcal{N}_{n}(\boldsymbol{X}\boldsymbol{\theta},a^{-1}\boldsymbol{I})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
 We place a $d$-dimensional diagonal Gaussian prior on the regression coefficients centered at $0$: $\boldsymbol{\theta} \sim \mathcal{N}_{d}(\boldsymbol{0},b^{-1}\boldsymbol{I})$ where $b$ is the precision of the prior ($b \equiv 1/\sigma_{\boldsymbol{\theta}}^{2}$).
@@
 \begin{align*}
-p(\boldsymbol{D}|\boldsymbol{\theta}) \propto \exp\left(-\frac{a}{2}\left(y-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(y - \boldsymbol{X}\boldsymbol{\theta}\right)\right) \numberthis 
+p(\boldsymbol{D}|\boldsymbol{\theta}) \propto \exp\left(-\frac{a}{2}\left(\boldsymbol{y}-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(\boldsymbol{y} - \boldsymbol{X}\boldsymbol{\theta}\right)\right) \numberthis
 \end{align*}
-where $\boldsymbol{X}$ denotes the design matrix containing all $\boldsymbol{x}^{(i)}$ stacked row-wise.
 Since the prior of $\boldsymbol{\theta}$ is conjugate to the likelihood (both are Gaussian distributions), the posterior of $\boldsymbol{\theta}$ is also Gaussian, given by:
 \begin{align*}
-p(\boldsymbol{\theta}|\boldsymbol{D}) \propto \exp\left(-\frac{a}{2}\left(y-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(y - \boldsymbol{X}\boldsymbol{\theta}\right)-\frac{b}{2}\boldsymbol{\theta}^{T}\boldsymbol{\theta}\right) \numberthis
+p(\boldsymbol{\theta}|\boldsymbol{D}) \propto \exp\left(-\frac{a}{2}\left(\boldsymbol{y}-\boldsymbol{X}\boldsymbol{\theta}\right)^{T}\left(\boldsymbol{y} - \boldsymbol{X}\boldsymbol{\theta}\right)-\frac{b}{2}\boldsymbol{\theta}^{T}\boldsymbol{\theta}\right) \numberthis
 \end{align*}
@@
 The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(50, 500)$ at each training iteration (Lines $2$-$9$ of \textbf{Algorithm} \ref{alg:1}).
+Training the networks took approx.
+half a day with the active learning approach.
+Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
+$5.5$ seconds.
 
@@
 The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+Training the networks took approx.
+half a day with the active learning approach.
+Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
+$6$ seconds.
 
@@
 The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(100, 1000)$ at each training iteration.
+Training the networks took a little less than a day with the active learning approach.
+Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
+$7.39$ seconds.
+
 \begin{figure}[H]
@@
 Thus, it is especially interesting to see if our method can recover the parameters, even if the process dynamics are still unfolding over time.
+Training the networks took approx.
+half a day with the active learning approach.
+Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
+$1.1$ seconds.
 
@@
 \centering
-\begin{subfigure}{.49\textwidth}
+\begin{subfigure}{.44\textwidth}
     \includegraphics[width=\textwidth]{SIR_plot.png}
@@
 \end{subfigure}
-\begin{subfigure}{.49\textwidth}
+\begin{subfigure}{.55\textwidth}
     \includegraphics[width=\textwidth]{SIR_states.png}
@@
 
-Our method combines the universal approximation power of deep learning methods with the important uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
-Besides being capable of performing rapid Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable \textit{parameter estimation machines} for various research domains.
-Moreover, the method is not confined solely to inference on intractable models, but can also prove as a viable alternative in modeling contexts where standard Bayesian or frequentist inference methods are available, but inference is nevertheless prohibitively slow.
+Our method combines the universal approximation power of deep learning methods with the crucial uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
+Besides being capable of performing full Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable \textit{parameter estimation machines} for various research domains.
+Moreover, it can also prove as a viable alternative in modeling contexts where standard inference methods are available, but inference is nevertheless computationally inconvenient.
 
@@
 We demonstrated that our method exhibits excellent recovery of the posterior means throughout the examples.
-We also showed that the recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters becomes smaller.
+We also showed that parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters becomes smaller.
 These are important and highly desirable properties of any parameter estimation method, as it mirrors the increase in information following increasing number of data points.
@@
 Indeed, there are many instances of research domains where a single model is extensively explored and independently fitted by multiple researches to test scientific hypotheses \cite{voss2019sequential, zappia2017splatter, ratcliff2008diffusion, de2002fitting}.
-These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for rapid inference on different datasets.
+These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for rapid inference on multiple datasets.
 In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
@@
 
-\section*{Conclusion}
-
-As formal theories in various scientific disciplines (especially in the younger sciences, such as, neuroscience, cognitive science, computational biology, etc.) become increasingly complex, the need for powerful and universally applicable likelihood-free estimation methods becomes increasingly pressing.
-In the present work, we addressed this need by introducing a method potentially applicable to \textit{any} modeling scenario in \textit{any} research domain where simulations from the mechanistic model can be obtained.
-We hope that the new method will enable researchers from a variety of fields to accelerate model-based inference and will further prove its utility beyond the examples considered in this paper.
-
 \acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

### 2019-07-11 latex/paper_supplementary.tex
New edit

--- 2019-06-28 latex/paper_supplementary.tex
@@
 
-\section*{Results}
+Complete code for all examples used in this paper is available at \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
-\subsection*{Performance metrics}
+\section*{Performance metrics}
 
@@
 
-\subsubsection*{Normalized Root Mean Squared Error}
+\subsection*{Normalized Root Mean Squared Error}
 The normalized root mean squared error (NRMSE) between a sample of true parameters $\{\theta^{(i)}\}_{i=1}^{n}$ and a sample of estimated parameters $\{\hat{\theta}^{(i)}\}_{i=1}^{n}$ is given by:
@@
 The NRMSE is zero when the estimates are exactly equal to the true values.
-\subsubsection*{Coefficient of Determination }
+\subsection*{Coefficient of Determination }
 The coefficient of determination $R^{2}$ gives the proportion of variance in a sample of true parameters $\{\theta^{(i)}\}_{i=1}^{n}$ that is "explained" by a sample of estimated parameters $\{\hat{\theta}^{(i)}\}_{i=1}^{n}$.
@@
 
-\subsubsection*{Kullback-Leibler Divergence} The Kullback-Leibler divergence ($D_{KL}$) quantifies the increase in entropy incurred by approximating a target probability distribution $P$ with a distribution $Q$.
+\subsection*{Kullback-Leibler Divergence} The Kullback-Leibler divergence ($D_{KL}$) quantifies the increase in entropy incurred by approximating a target probability distribution $P$ with a distribution $Q$.
 Its general form for absolutely continuous distributions is given by
@@
 
-\subsubsection*{Simulation-Based Calibration}
+\subsection*{Simulation-Based Calibration}
 Simulation-based calibration is a recently proposed method for validating the accuracy of posterior samples generated by a Bayesian sampling method \cite{talts2018validating}.
@@
 
-\subsection*{Model details}
+\section*{Invertible networks}
+Throughout all examples, we use a chain of $10$ conditional affine coupling blocks (cACB).
+Each internal network of each ACB is implemented as a fully connected neural network with $3$ to $4$ hidden layers with $64$ neurons each.
 
-\subsubsection*{Simulation of the Ricker model}
+\section*{Model details}
+
+\subsection*{Bayesian regression model}
+
+\subsubsection*{Summary network}
+We use a permutationally inariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+
+\subsection*{The Ricker model}
+
+\subsubsection*{Summary network}
+We use a bidirectional long short-term memory (LSTM) recurrent neural network for summarizing the Ricker time-series into fixed-size vectors.
+
+\subsubsection*{Simulation}
 We place the following uniform priors over the Ricker model parameters:
@@
 
-\subsubsection*{Simulation of the Lévy-Flight Model}
+\subsection*{The Lévy-Flight Model}
+
+\subsubsection*{Summary network}
+We use a permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
+
+\subsubsection*{Simulation}
 The following uniform priors over the LFM parameters are used for simulation:
@@
 
-\subsubsection*{Simulation of the stochastic SIR model}
+\subsection*{The stochastic SIR model}
+
+\subsubsection*{Summary network}
+We use a $1D$ convolutional neural network for summarizing the SIR time-series into fixed-size vectors.
+
+\subsubsection*{Simulation}
 We place the following uniform priors over the two rate parameters of the stochastic SIR model:
@@
 \end{align*}
-Figure XXX depicts some SIR datasets generated by parameters drawn from the specified prior.
+Figure XXX depicts some SIR datasets generated by parameters drawn from the specified priors.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{acb.png}
+\caption{Second figure}
+\end{figure}
+
+\subsection*{Single-Cell RNA Sequencing}
+
+\subsubsection*{Summary network}
+We use two permutationally invariant networks \cite{bloem2019probabilistic} for embedding the count matrices generated with the \textit{Splat} simulator \cite{zappia2017splatter} - one running over the cells and one running over the genes dimension.
+
+\subsubsection*{Simulation}
+We place the following uniform priors over the parameters of the \textit{Splat} simulation:
+\begin{align*}
+\beta &\sim \mathcal{U}(0.01, 1) \numberthis \\
+\gamma &\sim \mathcal{U}(0.01,\beta)  \numberthis 
+\end{align*}
+Figure XXX depicts some SIR datasets generated by parameters drawn from the specified priors.
 

### 2019-07-14 latex/paper_main.tex
Add attention and graph

@@
 
-An important goal of mathematical modeling is recovering the parameters of the model from observed data.
+An important goal of mathematical modeling is recovering parameters of the model from observed data.
 Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
@@
 It should be designed to follow the functional and probabilistic symmetries inherent in the data, e.g.
-a permutationally invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network for time-series data \cite{gers1999learning}, or a convolutional network for grid-like data \cite{goodfellow2016deep}.
+an invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network \cite{gers1999learning} or a convolutional network \cite{long2015fully} for data with temporal or spatial dependencies.
+Ideally, it should also be independent of the number of observations.
 The computation of summary statistics is a crucial aspect in likelihood-free inference.
-Previous approaches typically used hand-crafted summary statistics tailored to the specific application \cite{raynal2018abc, jiang2017learning, sunnaaker2013approximate}.
+Previous approaches typically used hand-crafted summary statistics tailored to the specific application \cite{raynal2018abc, sunnaaker2013approximate}.
 However, in many applications, finding an optimal set of summary statistics is far from obvious.
@@
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $500\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $50\%$ of the counts are $0$).
+We simulate $500\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
+Each count matrix contains the information about $20$ genes sequenced in $20$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{20\times20}$.
 The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
@@
 We demonstrated the utility of the method by applying it to models from various scientific domains dealing with several different data formats and data-generating mechanisms.
-Further, we explored two possible training approaches suitable for different simulation scenarios, namely an active learning approach, and a reference-table approach.
+Further, we explored two possible training approaches suitable for different simulation scenarios, namely, an active learning approach, and a reference-table approach.
 Both training approaches lead to excellent recovery of the true parameters throughout the examples considered in the current work.
 
-Our method combines the universal approximation power of deep learning methods with the crucial uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
+Our method combines the universal approximation power of deep learning methods \cite{goodfellow2016deep} with the crucial uncertainty quantification assets of Bayesian inference \cite{kendall2017uncertainties, gelman2013bayesian}.
 Besides being capable of performing full Bayesian inference on intractable mathematical models, our method provides a general framework for designing reusable \textit{parameter estimation machines} for various research domains.
@@
 
-First, the introduction of separate summary and inference neural network modules renders the invertible inference module independent of the shape or the size of the observed data.
-This is achieved by learning a fixed-size vector representation of the data through the summary module in an automatic, data-driven manner.
-This is particularly useful in settings where the most informative summary statistics are not known and thus potential information is lost through the choice of suboptimal summary function.
-However, if informative or even \textit{sufficient} statistics are available in a given domain, one might dispose with the summary module altogether and feed the summary statistics directly to the cINN.
+First, the introduction of separate summary and invertible networks renders the method independent of the shape or the size of the observed data.
+The summary network learns a fixed-size vector representation of the data in an automatic, data-driven manner.
+Since the summary network is optimized jointly with the invertible network, the learned data representation is encouraged to be maximally informative for parameter inference.
+This is particularly useful in settings where appropriate summary statistics are not known and, as a consequence, relevant information is lost through the choice of suboptimal summary function.
+However, if highly informative or even \textit{sufficient} statistics are available in a given domain, one might dispose with the summary network altogether and feed these statistics directly to the invertible network.
 
-Second, we showed that the ML loss directly maximizes the posterior over model parameters of interest, which is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
-Thus, inference is exact when the ML is minimized \cite{kingma2018glow, dinh2016density}.
-This \textit{optimal performance} claim is confirmed by our toy Bayesian regression example, in which we observe negligibly small deviations from the form of the true posterior.
-Further, researchers are often interested in some summary of the posterior, for instance, the posterior mean or the posterior variance \cite{radev2019towards, raynal2018abc}.
-We demonstrated that our method exhibits excellent recovery of the posterior means throughout the examples.
-We also showed that parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters becomes smaller.
-These are important and highly desirable properties of any parameter estimation method, as it mirrors the increase in information following increasing number of data points.
+Second, we showed that the ML loss directly maximizes the posterior over model parameters of interest.
+This is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
+Therefore, posterior inference is exact when the ML is globally minimized \cite{kingma2018glow, dinh2016density}.
+This theoretical claim is further confirmed by our results.
+Since researchers are often interested in some summary of the posterior (e.g., posterior means or variances), we also showed that our method exhibits excellent recovery of the posterior means throughout all examples.
+Further, parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters decreases.
+These are important and highly desirable properties of any Bayesian parameter estimation method, as it mirrors the increase in information following increasing number of observations.
 
@@
 Indeed, there are many instances of research domains where a single model is extensively explored and independently fitted by multiple researches to test scientific hypotheses \cite{voss2019sequential, zappia2017splatter, ratcliff2008diffusion, de2002fitting}.
-These research domains are expected to benefit the most from learning the “model universe” once and then inverting the model multiple times for rapid inference on multiple datasets.
+These research domains are expected to benefit the most from learning the \textit{model universe} once and then inverting the model multiple times for fast inference on multiple datasets.
 In this regard, our method is similar to the recently introduced prepaid method \cite{mestdagh2018prepaid} which uses a database of pre-computed summary statistics and nearest-neighbors for inference.
@@
 
-These advantages notwithstanding, some limitations of the method deserve mention.
+These advantages notwithstanding, some limitations of our method deserve mention.
 Even though high-level deep learning libraries, such as \textit{TensorFlow} or \textit{Torch}, allow for rapid and relatively straightforward development of various neural network architectures, the implementational burden associated with the current method is still reasonably high.
@@
 
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{We thank Jeffrey Rouder, Raphael Hartmann, David Izydorczyk, Hannes Wendler, Satya Almasian, and Karin Prillinger for their invaluable comments and suggestions that greatly improved the manuscript.}
 

### 2019-07-14 latex/paper_supplementary.tex
Add attention and graph

@@
 Each internal network of each ACB is implemented as a fully connected neural network with $3$ to $4$ hidden layers with $64$ neurons each.
+See \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN} for complete implementation details.
 
@@
 \subsubsection*{Summary network}
-We use a permutationally inariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+We use a permutationally inariant neural network as described in \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Regression_kl.png}
+\caption{Second figure}
+\end{figure}
 
@@
 \subsubsection*{Summary network}
-We use a permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
+We use a permutationally invariant neural network as described in \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{acb.png}
-\caption{Second figure}
-\end{figure}
-
 \subsection*{Single-Cell RNA Sequencing}
@@
 \subsubsection*{Summary network}
-We use two permutationally invariant networks \cite{bloem2019probabilistic} for embedding the count matrices generated with the \textit{Splat} simulator \cite{zappia2017splatter} - one running over the cells and one running over the genes dimension.
+We use two permutationally invariant networks as described in \cite{bloem2019probabilistic} for embedding the count matrices generated with the \textit{Splat} simulator \cite{zappia2017splatter}.
+The first network goes over the cells and the second over the genes axis of the count matrices.
 
@@
 \begin{align*}
-\beta &\sim \mathcal{U}(0.01, 1) \numberthis \\
-\gamma &\sim \mathcal{U}(0.01,\beta)  \numberthis 
+\alpha &\sim \mathcal{U}(1, 2) \numberthis \\
+\beta  &\sim \mathcal{U}(0.1, 5) \numberthis \\
+\mu^{L} &\sim \mathcal{U}(6, 10) \numberthis \\
+\sigma^{L} &\sim \mathcal{U}(1, 2) \numberthis \\
+\pi^{O} &\sim \mathcal{U}(0.01, 0.5) \numberthis \\
+\mu^{O} &\sim \mathcal{U}(0.5, 5) \numberthis \\
+\sigma^{O} &\sim \mathcal{U}(0.1, 2) \numberthis \\
+\phi &\sim \mathcal{U}(0.1, 0.9) \numberthis 
 \end{align*}
-Figure XXX depicts some SIR datasets generated by parameters drawn from the specified priors.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{acb.png}
-\caption{Second figure}
-\end{figure}
 

### 2019-07-15 latex/paper_main.tex
Splatter done

@@
 In what follows, we apply the method to a toy Bayesian regression example with conjugate priors, and then use it to estimate the parameters of challenging models from population dynamics, cognitive science, epidemiology, and genetics.
-Code for reproducing the results on all following examples is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
@@
 Details for computing the NRMSE, $R^2$, and SCB can be found in the \textbf{SI}.
+Code for reproducing the results on all following examples, as well as all network implementations, is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
@@
 \end{align*} 
-Thus, for known $a$ and $b$, the posterior can be easily computed.
+Thus, for known $a$ and $b$, the posterior of $\boldsymbol{\theta}$ can be easily computed.
 Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is good for testing the utility of our method.
@@
 
-What if the data does not contain information about a particular parameter?
+What if the data does not contain any information about a particular parameter?
 In this case, any good estimation method should detect this, and return the prior of the particular parameter.
@@
 Usually, at lower $T$s, the system has not converged to an equilibrium (i.e., not all individuals have transitioned from being $I$ to $R$).
-Thus, it is especially interesting to see if our method can recover the parameters, even if the process dynamics are still unfolding over time.
+Thus, it is especially interesting to see if our method can recover the rate parameters, even if the process dynamics are still unfolding over time.
 Training the networks took approx.
@@
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $500\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
-Each count matrix contains the information about $20$ genes sequenced in $20$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{20\times20}$.
+This is a challenging task for any Bayesian method, as the data-generating mechanism is a complicated algorithm, and no likelihood function for joint estimation of the parameters is available.
+We simulate $200\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation).
+Each count matrix contains the information about $40$ genes sequenced in $100$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{40\times100}$.
 The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
 The parameter estimation task is thus to recover $\boldsymbol{\theta} = (\alpha,\beta,\mu^{L},\sigma^{L},\pi^{O},\mu^{O},\sigma^{O},\phi)$ from an observed count matrix $\boldsymbol{X}$.
-We train the networks for 50 epochs through the entire dataset and evaluate the performance on a separate validation set of $500$ count matrices.
+We train the networks for 50 epochs through the entire dataset and evaluate the performance on a separate validation set of $300$ count matrices.
+To stabilize training and avoid exploding gradients due to large counts in the input matrices, we apply a log transformation to the data.
+Simulating the data took approx.
+$8$ hours.
+Training the networks took approx.
+one day with the reference table approach.
+Inference on $300$ datasets with $2000$ posterior samples per parameter took approx.
+$1.5$ seconds.
 \begin{figure}[H]

### 2019-07-15 latex/paper_supplementary.tex
Splatter done

--- 2019-07-11 latex/paper_supplementary.tex
@@
 \State {\textbf{end for}}
-\State {Create a histogram of $\{r^{(i)}\}_{i=1}^{n}$ and analyze it for uniformity}
+\State {Create a histogram of $\{r^{(i)}\}_{i=1}^{n}$ and inspect it for uniformity}
 \end{algorithmic}
@@
 Each internal network of each ACB is implemented as a fully connected neural network with $3$ to $4$ hidden layers with $64$ neurons each.
+See \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN} for complete implementation of all networks used throughout this paper.
 
@@
 \subsubsection*{Summary network}
-We use a permutationally inariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+We use the recently introduced permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+This network architecture is a good choice for this example, as each observation is independent of all other, and thus all permutations of a given dataset should lead to the same parameter estimates.
+Furthermore, invariant networks are independent of the number of observations.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Regression_kl.png}
+\caption{KL divergence ($D_{KL}$) between the true and estimated posteriors of the regression weights in our toy Bayesian regression example.
+The $D_{KL}$ is depicted at every $n$ observations used during training.}
+\end{figure}
 
@@
 \subsubsection*{Summary network}
-We use a bidirectional long short-term memory (LSTM) recurrent neural network for summarizing the Ricker time-series into fixed-size vectors.
+We use a bidirectional long short-term memory (LSTM) recurrent neural network \cite{gers1999learning} for summarizing the Ricker time-series into fixed-size vectors.
+The LSTM network architecture is a good choice for this example, as it is able to capture long-term in datasets with temporal or spatial autocorrelations.
+LSTMs can also easily deal with variable length time-series.
 
@@
 We use a permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
+Similarly to the toy Regression example, each response in an RT dataset is assumed to be independent of all others, so permutations of the dataset must lead to the same parameter estimates.
 
@@
 \subsubsection*{Summary network}
-We use a $1D$ convolutional neural network for summarizing the SIR time-series into fixed-size vectors.
+We use a $1D$ fully convolutional neural network \cite{long2015fully} for summarizing the SIR time-series into fixed-size vectors.
+Here, we chose a convolutional network architecture over the previously mentioned LSTM, as convolutional networks are more computationally efficient.
+Further, we wanted to underline the utility of 1D convolutional networks for multidimensional time-series data.
+Finally, convolutional networks can also deal with variable input sizes.
 
@@
 \subsubsection*{Summary network}
-We use two permutationally invariant networks \cite{bloem2019probabilistic} for embedding the count matrices generated with the \textit{Splat} simulator \cite{zappia2017splatter} - one running over the cells and one running over the genes dimension.
+We use a network implementing the recently introduced \textit{self-attention} mechanism \cite{vaswani2017attention}.
+Self-attention architectures can learn to represent complicated pairwise relationships between observations.
+This property aligns nicely with the task at hand, since each entry in the observed count matrices is not independent of the other entries.
+Moreover, these dependencies do not exhibit the typical spatial patterns observed, for instance, in natural images.
 
@@
 
-\movie{Type caption for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}
-
-\movie{A third movie, just for kicks.}
-
-\dataset{dataset_one.txt}{Type or paste caption here.}
-
-\dataset{dataset_two.txt}{Type or paste caption here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}
-
 \bibliography{references}

### 2019-07-16 latex/paper_main.tex
seqRNA finished

--- 2019-07-14 latex/paper_main.tex
@@
 With these settings, we perform $50\ 000$ to $100\ 000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
-Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINN).
+Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINNs and summary networks).
 All networks were implemented in Python using the \textit{TensorFlow} library \cite{abadi2016tensorflow} and trained on a single-GPU machine equipped with NVIDIA\textsuperscript{\textregistered} GTX1060 graphics card.
@@
 During training of the networks, we simulate time-series from the Ricker model with varying lengths.
-The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration.
 Training the networks took approx.
@@
 
-During training of the networks, we simulate response times data from two experimental conditions with two different drift rates (see \textbf{SI} for details of the simulation), since this such a design is often encountered in psychological research.
+During training of the networks, we simulate response times data from two experimental conditions with two different drift rates, since such a design is often encountered in psychological research.
 The parameter estimation task is thus to recover the parameters $\boldsymbol{\theta} = (v_{0}, v_{1}, a, t_{0}, zr, \alpha)$ from two-dimensional \textit{i.i.d.} RT data $\boldsymbol{X} = (\boldsymbol{x}_{1},\boldsymbol{x}_{2},...,\boldsymbol{x}_{n})$ containing variable number or RT trials.
@@
 During training of the networks, we simulate time-series from the stochastic SIR model with varying lengths.
-The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(200, 500)$ at each training iteration (see the \textbf{SI} for more details about the simulation).
+The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(200, 500)$ at each training iteration.
 Usually, at lower $T$s, the system has not converged to an equilibrium (i.e., not all individuals have transitioned from being $I$ to $R$).
-Thus, it is especially interesting to see if our method can recover the parameters, even if the process dynamics are still unfolding over time.
+Thus, it is especially interesting to see if our method can recover the rate parameters, even if the process dynamics are still unfolding over time.
 Training the networks took approx.
@@
     	\includegraphics[width=\textwidth]{SIR_density.png}
-    	\caption{Example posterior samples ($T=500$}
+    	\caption{Example posterior samples ($T=500$)}
     	\label{fig:Fig.6f}
@@
 The results on the SIR model are depicted in \autoref{fig:Fig.6}.
-Similar to the previous examples, we observe very good recovery of the true parameters, with NRMSE at $T=500$ around $0.04$, and $R^{2}$s around $0.98$.
+In line with the previous examples, we observe very good recovery of the true parameters, with NRMSE at $T=500$ around $0.04$, and $R^{2}$s around $0.98$.
 Further, we observe decent performance even at smaller $T$s, with performance increasing as $T$ increases.
@@
 \end{table}
+
 As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-We simulate $500\ 000$  \textit{Gene x Cell} count matrices with different parameter settings (see \textbf{SI} for details of the simulation) and discard implausible simulations (e.g., matrices where more than $90\%$ of the counts are $0$).
-Each count matrix contains the information about $20$ genes sequenced in $20$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{20\times20}$.
+This is a challenging task for any Bayesian method, as the data-generating mechanism is a complicated algorithm, and no likelihood function for joint estimation of the parameters is available.
+We simulate $200\ 000$  \textit{Gene x Cell} count matrices with different parameter settings.
+Each count matrix contains the information about $40$ genes sequenced in $100$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{40\times100}$.
 The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
 The parameter estimation task is thus to recover $\boldsymbol{\theta} = (\alpha,\beta,\mu^{L},\sigma^{L},\pi^{O},\mu^{O},\sigma^{O},\phi)$ from an observed count matrix $\boldsymbol{X}$.
-We train the networks for 50 epochs through the entire dataset and evaluate the performance on a separate validation set of $500$ count matrices.
+We train the networks for 30 epochs through the entire dataset and evaluate the performance on a separate validation set of $300$ count matrices.
+To stabilize training and avoid exploding gradients due to large counts in the input matrices, we apply a log transformation to each $\boldsymbol{X}$.
+Simulating the data took approx.
+$8$ hours.
+Training the networks took approx.
+one day with the reference table approach.
+Inference on $300$ datasets with $2000$ posterior samples per parameter took approx.
+$1.7$ seconds.
 \begin{figure}[H]
@@
 \begin{subfigure}{.49\textwidth}
-    \centering
     \includegraphics[width=\textwidth]{Splatter_plot.png}
@@
 \begin{subfigure}{.49\textwidth}
-    \centering
     \includegraphics[width=\textwidth]{Splatter_scatter.png}
-    \caption{Parameter recovery }
+    \caption{Parameter recovery}
     \label{fig:Fig.7b}
 \end{subfigure}
-\caption[short]{Results on the \textit{Splat} scRNA simulation model. \textbf{(a)} The \textit{Splat} simulation diagram; \textbf{(b)} Parameter recovery.} \label{fig:Fig.7}
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{Splatter_sbc.png}
+    \caption{Simulation-based calibration (SBC)}
+    \label{fig:Fig.7c}
+\end{subfigure}
+\begin{subfigure}{.49\textwidth}
+    \includegraphics[width=\textwidth]{Splatter_density.png}
+    \caption{Example posterior samples}
+    \label{fig:Fig.7d}
+\end{subfigure}
+
+\caption[short]{Results on the \textit{Splat} scRNA simulation model.
+\textbf{(a)} The \textit{Splat} simulation diagram; \textbf{(b)} Parameter recovery; \textbf{(c)} Simulation-based calibraton.
+The SCB plots indicate a slight underestimation of the true posterior variance.
+\textbf{(d)} Example full parameter posteriors from three datasets.} \label{fig:Fig.7}
 \end{figure}
 The results on the scRNA example are depicted in \autoref{fig:Fig.7}.
-We observe excellent parameter recovery on the validation set.
-\textbf{TODO: Train the full method.
-These results are preliminary}.
+First, recovery of most parameters is very good, with maximum $R^{2}$ of $0.98$ and minimum NRMSE of $0.039$ for the library location parameter $\mu^{L}$.
+Interestingly, the rate parameter $\beta$ appears to be nearly unrecoverable, as indexed by the scatterplots in \autoref{fig:Fig.7b}.
+Furthermore, the estimated posterior of $\beta$ is often bimodal, covering the whole prior range (\autoref{fig:Fig.7c}).
+This might indicate that the parameter is not identifiable, or that some multivariate trade-offs exist in the parameter space of the \textit{Splat} model.
+The latter is also confirmed by observing a few marked outliers in the scatterplots of $\alpha$, $\sigma^{O}$, and $\pi^{O}$.
+Finally, the SCB histograms indicate that our method tends to slightly underestimate the true posterior variance of the $\alpha$, $\beta$, $\mu^{O}$ $\pi^{O}$, and $\sigma^{O}$ parameters.
+This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear when employing the active learning approach.
 

### 2019-07-16 viz.py

@@ -251,7 +251,7 @@ def plot_true_est_posterior(model, n_samples, param_names, n_test=None, data_gen
         f.savefig("figures/{}_{}n_density.png".format(filename, X_test.shape[1]), dpi=600, bbox_inches='tight')
 
 
-def plot_sbc(model, n_samples, X_test, theta_test, param_names, 
+def plot_sbc(model, n_samples, X_test, theta_test, param_names, bins=None,
             figsize=(15, 5), show=True, filename=None, font_size=12):
     """
     Plots the simulation-based posterior checking histograms as advocated by Talts et al. (2018).
@@ -283,7 +283,7 @@ def plot_sbc(model, n_samples, X_test, theta_test, param_names,
 
     # Plot histograms
     for j in range(len(param_names)):
-        sns.distplot(ranks[:, j], kde=False, ax=axarr[j], rug=True, hist_kws=dict(edgecolor="k", linewidth=1))
+        sns.distplot(ranks[:, j], kde=False, ax=axarr[j], rug=True, hist_kws=dict(edgecolor="k", linewidth=1), bins=bins)
         axarr[j].set_title(param_names[j])
         axarr[j].spines['right'].set_visible(False)
         axarr[j].spines['top'].set_visible(False)

### 2019-07-16 latex/paper_main.tex
Prepare final notebooks

@@
 Details for computing the NRMSE, $R^2$, and SCB can be found in the \textbf{SI}.
-Code for reproducing the results on all following examples, as well as all network implementations, is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
+Code for reproducing the results on all following examples, as well as for all network implementations, is freely available at: \href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}.
 
@@
 To set the stage, assume we have observed a dataset $\boldsymbol{D} = \{(\boldsymbol{x}^{(i)},y^{(i)})\}_{i=1}^{n}$ with $\boldsymbol{x} \in \mathbb{R}^{d}$ and $y \in \mathbb{R}$.
-We stack all $\boldsymbol{x}^{(i)}$ row-wise in a design matrix $\boldsymbol{X}$ and model the vector of outcomes $\boldsymbol{y}$ as being conditionally Gaussian given $\boldsymbol{X}$, that is, $\boldsymbol{y} \sim \mathcal{N}_{n}(\boldsymbol{X}\boldsymbol{\theta},a^{-1}\boldsymbol{I})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
+We stack all $\boldsymbol{x}^{(i)}$ row-wise in a design matrix $\boldsymbol{X}$ and model the vector of outcomes $\boldsymbol{y}$ as being conditionally Gaussian given $\boldsymbol{X}$, that is, $\boldsymbol{y} \sim \mathcal{N}_{n}(\boldsymbol{X}\boldsymbol{\theta},a^{-1}\boldsymbol{I})$ where $\boldsymbol{\theta} \in \mathbb{R}^{d}$ is the vector of regression coefficients and $a$ is the precision (inverse noise variance, $a \equiv 1/\sigma_{y}^{2}$).
 We place a $d$-dimensional diagonal Gaussian prior on the regression coefficients centered at $0$: $\boldsymbol{\theta} \sim \mathcal{N}_{d}(\boldsymbol{0},b^{-1}\boldsymbol{I})$ where $b$ is the precision of the prior ($b \equiv 1/\sigma_{\boldsymbol{\theta}}^{2}$).
@@
 \end{align*}
-Therefore, the posterior has the form $p(\boldsymbol{\theta}|\boldsymbol{D}) = \mathcal{N}_{d}(\boldsymbol{\theta}|\boldsymbol{\mu},\boldsymbol{\Lambda}^{-1})$ where $\boldsymbol{\Lambda}$ denotes the posterior precision matrix (inverse covariance matrix), $\boldsymbol{\mu}$ the posterior mean vector, which are computed as follows:
+Therefore, the posterior has the form $p(\boldsymbol{\theta}|\boldsymbol{D}) = \mathcal{N}_{d}(\boldsymbol{\theta}|\boldsymbol{\mu},\boldsymbol{\Lambda}^{-1})$ where $\boldsymbol{\Lambda}$ denotes the posterior precision matrix (inverse covariance matrix), and $\boldsymbol{\mu}$ denotes the posterior mean vector.
+The posterior moments are computed as follows:
 \begin{align*}
@@
 Thus, for known $a$ and $b$, the posterior of $\boldsymbol{\theta}$ can be easily computed.
-Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is good for testing the utility of our method.
+Even though in real-world applications $a$ is usually not known and a hierarchical model is used instead, the current example is well suited for testing the utility of our method.
 
@@
 Visual inspections reveals a close match between analytic and estimated posteriors; \textbf{(c)} Posterior mean recovery is almost perfect at $n=500$; \textbf{(d)} NRMSE and $R^{2}$ over all $n$.
-Performance improves with increasing number of data points; \textbf{(e)} Analytic vs.
+Performance improves with increasing number of observed data points; \textbf{(e)} Analytic vs.
 estimated variance over all $n$.
@@
 The parameter $\sigma$ seems to be the hardest to recover.
-Inspecting the full posteriors, we further see that the posterior distribution of the dummy noise variable $u$ closely resembles the prior, as expected due to the complete lack of mutual information between the data $x$ and $u$ (\autoref{fig:Fig.4b}).
+Inspecting the full posteriors, we further see that the posterior distribution of the dummy noise variable $u$ closely resembles the prior.
+This is expected due to the complete lack of mutual information between the observed data $x$ and $u$ (\autoref{fig:Fig.4b}).
 Finally, the plots of the rank statistic computed for SCB suggest no systematic distortions of the posterior across all parameters (\autoref{fig:Fig.4d}).
@@
 
-As a last example for our method, we use simulations from the \textit{Splat} model and attempt to recover the data-generating parameters.
-This is a challenging task for any Bayesian method, as the data-generating mechanism is a complicated algorithm, and no likelihood function for joint estimation of the parameters is available.
+As a last example for our method, we use simulations from the \textit{Splat} model to recover the data-generating parameters.
+This is a challenging task for any Bayesian method, as the data-generating mechanism is a complicated algorithm with no available likelihood for joint estimation of the parameters.
+Moreover, the observed data exhibits multiple local and non-local dependencies.
 We simulate $200\ 000$  \textit{Gene x Cell} count matrices with different parameter settings.
-Each count matrix contains the information about $40$ genes sequenced in $100$ different cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{40\times100}$.
+Each entry in a count matrix contains the number of times a gene has been sequenced in a different cell.
+We simulate counts for $40$ genes and $100$ cells, that is, each $\boldsymbol{X} \in \mathbb{R}^{40\times100}$.
 The parameters considered for estimation, along with a short description, are listed in \autoref{table:Table 1}.
+
 The parameter estimation task is thus to recover $\boldsymbol{\theta} = (\alpha,\beta,\mu^{L},\sigma^{L},\pi^{O},\mu^{O},\sigma^{O},\phi)$ from an observed count matrix $\boldsymbol{X}$.
@@
 Finally, the SCB histograms indicate that our method tends to slightly underestimate the true posterior variance of the $\alpha$, $\beta$, $\mu^{O}$ $\pi^{O}$, and $\sigma^{O}$ parameters.
-This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear when employing the active learning approach.
+This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear if the active learning approach is employed.
 
@@
 
-In the current work, we proposed and explored a novel end-to-end likelihood-free method which uses invertible neural networks to perform approximate Bayesian inference on any mathematical process model.
-We demonstrated the utility of the method by applying it to models from various scientific domains dealing with several different data formats and data-generating mechanisms.
+In the current work, we proposed and explored a novel method which uses invertible neural networks to perform approximate Bayesian inference on any mathematical process model.
+The method requires only simulations from a model to learn a probabilistic mapping between data and parameters.
+We demonstrated the utility of the method by applying it to models and data from various scientific domains.
 Further, we explored two possible training approaches suitable for different simulation scenarios, namely, an active learning approach, and a reference-table approach.
-Both training approaches lead to excellent recovery of the true parameters throughout the examples considered in the current work.
+Both training approaches lead to excellent parameter estimation throughout the examples considered in the current work.
 
@@
 
-Inspired by previous machine learning approaches to likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, our method shares many of the advantages of these methods and further overcomes some of their important limitations.
+Inspired by previous machine learning approaches to likelihood-free inference \cite{radev2019towards, hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, our method shares many of the advantages of these methods and, further, overcomes some of their important limitations.
 
@@
 The summary network learns a fixed-size vector representation of the data in an automatic, data-driven manner.
-Since the summary network is optimized jointly with the invertible network, the learned data representation is encouraged to be maximally informative for parameter inference.
-This is particularly useful in settings where appropriate summary statistics are not known and, as a consequence, relevant information is lost through the choice of suboptimal summary function.
+Since the summary network is optimized jointly with the invertible network, the learned data representation is encouraged to be maximally informative for parameter estimation.
+This is particularly useful in settings where appropriate summary statistics are not known and, as a consequence, relevant information is lost through the choice of suboptimal summary functions.
 However, if highly informative or even \textit{sufficient} statistics are available in a given domain, one might dispose with the summary network altogether and feed these statistics directly to the invertible network.
 
-Second, we showed that the ML loss directly maximizes the posterior over model parameters of interest.
-This is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
+Second, we showed that the ML loss directly maximizes the posterior over the target model parameters without any assumptions on the shape of the posterior.
+This is in contrast to ELBO-based methods which optimize a lower-bound on the posterior \cite{papamakarios2016fast, kingma2014auto}, and usually assume Gaussian posteriors.
 Therefore, posterior inference is exact when the ML is globally minimized \cite{kingma2018glow, dinh2016density}.
-This theoretical claim is further confirmed by our results.
-Since researchers are often interested in some summary of the posterior (e.g., posterior means or variances), we also showed that our method exhibits excellent recovery of the posterior means throughout all examples.
+Our results support this theoretical claim.
+Further, since researchers are often interested in some summary of the posterior (e.g., posterior means or variances), we also showed that our method exhibits excellent recovery of the posterior means throughout all examples.
 Further, parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters decreases.
@@
 This not only makes the global sharing of pre-trained parameter estimation networks across researchers extremely easy, but also makes their local storage very efficient.
-Moreover, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern GPUs.
+Finally, all computations involved in our method benefit from a high degree of parallelism and can thus utilize the advantages of modern parallel computing paradigms.
 
 These advantages notwithstanding, some limitations of our method deserve mention.
-Even though high-level deep learning libraries, such as \textit{TensorFlow} or \textit{Torch}, allow for rapid and relatively straightforward development of various neural network architectures, the implementational burden associated with the current method is still reasonably high.
+Although high-level deep learning libraries, such as \textit{TensorFlow} or \textit{Torch}, allow for rapid and relatively straightforward development of various neural network architectures, the implementational burden associated with the current method is still reasonably high.
 In order to ease the understanding and independent application of the method, we provide fully functioning code to reproduce and study all of the examples tackled in this paper (\href{https://github.com/stefanradev93/cINN}{https://github.com/stefanradev93/cINN}).
-In addition, we also provide implementation of all tools for performance validation used throughout the paper.
+In addition, we provide implementation of all metrics and visualization tools for performance validation used throughout the paper.
 Moreover, we are currently developing a general user-friendly software, which should abstract away most of the methodological complexities from the user.
@@
 However, we observe that many of the default hyperparameter values are sufficient to achieve excellent performance, and starting with a relatively large default network of $10$ ACBs does not appear to hurt performance or destabilize training, even if the mathematical model to be learned is relatively simple.
-We expect that a single architecture should be able to perform well on almost all models from a given domain (i.e., a single architecture for decision-making models).
+Based on our results, we expect that a single architecture should be able to perform well on almost all models from a given domain.
 Future research should investigate the question of generality by applying the method to challenging parameter estimation tasks across different research domains.
+
+We hope that the new method will enable researchers from a variety of fields to accelerate model-based inference and will further prove its utility beyond the examples considered in this paper.
 

### 2019-07-16 models.py

@@ -417,3 +417,105 @@ class InvariantNetwork(tf.keras.Model):
         x = self.equiv(x)
         out = self.inv(x)
         return out
+
+
+class SelfAttentionHead(tf.keras.Model):
+    """Implements a single self-attention head."""
+
+    def __init__(self, key_dim=64, dense_dim=128):
+        super(SelfAttentionHead, self).__init__()
+        
+        self.Wq = tf.keras.layers.Dense(key_dim, use_bias=False, kernel_initializer='glorot_uniform')
+        self.Wk = tf.keras.layers.Dense(key_dim, use_bias=False, kernel_initializer='glorot_uniform')
+        self.Wv = tf.keras.layers.Dense(key_dim, use_bias=False, kernel_initializer='glorot_uniform')
+        
+        self.dense_part = tf.keras.models.Sequential([
+            tf.keras.layers.Dense(dense_dim, activation='elu', kernel_initializer='glorot_uniform'),
+            tf.keras.layers.Dense(dense_dim, activation='elu', kernel_initializer='glorot_uniform'),
+            tf.keras.layers.Dense(dense_dim, activation='elu', kernel_initializer='glorot_uniform'),
+        ])
+        
+        
+        self.scale = np.sqrt(key_dim)
+              
+    def call(self, x):
+        """
+        Passes the input through an attention head and then through a shared dense layer.
+        ----------
+
+        Arguments:
+        x : tf.Tensor of shape (batch_size, n, m) - the input where n is the 'time' or 'samples' dimensions 
+            over which pooling is performed and m is the input dimensionality
+        ----------
+
+        Returns:
+        out : tf.Tensor of shape (batch_size, n, h_dim) -- the trasnformed representation of the input
+        """
+
+        # Obtain queries, keys, and values
+        queries = self.Wq(x)
+        keys = self.Wk(x)
+        values = self.Wv(x)
+        
+        # Compute self-attention matrices
+        qk_t = tf.einsum('bij,bkj->bik', queries, keys)
+        weights = tf.nn.softmax(qk_t / self.scale, axis=-1)
+        z = tf.einsum('bik,bkj->bij', weights, values)
+        
+        # Pass through final dense
+        out = self.dense_part(z)
+        return out
+
+
+class AttentionNetwork(tf.keras.Model):
+    """
+    Implements a network with multi-head self-attention architecture after Vaswani et al. (2017).
+    """
+
+    def __init__(self, n_heads=8, key_dim=64, dense_dim=128):
+        """
[…]

### 2019-07-17 latex/paper_main.tex
Add new version of text

@@
 
-\title{Learning complex stochastic models with invertible neural networks: a likelihood-free Bayesian approach}
+\title{Learning complex stochastic models with invertible neural networks: a likelihood-free approach to Bayesian inference}
 
@@
 \affil[1]{Institute of Psychology, Heidelberg University, Hauptstr. 47-51, 69117 Heidelberg, Germany}
-\affil[2]{Heidelberg Collaboratory for Image Processing (HCI), Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany}
+\affil[2]{Visual Learning Lab, Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany}
 
@@
 
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
 \authordeclaration{This research was supported by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation; grant number GRK 2277 "Statistical Modeling in Psychology")}
@@
 \begin{abstract}
-Mathematical models of complex processes are ubiquitous throughout the sciences.
-As the processes under study and the models describing them become increasingly complex, parameter estimation with standard Bayesian and frequentist methods can quickly become intractable.
-To address this issue, we propose a novel method for likelihood-free inference based on invertible neural networks.
+Mathematical models of complex stochastic processes are ubiquitous throughout the sciences.
+Estimating the parameters of a mathematical model becomes notably difficult when processes and model descriptions become increasingly complex.
+We address this issue with a novel method for likelihood-free parameter inference based on invertible neural networks.
 The method is capable of performing full Bayesian inference on large datasets by training the networks on simulated data to learn a probabilistic mapping between parameters and data.
 The method is independent of any particular data representation, as it incorporates a summary network trained to embed the observed data into fixed-size vectors in a data-driven way.
-This makes the method applicable to various scenarios where standard inference techniques fail.
-We demonstrate the utility of the method on a toy model with known analytic posterior and on example models from population dynamics, epidemiology, cognitive science and genetics.
+This makes the method applicable to various modeling scenarios where standard inference techniques fail.
+We demonstrate the utility of the method on a simple toy regression model with known analytic posterior and on various models from population dynamics, epidemiology, cognitive science and genetics.
 We argue that our method provides a general framework for building reusable parameter estimation machines for any process model from which data can be simulated.
@@
 \dropcap{M}athematical models are formal descriptions of scientific theories.
-In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a forward model $q$ mimicking the process by which manifest data $\boldsymbol{x}$ arise from latent parameters:
+In its most abstract form, a mathematical model is specified by a set of parameters $\boldsymbol{\theta}$ and a forward model $S$ mimicking the process by which manifest data $\boldsymbol{x}$ arise from latent parameters.
+Independent noise $\xi$ is often included, giving:
 \begin{align*}
-\boldsymbol{x}= q(\boldsymbol{\theta}) \numberthis \label{eqn:1} 
+\boldsymbol{x}= S(\boldsymbol{\theta},\xi) \numberthis \label{eqn:1} 
 \end{align*}
-The functional form of $q$, which can represent an arbitrarily complex process by an arbitrarily complicated expression, is usually guided by a well-founded theoretical framework.
-For instance, $q$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm controlling the rate of gene expression in certain cells.
+The functional form of $S$, which can represent an arbitrarily complex process by an arbitrarily complicated expression, is usually guided by a well-founded theoretical framework.
+For instance, $S$ can be a stochastic differential equation describing the dynamics of single neurons in the brain, or a step-by-step biological algorithm controlling the rate of gene expression in certain cells.
 Thus, it is only within the context of a theory that a meaningful interpretation in terms of some mechanism can be attached to the parameters of a mathematical model.
-Examples of mathematical models can be found in various scientific domains, for instance, genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical, geritz2004mechanistic}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, to name just a few.
+Examples of mathematical models can be found in all scientific domains, for instance, genetics \cite{zappia2017splatter, beaumont2002approximate}, cognitive science \cite{palestro2018likelihood, usher2001time}, neuroscience \cite{hwang2018conditional, lueckmann2017flexible}, population dynamics \cite{wood2010statistical, geritz2004mechanistic}, epidemiology \cite{keeling2011modeling,hethcote2000mathematics}, to name just a few.
 
 An important goal of mathematical modeling is recovering parameters of the model from observed data.
-Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$ exactly.
+Idealized parameter estimation involves computing the inverse (backward) model $\boldsymbol{\theta} = S^{-1}(\boldsymbol{x},\xi)$ exactly.
 Unfortunately, this inverse rarely exists, and estimating parameters remains computationally inconvenient in a wide range of settings.
@@
 A subset of likelihood-free methods includes approximate Bayesian computation (ABC) methods, which aim at preserving the advantages of Bayesian data analysis even when the likelihood function is intractable or practically impossible to compute \cite{turner2014generalized, sunnaaker2013approximate, csillery2010approximate}.
-ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the forward model $q(\boldsymbol{\theta})$ using the sampled parameters.
-Correspondingly, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $q(\boldsymbol{\theta})$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
+ABC methods approximate the likelihood function by repeatedly sampling parameters from a pre-specified prior distribution $p(\boldsymbol{\theta})$ and then simulating multiple datasets by running the forward model $S(\boldsymbol{\theta},\xi)$ using the sampled parameters.
+Correspondingly, the core ingredients of ABC methods are a prior on $\boldsymbol{\theta}$, and a generative model $S(\boldsymbol{\theta},\xi)$, usually specified as a function code in a general-purpose programming language \cite{csillery2010approximate, mertens2018abrox}.
 
@@
 The most common approach has been to cast the problem of parameter estimation as a supervised learning task.
-In this setting, a large dataset of the form $\boldsymbol{D} = \{h(\boldsymbol{x}^{(i)}), \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ is created by repeatedly sampling from $p(\boldsymbol{\theta})$ and simulating an artificial dataset $\boldsymbol{x}$ by running $q(\boldsymbol{\theta})$ with the sampled parameters.
+In this setting, a large dataset of the form $\boldsymbol{D} = \{h(\boldsymbol{x}^{(i)}), \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ is created by repeatedly sampling from $p(\boldsymbol{\theta})$ and simulating an artificial dataset $\boldsymbol{x}$ by running $S(\boldsymbol{\theta},\xi)$ with the sampled parameters.
 Usually, the dimensionality of the simulated data is reduced by computing summary statistics  with a fixed summary function $h(\boldsymbol{x})$.
 Then, a supervised learning algorithm $f(h(\boldsymbol{x});\boldsymbol{\phi}) = \widehat{\boldsymbol{\theta}}$ with learnable parameters $\boldsymbol{\phi}$ (e.g., linear regression, random forest, neural network) is trained on the simulated data to output an estimate of the true data generating parameters.
-Thus, $f(h(\boldsymbol{x});\boldsymbol{\phi})$ essentially attempts to “learn” the intractable inverse model $\boldsymbol{\theta} = q^{-1}(\boldsymbol{x})$.
+Thus, $f(h(\boldsymbol{x});\boldsymbol{\phi})$ essentially attempts to \textit{learn} the intractable inverse model $\boldsymbol{\theta} = S^{-1}(\boldsymbol{x},\xi)$.
 
 Inspired by previous machine learning approaches \cite{hwang2018conditional, mestdagh2018prepaid, raynal2018abc, jiang2017learning, lueckmann2017flexible, papamakarios2016fast}, the current work proposes a novel and universal likelihood-free method capable of performing full Bayesian inference on any mathematical process model from which simulations can be obtained.
-It treats parameter inference as a task of inverting the forward model in Eq.\ref{eqn:1} and achieves this by drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
+It follows the pioneering work of \cite{ardizzone2018analyzing} by treating parameter inference as a task of inverting the forward model in Eq.\ref{eqn:1} and drawing on the modern framework of deep probabilistic modeling for tackling intractable posteriors \cite{ardizzone2018analyzing, kingma2018glow, grover2018flow, dinh2016density}.
 The method integrates two separate deep neural networks modules (detailed in the \textbf{Methods} section; see \autoref{fig:Fig.1}) trained jointly on simulated data: a \textit{summary network} and an \textit{invertible network}.
@@
 It should be designed to follow the functional and probabilistic symmetries inherent in the data, e.g.
-an invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network \cite{gers1999learning} or a convolutional network \cite{long2015fully} for data with temporal or spatial dependencies.
+a permutation invariant network for \textit{i.i.d.} data \cite{bloem2019probabilistic}, a recurrent network \cite{gers1999learning} or a convolutional network \cite{long2015fully} for data with temporal or spatial dependencies.
 Ideally, it should also be independent of the number of observations.
@@
 The \textit{invertible network} is responsible for learning the posterior of the model parameters given the observed and summarized data.
-It is based on the recently developed flow-based architecture \cite{kingma2018glow, grover2018flow, dinh2016density}.
+It is based on the recently developed flow-based architecture \cite{ardizzone2019guided, kingma2018glow, grover2018flow, dinh2016density}.
 Flow-based methods provide exact latent-variable inference and log-likelihood evaluation when operating at optimum.
@@
 To illustrate the utility of the new method, we first apply it to a toy Bayesian regression model with known posterior.
-Then, we present applications to intractable models from cognitive science, population dynamics, epidemiology, and genetics and demonstrate state-of-the art parameter recovery.
+Then, we present applications to intractable models from cognitive science, population dynamics, epidemiology, and genetics and demonstrate excellent parameter recovery.
 Across the examples, we introduce multiple tools to validate the performance of our method.
@@
 Most popular deep probabilistic models entail two phases.
-During the \textit{training phase}, a transformation from the simple to the desired target distribution is learned by optimizing a cost function via backpropagation (see \autoref{fig:Fig.1a}).
+During the \textit{training phase}, a transformation from the simple to the desired target distribution (a so called \textit{pushforward}) is learned by optimizing a cost function via backpropagation (see \autoref{fig:Fig.1a}).
 During the \textit{inference phase}, samples from the target distribution are obtained by sampling from the simple distribution and applying the transformation learned during the training phase (see \autoref{fig:Fig.1b}).
@@
 By specifying a prior distribution over the model parameters $p(\boldsymbol{\theta})$, one can generate arbitrarily large datasets of the form $\boldsymbol{D} = \{\boldsymbol{x}^{(i)}, \boldsymbol{\theta}^{(i)}\}_{i=1}^{n}$ and use a deep generative model to learn a probabilistic mapping from data to parameters.
-Thus, during the inference phase, the model generates samples $\boldsymbol{\theta}^{(1)}, \boldsymbol{\theta}^{(2)},...,\boldsymbol{\theta}^{(L)}$ from the observed data $\boldsymbol{x}_{obs}$, approximating the target posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$.
+Thus, during the inference phase, the model generates samples $\boldsymbol{\theta}^{(1)}, \boldsymbol{\theta}^{(2)},...,\boldsymbol{\theta}^{(L)}$ given the observed data $\boldsymbol{x}_{obs}$, approximating the target posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs})$.
 
 In the current work, we propose to implement and use a conditional invertible neural network (cINN) architecture.
-Previously, INNs have been successfully employed to model data from astrophysics and medicine\cite{ardizzone2018analyzing}.
+Previously, INNs have been successfully employed to model data from astrophysics and medicine \cite{ardizzone2018analyzing}.
 We adapt the model to suit the task of parameter estimation in the context of mathematical modeling (see Figure 1 for a full graphical illustration of the method) and develop a reusable probabilistic architecture for full Bayesian likelihood-free inference on complex mathematical models.
@@
 Each cACB consists of four separate fully connected neural networks denoted as $s_{1}(\cdot), s_{2}(\cdot), t_{1}(\cdot), t_{2}(\cdot)$.
-An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ it also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ "for free".
+An ACB is specifically designed to be invertible, which means that in addition to a parametric mapping $f_{\boldsymbol{\phi}_{inv}}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ it also learns the inverse mapping $f_{\boldsymbol{\phi}_{inv}}^{-1}: \mathbb{R}^{d} \rightarrow \mathbb{R}^{d}$ \textit{for free}.
 Denoting the input vector of $f_{\boldsymbol{\phi}_{inv}}$ as $\boldsymbol{u}$ and the output vector as $\boldsymbol{v}$, it follows that $f(\boldsymbol{u}; \boldsymbol{\phi}_{inv}) = \boldsymbol{v}$ and $f^{-1}(\boldsymbol{v}; \boldsymbol{\phi}_{inv}) = \boldsymbol{u}$.
@@
 Therefore, instead of feeding the raw simulated or observed data to each ACB, we pass the data through an additional summary network to obtain a fixed-sized vector of learned summary statistics $\tilde{\boldsymbol{x}} = h(\boldsymbol{x};\boldsymbol{\phi}_{sum})$ and learn the parameters of the summary network $h$ jointly with those of the cINN chain via backpropagation.
-Thus, the current method remains completely end-to-end and is capable of generalizing to data sets of variable input size and structure.
+Thus, the current method remains completely end-to-end and is capable of generalizing to datasets of variable input sizes and structure.
 
@@
 where $\partial f/\partial \boldsymbol{\theta} = \boldsymbol{J}_{f}$ is the Jacobian of the learned transformation $f(\boldsymbol{\theta};\boldsymbol{x},\boldsymbol{\phi})$ with respect to the input.
-Both terms in Eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-2/d}\exp(\norm{\boldsymbol{z}}_{2}^{2})$ and the determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
+Both terms in Eq.\ref{eqn:11} are now tractable, since we have previously defined $\boldsymbol{z}$ as following a spherical unit Gaussian, that is, $p(\boldsymbol{z}) = (2\pi)^{-d/2}\exp(\norm{-\frac{1}{2}\boldsymbol{z}}_{2}^{2})$ and the determinant of the Jacobian is easily computed as $s_{1}(\boldsymbol{u}_{2}, \boldsymbol{x}) + s_{2}(\boldsymbol{v}_{1}, \boldsymbol{x})$ due to eqs.
 \ref{eqn:2} and \ref{eqn:3}.
@@
 &= -\frac{1}{m}\sum_{i=1}^{m}\log\left(p(\boldsymbol{\theta}^{(i)}|\boldsymbol{x}^{(i)},\boldsymbol{\phi})p(\boldsymbol{\phi})\right) \numberthis \label{eqn:13} \\
-&= -\frac{1}{m}\sum_{i=1}^{m}\left(\frac{\norm{f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi})}_{2}^{2}}{2} + \log\left|\det\left(\boldsymbol{J}_{f}^{(i)}\right)\right|\right) + \tau\norm{\boldsymbol{\phi}}_{2}^{2} \numberthis \label{eqn:14}
+&= \frac{1}{m}\sum_{i=1}^{m}\left(\frac{\norm{f(\boldsymbol{\theta}^{(i)};\boldsymbol{x}^{(i)},\boldsymbol{\phi})}_{2}^{2}}{2} - \log\left|\det\left(\boldsymbol{J}_{f}^{(i)}\right)\right|\right) + \tau\norm{\boldsymbol{\phi}}_{2}^{2} \numberthis \label{eqn:14}
 \end{align*}
@@
 Note that our formulation maximizes the posterior of model parameters directly, in contrast to variational methods which optimize a lower bound on the posterior \cite{papamakarios2016fast, kingma2014auto}.
-Once the backpropagation algorithm has settled to a local minimum of the ML loss, one can easily obtain samples from the approximate posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs},\boldsymbol{\phi}=\boldsymbol{\phi}^{*})$, based on an observed dataset $\boldsymbol{x}_{obs}$, by repeatedly sampling $\boldsymbol{z}^{(l)} \sim \mathcal{N}_{d}(\boldsymbol{0},\boldsymbol{I})$, and then passing $\boldsymbol{z}^{(l)}$ in reverse to the cINN in order to compute $\boldsymbol{\theta}^{(l)} = f^{-1}(\boldsymbol{z}^{(l)};\boldsymbol{x}_{obs},\boldsymbol{\phi}^{*})$ for $l=1,...,L$.
+Once the backpropagation algorithm has settled to a local minimum of the ML loss, one can easily obtain samples from the approximate posterior $p(\boldsymbol{\theta}|\boldsymbol{x}=\boldsymbol{x}_{obs},\boldsymbol{\phi}=\boldsymbol{\phi}^{*})$, based on an observed dataset $\boldsymbol{x}_{obs}$.
+This is achieved by repeatedly sampling $\boldsymbol{z}^{(l)} \sim \mathcal{N}_{d}(\boldsymbol{0},\boldsymbol{I})$, and then passing $\boldsymbol{z}^{(l)}$ in reverse to the cINN in order to compute $\boldsymbol{\theta}^{(l)} = f^{-1}(\boldsymbol{z}^{(l)};\boldsymbol{x}_{obs},\boldsymbol{\phi}^{*})$ for $l=1,...,L$.
 Figure 1b illustrates the inference phase of the method.
@@
 With these settings, we perform $50\ 000$ to $100\ 000$ iterations (network updates) for each of the examples in this paper, and report the results on the trained network.
-Note, that we did not perform an extensive search for optimal values of the cINN hyperparameter, but use a cINN with 10 ACBs all examples in this paper (see \textbf{SI} for details of the cINNs and summary networks).
+Note, that we did not perform an extensive search for optimal values of the cINN hyperparameters, but use a cINN with 10 ACBs for all examples in this paper (see \textbf{SI} for details of the cINNs and summary networks).
 All networks were implemented in Python using the \textit{TensorFlow} library \cite{abadi2016tensorflow} and trained on a single-GPU machine equipped with NVIDIA\textsuperscript{\textregistered} GTX1060 graphics card.
@@
 
-The second approach follows a different strategy, which resembles ideas from \textit{active learning} \cite{mnih2015human}.
-Correspondingly, a dataset, or a batch of datasets, is created on the fly and then passed through the neural network.
-This training regime has the advantage that the network never \textit{experiences} the same input data twice.
+The second approach follows a different strategy, which resembles ideas from \textit{online learning} \cite{mnih2015human} where data are stimulated by $S(\boldsymbol{\theta},\xi)$ on demand.
+Correspondingly, a dataset, or a batch of datasets, is generated on the fly and then passed through the neural network.
+This training approach has the advantage that the network never \textit{experiences} the same input data twice.
 Moreover, training can continue as long as the network keeps improving (i.e., the loss keeps decreasing), since overfitting in the classical sense is nearly impossible.
-However, if simulations are computationally expensive and researchers need to experiment with different networks or training hyperparameters, it might be beneficial to switch to the first regime, since simulation and training in the active learning regime are tightly intertwined.
+However, if simulations are computationally expensive and researchers need to experiment with different networks or training hyperparameters, it might be beneficial to switch to the first approach, since simulation and training in online learning are tightly intertwined.
 
@@
 
-On an abstract level, our method requires three key ingredients: 1) a mathematical process model $q(\boldsymbol{\theta})$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\phi}}$ capable of approximating a large enough family of probability distributions (see Figure 2).
+On an abstract level, our method requires three key ingredients: 1) a mathematical process model $S(\boldsymbol{\theta},\xi)$ capable of simulating data $\boldsymbol{x}$; 2) a prior distribution over the model parameters $p(\boldsymbol{\theta})$ encoding our prior beliefs about plausible parameter values; 3) an invertible neural network $f_{\boldsymbol{\phi}}$ capable of approximating a large enough family of probability distributions (see \autoref{fig:Fig.1}).
 In practice, a chain of up to 10 ACBs should suffice to learn most distributions encountered in the cognitive or life sciences, since they tend to be unimodal and relatively simple (in contrast to the distributions required to represent images or words in a spoken language).
 From these three ingredients, a universal and reusable sampler can be designed for likelihood-free Bayesian estimation of both tractable and intractable mathematical models.
-\textbf{Algorithm} \ref{alg:1} describes the essential steps of the method using an arbitrary summary network and employing the active learning training regime.
+\textbf{Algorithm} \ref{alg:1} describes the essential steps of the method using an arbitrary summary network and employing the online learning training regime.
 
@@
 \begin{algorithmic}[1]
-\State \emph{Training (via active learning):}
+\State \emph{Training (via online learning):}
 \Repeat
 \State {Sample a batch of $\{\boldsymbol{\theta}^{(i)}\}_{i=1}^{m}$ from prior $p(\boldsymbol{\theta})$}
-\State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = q(\boldsymbol{\theta}^{(i)})$ for $i=1,...,m$}
+\State {Simulate a batch of datasets $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ by running $\boldsymbol{x}^{(i)} = S(\boldsymbol{\theta}^{(i)},\xi)$ for $i=1,...,m$}
 \State {Pass $\{\boldsymbol{x}^{(i)}\}_{i=1}^{m}$ through summary network $h(\boldsymbol{x}^{(i)};\boldsymbol{\phi}_{sum})$ to obtain $\{\tilde{\boldsymbol{x}}^{(i)}\}_{i=1}^{m}$}
@@
 
-To assess the performance of our method on the following application examples, we consider a number of different metrics.
+To evaluate the performance of our method on the following application examples, we consider a number of different metrics.
 To assess the precision of point estimates (posterior means), we compute the normalized root mean squared error (NRMSE) and the coefficient of determination ($R^{2}$) between estimated and true parameter values.
@@
 Training the networks took approx.
-half a day with the active learning approach.
+half a day with the online learning approach.
 Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
@@
 x_{t} &\sim Pois(\rho N_{t}) \numberthis \\
-N_{t+1} &= rN_{t}e^{-N_{t} + \epsilon_{t}} \numberthis 
+N_{t+1} &= rN_{t}e^{-N_{t} + \xi_{t}} \numberthis 
 \end{align*}
-for $t = 1,...,T$ where $N_{t}$ is the expected number of individuals at time $t$, $r$ is the growth rate, $\rho$ is a scaling parameter and $\epsilon_{t} \sim \mathcal{N}(0, \sigma^{2})$ is random Gaussian noise \cite{mestdagh2018prepaid}.
+for $t = 1,...,T$ where $N_{t}$ is the expected number of individuals at time $t$, $r$ is the growth rate, $\rho$ is a scaling parameter and $\xi_{t} \sim \mathcal{N}(0, \sigma^{2})$ is random Gaussian noise \cite{mestdagh2018prepaid}.
 The likelihood function for the Ricker model is not available in closed-form, and the model is known to exhibit chaotic behavior.
@@
 
-During training of the networks, we simulate time-series from the Ricker model with varying lengths.
+For training the networks, we simulate time-series from the Ricker model with varying lengths.
 The number of time points $T$ is drawn from a uniform distribution $T \sim \mathcal{U}(100, 500)$ at each training iteration.
 Training the networks took approx.
-half a day with the active learning approach.
+half a day with the online learning approach.
 Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
@@
 Here, we apply our method to estimate the parameters of the recently proposed Lévy-Flight Model (LFM, \cite{voss2019sequential}).
-The LFM assumes an \textit{alpha-stable} noise distribution of the evidence accumulation process in order to model "jumps" in the decision process.
+The LFM assumes an \textit{alpha-stable} noise distribution of the evidence accumulation process in order to model \textit{jumps} in the decision process.
 However, the inclusion of \textit{alpha-stable} noise leads to a model with intractable likelihood; further, to our knowledge, a fully Bayesian treatment of the model is still missing in the literature.
@@
 The number of trials is drawn from a uniform distribution $n \sim \mathcal{U}(100, 1000)$ at each training iteration.
-Training the networks took a little less than a day with the active learning approach.
+Training the networks took a little less than a day with the online learning approach.
 Inference on $1000$ datasets with $2000$ posterior samples per parameter took approx.
@@
 Finally, the SCB histograms indicate that our method tends to slightly underestimate the true posterior variance of the $\alpha$, $\beta$, $\mu^{O}$ $\pi^{O}$, and $\sigma^{O}$ parameters.
-This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear if the active learning approach is employed.
+This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear if the online learning approach is employed.
 
@@
 We demonstrated the utility of the method by applying it to models and data from various scientific domains.
-Further, we explored two possible training approaches suitable for different simulation scenarios, namely, an active learning approach, and a reference-table approach.
+Further, we explored two possible training approaches suitable for different simulation scenarios, namely, an online learning approach, and a reference-table approach.
 Both training approaches lead to excellent parameter estimation throughout the examples considered in the current work.
@@
 Our results support this theoretical claim.
-Further, since researchers are often interested in some summary of the posterior (e.g., posterior means or variances), we also showed that our method exhibits excellent recovery of the posterior means throughout all examples.
-Further, parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters decreases.
-These are important and highly desirable properties of any Bayesian parameter estimation method, as it mirrors the increase in information following increasing number of observations.
+Additionally, since researchers are often interested in some summary of the posterior (e.g., posterior means or variances), we also showed that our method exhibits excellent recovery of the posterior means throughout all examples.
+Further, we observe the desired posterior contraction (i.e., parameter recovery becomes better with increasing number of observations, whereas the posterior variance of the parameters decreases).
+These are indispensable properties of any Bayesian parameter estimation method, as it mirrors the increase in information following increasing number of observations.
 

### 2019-07-17 latex/paper_supplementary.tex
Add new version of text

@@
 \includegraphics[width=\textwidth]{Regression_kl.png}
-\caption{KL divergence ($D_{KL}$) between the true and estimated posteriors of the regression weights in our toy Bayesian regression example.
+\caption{KL divergence ($D_{KL}$) between the true and estimated posteriors of the regression coefficients in our toy Bayesian regression example.
 The $D_{KL}$ is depicted at every $n$ observations used during training.}
@@
 We use a bidirectional long short-term memory (LSTM) recurrent neural network \cite{gers1999learning} for summarizing the Ricker time-series into fixed-size vectors.
-The LSTM network architecture is a good choice for this example, as it is able to capture long-term in datasets with temporal or spatial autocorrelations.
-LSTMs can also easily deal with variable length time-series.
+The LSTM network architecture is a reasonable choice for this example, as it is able to capture long-term dependencies in datasets with temporal or spatial autocorrelations.
+LSTMs can also easily deal with variable-length time-series.
 
@@
 \end{align*}
-These ranges are very broad, as datasets generated by extreme parameter values appear implausible in real-world scenarios.
+These ranges appear to be very broad, as datasets generated by extreme parameter values appear implausible in real-world scenarios.
 Nevertheless, we stick to broad priors for training, even though parameter recovery might degrade at the extremes.
-Figure XXX depicts some Ricker datasets generated by parameters drawn from the specified prior.
 
@@
 \end{align*}
-Figure XXX depicts some SIR datasets generated by parameters drawn from the specified priors.
 

### 2019-07-17 splatter_test.R

@@ -1,0 +1,78 @@
+library(splatter)
+library(RcppCNPy)
+
+setwd("D:/seqRNA")
+
+
+
+simulate_batch = function(sim_id, n_sim=32, n_genes=20, n_cells=20, test=F, n_params=8) {
+  # Simulates a batch of gene-cell matrices and save it to file
+  
+  if (test) set.seed(42)
+  else set.seed(sim_id)
+    
+  sim_data = array(0, c(n_sim, n_genes, n_cells))
+  sim_params = array(0., c(n_sim, n_params))
+  
+  # Sample parameters 
+  mu = runif(n_sim, 1.0, 2.0)
+  rate = runif(n_sim, 0.1, 5.0)
+  out_loc = runif(n_sim, 0.5, 5)
+  out_scale = runif(n_sim, 0.1, 2.0)
+  lib_loc = runif(n_sim, 6, 10)
+  lib_scale = runif(n_sim, 1.0, 2.0)
+  out_prob = runif(n_sim, 0.01, 0.5)
+  bcv_common = runif(n_sim, 0.1, 0.9)
+  
+  for (i in 1:n_sim) {
+    
+    
+    # Generate a parameter instance
+    params = newSplatParams(mean.shape=mu[i], 
+                            mean.rate=rate[i], 
+                            out.facLoc=out_loc[i], 
+                            out.facScale=out_scale[i], 
+                            lib.loc=lib_loc[i], 
+                            lib.scale=lib_scale[i], 
+                            out.prob=out_prob[i],
+                            bcv.common=bcv_common[i])
+    
+    # Simulate and store
+    sim = splatSimulate(params, nGenes = n_genes, batchCells=n_cells, verbose=F)
+    sim_data[i, , ] = matrix(counts(sim), nrow=n_genes, ncol=n_cells)
+    sim_params[i, ] = c(mu[i], rate[i], out_loc[i], out_prob[i], 
+                        out_scale[i], lib_loc[i], lib_scale[i], bcv_common[i])
+  }
+  
+  
+  
+  if (test) {
+    npySave("test/rna_data_test.npy", sim_data)
+    npySave("test/rna_params_test.npy", sim_params)
+  } else {
+    npySave(paste0("data/rna_data_", sim_id + offset, ".npy"), sim_data)
+    npySave(paste0("params/rna_params_", sim_id + offset, ".npy"), sim_params)
+  }
+  
+  
+}
+
[…]

### 2019-07-18 latex/paper_main.tex
Add seqRNA example

@@
 
-\title{Learning complex stochastic models with invertible neural networks: a likelihood-free approach to Bayesian inference}
+\title{Learning complex stochastic models with invertible neural networks: a new approach to Bayesian inference}
 
@@
 However, in many applications, finding an optimal set of summary statistics is far from obvious.
+This also applies to the models considered in this work.
 The summary network eliminates the need to manually decide on and compute a fixed set of summary statistics and, thus, makes the method independent of the format or the size of the data.
@@
 Moreover, the observed data exhibits multiple local and non-local dependencies.
-We simulate $200\ 000$  \textit{Gene x Cell} count matrices with different parameter settings.
+We simulate $250\ 000$  \textit{Gene x Cell} count matrices with different parameter settings.
 Each entry in a count matrix contains the number of times a gene has been sequenced in a different cell.
@@
 $1.7$ seconds.
+To our knowledge, this is the first Bayesian treatment of the \textit{Splat} model.
 \begin{figure}[H]
@@
 \caption[short]{Results on the \textit{Splat} scRNA simulation model.
-\textbf{(a)} The \textit{Splat} simulation diagram; \textbf{(b)} Parameter recovery; \textbf{(c)} Simulation-based calibraton.
-The SCB plots indicate a slight underestimation of the true posterior variance.
-\textbf{(d)} Example full parameter posteriors from three datasets.} \label{fig:Fig.7}
+\textbf{(a)} The \textit{Splat} simulation diagram; \textbf{(b)} Parameter recovery; \textbf{(c)} Histograms of the rank statistic for simulation-based calibration (SCB); \textbf{(d)} Example full parameter posteriors from three datasets.} \label{fig:Fig.7}
 \end{figure}
 The results on the scRNA example are depicted in \autoref{fig:Fig.7}.
-First, recovery of most parameters is very good, with maximum $R^{2}$ of $0.98$ and minimum NRMSE of $0.039$ for the library location parameter $\mu^{L}$.
-Interestingly, the rate parameter $\beta$ appears to be nearly unrecoverable, as indexed by the scatterplots in \autoref{fig:Fig.7b}.
-Furthermore, the estimated posterior of $\beta$ is often bimodal, covering the whole prior range (\autoref{fig:Fig.7c}).
-This might indicate that the parameter is not identifiable, or that some multivariate trade-offs exist in the parameter space of the \textit{Splat} model.
-The latter is also confirmed by observing a few marked outliers in the scatterplots of $\alpha$, $\sigma^{O}$, and $\pi^{O}$.
-Finally, the SCB histograms indicate that our method tends to slightly underestimate the true posterior variance of the $\alpha$, $\beta$, $\mu^{O}$ $\pi^{O}$, and $\sigma^{O}$ parameters.
-This issue might be due to the fact that the networks were trained with the reference-table approach and might disappear if the online learning approach is employed.
+First, recovery of most parameters is very good, with maximum $R^{2}$ of $0.999$ and minimum NRMSE of $0.01$ for the library location parameter $\mu^{L}$.
+Interestingly, the rate parameter $\beta$ appears to be unrecoverable, as indexed by the scatterplots in \autoref{fig:Fig.7b}.
+Furthermore, the estimated posterior of $\beta$ is approximately uniform, covering the entire prior range (\autoref{fig:Fig.7c}).
+This indicates that no information about the parameter is present in the data.
+We also observe some multivariate trade-offs in the parameter space suggested by the marked outliers in the scatterplots of $\mu^{O}$ and $\sigma^{O}$.
+Finally, the SCB histograms indicate no systematic bias in the approximate posteriors across all parameters.
+The posterior mean of $\alpha$ appears to be slightly overestimated, whereas the opposite trend is present for $\pi^{O}$ and $\sigma^{L}$.
+These inaccuracies might be due to the fact that the networks were trained with the reference-table approach and might disappear by employing the online learning approach.
 

### 2019-07-18 latex/paper_supplementary.tex
Add seqRNA example

@@
 \subsubsection*{Summary network}
-We use the recently introduced permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
+We use the recently introduced permutation invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} regression data.
 This network architecture is a good choice for this example, as each observation is independent of all other, and thus all permutations of a given dataset should lead to the same parameter estimates.
-Furthermore, invariant networks are independent of the number of observations.
+Furthermore, permutation invariant networks are independent of the number of observations.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Ricker_plot_multiple.png}
+\caption{Some example Ricker time-series generated with different parameters drawn from the prior.} 
+\end{figure}
+
 \subsection*{The Lévy-Flight Model}
@@
 \subsubsection*{Summary network}
-We use a permutationally invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
+We use a permutation invariant neural network \cite{bloem2019probabilistic} for the \textit{i.i.d.} reaction times (RT) data.
 Similarly to the toy Regression example, each response in an RT dataset is assumed to be independent of all others, so permutations of the dataset must lead to the same parameter estimates.
@@
 These priors are broad enough to cover the range of realistic reaction times observed in choice RT experiments.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Levy_plot_multiple.png}
+\caption{Some example Lévy-Flight datasets generated with different parameters drawn from the prior.} 
+\end{figure}
 
@@
 \end{align*}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{SIR_plot_multiple.png}
+\caption{Some example SIR time-series generated with different parameters drawn from the prior.} 
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Splatter_plot_multiple.png}
+\caption{A heatmap of a single $Genes \star Cells$ count matrix generated with different parameters drawn from the prior.} 
+\end{figure}
+
 %%% Add this line AFTER all your figures and tables
 \FloatBarrier
-
-\movie{Type caption for the movie here.}
 

### 2019-07-18 process_models.py

@@ -1,9 +1,21 @@
 from numba import jit, prange
+import ctypes
+from numba.extending import get_cython_function_address
 import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
 import scipy
 import tensorflow as tf
+
+
+# Get a pointer to the C function diffusion.c
+addr_diffusion = get_cython_function_address("diffusion", "diffusion_trial")
+functype = ctypes.CFUNCTYPE(ctypes.c_double, ctypes.c_double, ctypes.c_double, 
+                            ctypes.c_double, ctypes.c_double, ctypes.c_double,
+                            ctypes.c_double, ctypes.c_double, ctypes.c_double, ctypes.c_double,
+                            ctypes.c_int)
+
+diffusion_trial = functype(addr_diffusion)
 
 
 @jit(nopython=True)
@@ -73,7 +85,7 @@ def simulate_ricker(batch_size=64, n_points=None, t_obs_min=100, t_obs_max=500, 
     simulate_ricker_batch(X, theta, batch_size, n_points)
     if to_tensor:
         return tf.convert_to_tensor(X[:, :, np.newaxis], dtype=tf.float32), tf.convert_to_tensor(theta, dtype=tf.float32)
-    return X[:, :, np.newaxis]
+    return X[:, :, np.newaxis], theta
 
 
 def sir(u, beta, gamma, t, N=1000, dt=0.1, iota=0.5):
@@ -151,6 +163,72 @@ def simulate_sir(batch_size, n_points=None, low_beta=0.01, high_beta=1., low_gam
     return X, theta
 
 
+@jit(nopython=True, cache=True, parallel=True)
+def simulate_batch_diffusion_p(x, params):
+    """
+    Simulate a batch from the diffusion model.
+    ----------
+    INPUT:
+    x - np.array of shape (n_batch, n_trials, n_cond)
+    params - np.array of shape (n_batch, n_params):
+        param index 0 - drift_rate1
+        param index 1 - drift_rate2
+        param index 2 - variability in drift rate
+        param index 3 - relative starting point
+        param index 4 - variability in starting point
+        param index 5 - threshold
+		param index 6 - NDT
+		param index 7 - variability in NDT
+        param index 8 - alpha
+    """
+
+    # For each batch
+    for i in prange(x.shape[0]):
+        # For each condition
+        for j in prange(x.shape[1]):
[…]

### 2019-07-18 simulate_diffusion.py

@@ -1,6 +1,5 @@
 from numba import jit, prange
-import ctypes
-from numba.extending import get_cython_function_address
+
 import tensorflow as tf
 import numpy as np
 
@@ -79,3 +78,31 @@ def simulate_diffusion(batch_size, pbounds, n_points=None, n_cond=2,
     if to_tensor:
         X_batch, theta_batch = tf.convert_to_tensor(X_batch, dtype=tf.float32), tf.convert_to_tensor(theta_batch, dtype=tf.float32)
     return X_batch, theta_batch
+
+
+def plot_diffusion_multiple(n=1000, figsize=(20, 20), filename='levy'):
+    """Plots example datasets from the SIR model."""
+
+    X, theta = simulate_diffusion(25, n_points=n, to_tensor=False)
+    n = np.arange(1, n+1)
+
+    f, axarr = plt.subplots(5, 5, figsize=figsize)
+
+    for i, ax in enumerate(axarr.flat):
+
+        # Plot just a single condition
+        sns.distplot(X[i, :, 0], ax=ax, label='Upper threshold', rug=True, color="#4873b8")
+
+        if i == 0:
+            ax.set_xlabel(r'Number of trials ($n$)', fontsize=10)
+            ax.set_ylabel('Density', fontsize=10)
+            ax.legend(fontsize=10)
+        else:
+            ax.get_legend().remove()
+        ax.spines['right'].set_visible(False)
+        ax.spines['top'].set_visible(False)
+    
+    f.tight_layout()
+
+    if filename is not None:
+        f.savefig("figures/{}_levy_multiple.png".format(filename), dpi=600, bbox_inches='tight')
